---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'Trail Making Test'"
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
# bibliography: "`r ecs.data::get_proj_bib_file()`"
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib
csl: "`r ecs.data::get_proj_csl_file()`"
# nocite: '@<cita_1>, @<cita_2>' ## TODO: Descomentar y añadir cita(s) o borrar campo
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

# Opciones de configuración de las salidas de R y Stata:

options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( # Para todos los 'chunks' de código:
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

| Main variable name | Label                                            | Format       | Code |
|:-------------------|:-------------------------------------------------|:-------------|:-----|
| tmt_time_a\_s3     | TMT completion time (seconds), part A (c2011_w4) | Cuantitative | 0+   |
| tmt_errors_a\_s3   | TMT errors, part A (c2011_w4)                    | Cuantitative | 0+   |
| tmt_time_b\_s3     | TMT completion time (seconds), part B (c2011_w4) | Cuantitative | 0+   |
| tmt_errors_b\_s3   | TMT errors, part A (c2011_w4)                    | Cuantitative | 0+   |
| tmt_ratio_s3       | Trail making ratio (seconds) (c2011_w4)          | Cuantitative | 0+   |

<!--# TODO: Añadir variables outcome a la tabla -->

# Variable description

The Trail Making Test is a neuropsychological assessment tool used to evaluate a
person's visual-motor processing speed, attention, and executive function. This
version is adapted from a neuropsychological battery [@benedet_evaluacion_2013],
and has been previously validated in Spanish in a sample of adult argentinian
population [@fernández2002]. Herein we apply the same scoring and application
method, which consists of correcting the errors the participant makes in
application time; as a result, the errors are taken into account in the
completion time [@reitan1971; @reitan1958], although the method has been
criticized for its dependence of the examiner's response time [@fernández2002].

The test consists of two parts, A and B. Part A requires the person to connect a
series of 25 numbered circles, arranged randomly on a sheet of paper, in
ascending order as quickly as possible. Part B is similar, but it includes both
numbered circles (1-13) and letter circles (A-L), and the person must alternate
between connecting the two types of circles in ascending order (1-A-2-B-3-C,
etc.).

The time taken to complete each part is recorded, and the difference between the
times for Part B and Part A (known as the "Trail Making Ratio") can provide
information about the person's cognitive flexibility and ability to shift
between different mental tasks.

# Explanation of how the variables were created

Missing value categories from the two time variables (`q2549_sequence1_time_s3`
and `q2552_sequence2_time_s3`) are recoded as system missing values, prior to
renaming them as their corresponding outcomes (`tmt_time_a_s3` and
`tmt_time_b_s3`, respectively). Variable `tmt_ratio_s3` is computed as the
difference of the two previous ones, i.e. `tmt_time_b_s3` - `tmt_time_a_s3`.

The two "number of error" variables are simply renamed as their corresponding
outcomes (`tmt_errors_a_s3` and `tmt_errors_a_s3`, respectively).

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* # Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
* "Recode the missing value categories as system missing values:"
recode q2549_sequence1_time_s3 q2552_sequence2_time_s3 (888=.) (999=.)
label values q2549_sequence1_time_s3
label values q2552_sequence2_time_s3

rename q2549_sequence1_time_s3  tmt_time_a_s3
rename q2550_sequence1_error_s3 tmt_errors_a_s3
rename q2552_sequence2_time_s3  tmt_time_b_s3
rename q2553_sequence2_error_s3 tmt_errors_b_s3

gen tmt_ratio_s3 = tmt_time_b_s3 - tmt_time_a_s3

label variable tmt_time_a_s3   "TMT completion time (seconds), part A (c2011_w4)"
label variable tmt_errors_a_s3 "TMT errors, part A (c2011_w4)"
label variable tmt_time_b_s3   "TMT completion time (seconds), part B (c2011_w4)"
label variable tmt_errors_b_s3 "TMT errors, part B (c2011_w4)"
label variable tmt_ratio_s3    "Trail making ratio (seconds) (c2011_w4)"
```

<br>

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result_s3 proxy_s3 tmt_time_a_s3 tmt_errors_a_s3 tmt_time_b_s3 tmt_errors_b_s3 tmt_ratio_s3

codebook tmt_time_a_s3 tmt_errors_a_s3 tmt_time_b_s3 tmt_errors_b_s3 tmt_ratio_s3
```

```{stata save-data-test, eval=TEST}
saveold "Outcome_trailmakingtest.dta", replace version(13)
```

```{stata save-data, eval=!TEST}
* # Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_trailmakingtest.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Daniel Morillo Cuadrado
-   e-mail:
    [\<daniel.morillo\@inv.uam.es\>](mailto:daniel.morillo@inv.uam.es){.email}
-   Phone number: (+34) 91 497 58 16

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
