---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'cognitive decline (IQCODE)'"
output:
  word_document:
   reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib 
csl: "`r ecs.data::get_proj_csl_file()`"
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"


options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set(
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

| Main variable name | Label                              | Format     | Code |
|:-------------------|:-----------------------------------|:-----------|:-----|
| IQCODE_s3          | Cognitive decline level (c2011_w4) | Continuous | 0-5  |

# Variable description

Cognitive decline refers to the impairment of cognitive functions such as memory
or information processing, to a greater extent than natural age-related decline.
It may appear suddenly or gradually and can be permanent or temporary
[@aarsland2017]. The IQCODE is a questionnaire which asks an informant about
changes in an elderly person's everyday cognitive functions. The questionnaire
aims to assess cognitive decline independent of premorbid ability
[@forcanogarcía2002]. Using the IQCODE allow us to assess possible changes in
the functioning of cognitive functions over time.

# Explanation of how the variables were created

We used the Short Form of the Informant Questionnaire on Cognitive Decline
(SF-IQCD) including 16 questions about remember things about family
(p1021_family_s3), recent events (p1022_recent_s3), past conversations
(p1023_recall_s3), personal infromations (p1024_address_s3), localisation
(p1025_day_s3), location of things (p1026_kept_s3), find things (p1027_find_s3),
use of household machines (p1028_machin_s3), use of new gadget
(p1029_gadget_s3), learn new things (p1030_newthin_s3), read a book
(p1031_book_s3), take decision (p1032_decisio_s3), manege money
(p1033_money_s3), finance (p1034_financial_s3), daily arithmetic problems
(p1035_other_s3 and p1036_reason_s3). All the items previously cited were
assessed on a scale ranging from 1 (much better functioning of cognitive
functions) to 5 (much worst functioning of cognitive functions over time).

To assess the cognitive decline through the IQCODE, we computed the mean of the
16 items by adding up the score for each question and divide by the number of
questions (16). The result is a score that ranges from 1 to 5. Higher scores
indicate greater impairment, 3 = any changes, and 1 and 2 = improvement of
cognitive functions.

Note: An average score above 3,44 may indicate that participant is suffering
from demencia (see [@forcanogarcía2002]).

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
recode  p1021_family_s3 p1022_recent_s3 p1023_recall_s3 p1024_address_s3 p1025_day_s3 p1026_kept_s3 p1027_find_s3 p1028_machin_s3 p1029_gadget_s3 p1030_newthin_s3 p1031_book_s3 p1032_decisio_s3 p1033_money_s3 p1034_financial_s3 p1035_other_s3 p1036_reason_s3 (888=.) (999=.)

gen IQCODE_s3 = (p1021_family_s3 + p1022_recent_s3 + p1023_recall_s3 + p1024_address_s3 + p1025_day_s3 + p1026_kept_s3 + p1027_find_s3 + p1028_machin_s3 + p1029_gadget_s3 + p1030_newthin_s3 + p1031_book_s3 + p1032_decisio_s3 + p1033_money_s3 + p1034_financial_s3 + p1035_other_s3 + p1036_reason_s3)/16

label variable IQCODE_s3 "Cognitive decline level (c2011_w4)"

```

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result proxy IQCODE_s3

codebook IQCODE_s3
```

```{stata save-data-test, eval=TEST}
saveold "Outcome_cognitivedecline.dta", replace version(13)
```

```{stata save-data, eval=!TEST}

* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_cognitivedecline.dta", replace version(13)

```

# Contact person:

-   Name: Lea Francia
-   e-mail: [lea.francia\@inv.uam.es](mailto:lea.francia@inv.uam.es){.email}
-   Phone number: (+34) 914974601

<br>

\< Original version:

-   Name: Joan Domènech
-   e-mail: [\j.domenech\@pssjd.org\>](mailto:j.domenech@pssjd.org){.email}

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
