--- 
title: "Edad con Salud -- 2011 Cohort -- Wave 4 -- 'Positive and Negative Affect'"
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: ../../../../www/EcS_DRM.bib 
csl: "`r ecs.data::get_proj_csl_file()`"
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
--- 

```{r setup, include=FALSE}

library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"

options(width = 88) 
knitr::opts_chunk$set( 
  echo        = FALSE, 
  results     = 'hide',
  cleanlog    = TRUE,  
  collectcode = TRUE,  
  comment     = ''    
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

+--------------------+------------------------------------------+------------+----------------------+
| Main variable name | Label                                    | Format     | Code                 |
+:===================+:=========================================+:===========+:=====================+
| positiveaffect_s3  | Positive affect (DRM) (c2011w4)          | Continuous | 0-6                  |
+--------------------+------------------------------------------+------------+----------------------+
| negativeaffect_s3  | Negative affect (DRM) (c2011_w4)         | Continuous | 0-6                  |
+--------------------+------------------------------------------+------------+----------------------+
| netaffect_s3       | Net affect (DRM) (c2011_w4)              | Continuous | -6 - +6              |
+--------------------+------------------------------------------+------------+----------------------+

# Variable description

The Day Reconstruction Method (DRM) (@kahneman_survey_2004,
@kahneman_developments_2006) measures how people spend their time and how they
emotionally experience the various activities of life. In the extended version,
it contains 12 items, including 4 positive and 8 negative. An example of a DRM
item follows: \_"Indicate to what extent you\_ *felt happy during specific
activities of your day"*. Response options are conformed by a 7-point Likert
scale ranging from 0,"not at all" to 6, "very much".

In Aging with Health, respondents were asked to complete an abbreviated version
of the DRM (@miret_validation_2012, @ayuso-mateos_multi-country_2013,
@caballero_evaluation_2014, @mellor-marsa_measurement_2016). Each participant
provided information about the activities conducted the previous day and rated
the emotions experienced during these activities. By means of Computer-Assisted
Personal Interviewing, one activity was randomly selected for each time frame
(morning, afternoon and evening).

-   **Positive affect** refers to emotions of happiness, warminess, friendliness
    and enjoyment with myself.
-   **Negative affect** is defined by emotions of frustration, annoyment,
    depression, feeling blue, hassled/pushed around, angry/hostile,
    worried/anxious, criticized/put down.
-   **Net affect** is defined as the average of two positive emotions (calm/
    relaxed and enjoying) minus the average of the five negative ones (worried,
    rushed, irritated/angry, depressed, and tense/stressed), i.e., positive
    affect minus negative affect.
    
\newpage

# Explanation of how the variables were created

Three measures were obtained from the Day Reconstruction Method (DRM): *positive
affect*, *negative affect* and *net affect*.

-   **Positive affect**: the average of the two positive emotions (calm/relaxed
    and enjoying).

-   **Negative affect**: the average of the five negative ones (worried, rushed,
    irritated/angry, depressed, and tense/stressed).

-   **Net affect**: the average of the two positive emotions (calm/relaxed and
    enjoying) minus the average of the five negative ones (worried, rushed,
    irritated/angry, depressed, and tense/stressed), i.e., positive affect minus
    negative affect (Kahneman & Krueger, 2006). Individual net affect was
    calculated adding up the net-affect in each activity performed weighted by
    its duration. Net affect scores ranged from -6 to 6, with higher scores
    representing a better affective state.

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
* Duration of the trhee activities randomly chosen in the DRM at wave 4
gen duration_DRM4=q7026_s3 + q7066_s3 + q7111_s3
label variable duration_DRM4 "Total duration of the three activities randomly chosen (c2011w4)"


* Negative affect per activity
egen neg_activity1=rowmean(q7029_worr_s3-q7033_tense_s3)
egen neg_activity2=rowmean(q7069_worr_s3-q7073_tense_s3)
egen neg_activity3=rowmean(q7114_worr_s3-q7118_tense_s3)

* Positive affect per activity
egen pos_activity1=rowmean(q7034_calm_s3 q7035_enjoy_s3)
egen pos_activity2=rowmean(q7074_calm_s3 q7075_enjoy_s3)
egen pos_activity3=rowmean(q7119_calm_s3 q7120_enjoy_s3)

* Mean of negative and positive affect
gen negativeaffect_s3=((q7026_s3*neg_activity1) + (q7066_s3*neg_activity2) + (q7111_s3*neg_activity3))/duration_DRM4
gen positiveaffect_s3=((q7026_s3*pos_activity1) + (q7066_s3*pos_activity2) + (q7111_s3*pos_activity3))/duration_DRM4

replace negativeaffect_s3=((neg_activity1+neg_activity2+neg_activity3)/3) if duration_DRM4>1
replace positiveaffect_s3=((pos_activity1+pos_activity2+pos_activity3)/3) if duration_DRM4>1

*Net affect
gen netaffect_s3=positiveaffect_s3-negativeaffect_s3

label variable positiveaffect_s3 "Positive affect (DRM) (c2011w4) "
label variable negativeaffect_s3 "Negative affect (DRM) (c2011w4) "
label variable netaffect_s3 "Net affect (DRM) (c2011w4) "
```

<br>

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result proxy positiveaffect_s3 negativeaffect_s3 netaffect_s3

codebook positiveaffect_s3 negativeaffect_s3 netaffect_s3
```


```{stata save-data-test, eval=TEST}
saveold "Outcome_positive_negative_net_affects_DRM.dta", replace version(13)
```

```{stata save-data, eval=!TEST}
* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_positive_negative_net_affects_DRM.dta", replace version(13)
```

\newpage

# Contact person:

-   Name: Cristina Rodríguez Prada *(adapted the algorithm for wave 4)*
-   e-mail:
    [cristina.rodriguez\@inv.uam.es](mailto:cristina.rodriguez@inv.uam.es){.email}
-   Institution: Universidad Autónoma de Madrid / FIB-HU La Princesa
-   Phone number: (+34) 91 497 54 92

<br>

Original version:

-   Name: Félix Caballero
-   e-mail: [felix.caballero\@uam.es](mailto:felix.caballero@uam.es){.email}
-   Institution: Universidad Autónoma de Madrid
-   Phone number: (+34) 91 497 50 88

<br>

-   Name: Marta Miret
-   e-mail: [marta.miret\@uam.es](mailto:marta.miret@uam.es){.email}
-   Institution: Universidad Autónoma de Madrid
-   Phone number: (+34) 91 497 27 16

<br>

-   Name: Elvira Lara *(adapted the algorithm for wave 3)*
-   e-mail: [elvira.lara\@uam.es](mailto:elvira.lara@uam.es){.email}
-   Institution: Universidad Autónoma de Madrid / Universidad Complutense de Ma
    drid
-   Phone number: (+34) 91 497 46 01

# References

<!--# TODO: Borrar párrafo anterior -->

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
