---
title: "Edad con Salud -- cohort 2011 -- Wave 4 -- 'cognition'" 
output:
  word_document:
       reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib 
csl: "`r ecs.data::get_proj_csl_file()`"
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"

options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( # Para todos los 'chunks' de código:
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

| Main variable name | Label                                    | Format     | Code     |
|:-----------------|:--------------------------|:-----------------|:-----------------|
| cognition_std_s3   | Cognition score, standardized (c2011_w4) | continuous | standard |
| cognition_s3       | Cognition score, harmonized (c2011_w4)   | continuous | 0-100    |

<br>

# Variable description

Cognitive functioning was assessed by evaluating verbal fluency with the
animal-naming technique and immediate and delayed verbal recall was assessed
with the Consortium to Establish a Registry for Alzheimer Disease Word List
Memory [@moms1989]. Short-term memory was assessed with digit span backward and
forward tests from the Weschler Adult Intelligence Scale [@wechsler1997]. A
factor analysis was employed at previous waves to confirm that verbal fluency,
immediate verbal recall, delayed verbal recall, digit span backward and digit
span forward represented one dimension. Then, we calculated a global score for
cognitive functioning as the average of the z-scores on each of the variables.
We transformed this score into a 0-100 scale, with higher scores indicating
better cognitive functioning. To realize this step we use the following formula
$(x-min(x))/(max(x)-min(x))*100$ where X represent the non-standardized values.

# Explanation of how the variables were created

A global score in cognition was obtained from the standardized averages of
verbal fluency (`q2536`), immediate verbal recall (`q2525`, `q2528`, and
`q2531`), delayed verbal recall (`q2544`), digit span forward (`q2534`), and
digit span backward (`q2535`). The score is given as a standardized variable,
and transformed to a 0-100 range , with higher scores indicating better
cognitive functioning.

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
recode q2525_recal1_s3 q2528_recal2_s3 q2531_recal3_s3 q2534_score_s3 q2535_score_s3 q2536_score_s3 q2544_recall_s3 (999=.)

egen immediate_c = rowmean(q2525_recal1_s3 q2528_recal2_s3 q2531_recal3_s3)

generate delayed_c = q2544_recall_s3
generate ds_forward = q2534_score_s3
generate ds_backward = q2535_score_s3
generate animals = q2536_score_s3

label variable immediate_c "Number of words recalled correctly in immediate Verbal recall"
label variable delayed_c "Number of words recalled correctly in delayed Verbal recall"

label variable ds_forward "Digit span - Digits forward"
label variable ds_backward "Digit span - Digits backward"
label variable animals "Number of animals named correctly"

egen z_immediate_c = std(immediate_c)
egen z_delayed_c = std(delayed_c)
egen z_ds_forward = std(ds_forward)
egen z_ds_backward = std(ds_backward)
egen z_animals = std(animals)
egen missing_cognitive = rowmiss(z_immediate_c - z_animals)
egen z_cognitive_test = rowmean(z_immediate_c - z_animals) if missing_cognitive<=1

egen cognition_std_s3 = std(z_cognitive_test)
label variable cognition_std_s3 "Cognition score, standardized (c2011_w4)"
 
/* The following step transforms the score generated into a 0-100 scale. */
generate cognition_s3 = (cognition_std_s3 + 2.49287) / (3.002676 + 2.49287)*100
label variable cognition_s3 "Cognition score, harmonized (c2011_w4)"
```

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
sum cognition_s3 cognition_std_s3

keep ID_ECS q0007a_result proxy cognition_std_s3 cognition_s3

codebook cognition_std_s3 cognition_s3
```

```{stata save-data-test, eval=TEST}
saveold "Outcome_cognition.dta", replace version(13)
```

```{stata save-data, eval=!TEST}
* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_cognition.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Lea Francia
-   e-mail: [lea.francia\@inv.uam.es](mailto:lea.francia@inv.uam.es){.email}
-   Phone number: (+34) 914974601

Original version:

-   Name: Félix Caballero

-   e-mail: [felix.caballero\@uam.es](mailto:felix.caballero@uam.es){.email}

-   Phone number: (+34) 91 497 24 47 <br>

-   Name: Elvira Lara

-   e-mail: [elvira.lara\@uam.es](mailto:elvira.lara@uam.es){.email}

-   Phone number: (+34) 914974601

<br>

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
