---
title: "Edad con Salud -- cohort 2011 -- Wave 4 -- 'Chronic Conditions'" 
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
# bibliography: "`r ecs.data::get_proj_bib_file()`" ## TODO: Borrar esta línea, O BIEN deescomentar esta línea y borrar siguiente UNA VEZ completo el documento
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib ## TODO: Descomentar esta línea SI se utiliza el gestor bibliográfico, en otro caso BORRAR esta línea
csl: "`r ecs.data::get_proj_csl_file()`"
# nocite: '@<cita_1>, @<cita_2>' ## TODO: Descomentar y añadir cita(s) o borrar campo
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}

library(Statamarkdown)

STATA_PATH <- r"(Program Files (x86)StataSE-64.exe)"


options(width = 88) 

knitr::opts_chunk$set( 
  echo        = FALSE, 
  results     = 'hide',
  cleanlog    = TRUE, 
  collectcode = TRUE,  
  comment     = ''     
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

+----------------+-------------------------+----------------+----------------+
| Main variable  | Label                   | Format         | Code           |
| name           |                         |                |                |
+:===============+:========================+:===============+:===============+
| arthitis_s3    | arthritis algorithm     | dummy          | 2 no           |
|                | test or self-reported   |                |                |
|                | (c2011_w4)              |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+
| oste           | osteoarthritis          | dummy          | 2 no           |
| oarthritis_s3  | algorithm test or       |                |                |
|                | self-reported           |                | 1 yes          |
|                | (c2011_w4)              |                |                |
+----------------+-------------------------+----------------+----------------+
| stroke_s3      | stroke algorithm test   | dummy          | 2 no           |
|                | or self-reported        |                |                |
|                | (c2011_w4)              |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+
| angina_s3      | angina algorithm test   | dummy          | 2 no           |
|                | or self-reported        |                |                |
|                | (c2011_w4)              |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+
| diab_s3        | diabetes self reported  | dummy          | 2 no           |
|                | (c2011_w4)              |                |                |
|                |                         |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+
| lung_s3        | lung disease algorithm  | dummy          | 2 no           |
|                | test or self-reported   |                |                |
|                | (c2011_w4)              |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+
| asthma_s3      | asthma algorithm test   | dummy          | 2 no           |
|                | or self reported        |                |                |
|                | (c2011_w4)              |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+
| hy             | hypertension algorithm  | dummy          | 2 no           |
| pertension_s3  | test or self reported   |                |                |
|                | (c2011_w4)              |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+
| cataracts_s3   | cataracts algorithm     | dummy          | 2 no           |
|                | test or self reported   |                |                |
|                | (c2011_w4)              |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+
| ora            | oral problems algorithm | dummy          | 2 no           |
| l_problems_s3  | test or self reported   |                |                |
|                | (c2011_w4)              |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+
| kidney_s3      | kidney self reported    | dummy          | 2 no           |
|                | (c2011_w4)              |                |                |
|                |                         |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+
| diges_s3       | digestive self reported | dummy          | 2 no           |
|                | (c2011_w4)              |                |                |
|                |                         |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+
| cancer_s3      | cancer self reported    | dummy          | 2 no           |
|                | (c2011_w4)              |                |                |
|                |                         |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+
| liver_s3       | liver self reported     | dummy          | 2 no           |
|                | (c2011_w4)              |                |                |
+----------------+-------------------------+----------------+----------------+
| osteop_s3      | osteoporosis self       | dummy          | 2 no           |
|                | reported (c2011_w4)     |                |                |
|                |                         |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+
| anaemia_s3     | anaemia self reported   | dummy          | 2 no           |
|                | (c2011_w4)              |                |                |
|                |                         |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+
| thyroid_s3     | thyroid self reported   | dummy          | 2 no           |
|                | (c2011_w4)              |                |                |
|                |                         |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+
| glaucoma_s3    | glaucoma self reported  | dummy          | 2 no           |
|                | (c2011_w4)              |                |                |
|                |                         |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+
| heart_s3       | heart self reported     | dummy          | 2 no           |
|                | (c2011_w4)              |                |                |
|                |                         |                | 1 yes          |
+----------------+-------------------------+----------------+----------------+

# Variables description

General comment: Chronic condition algorithms were developed for the World
Health Survey and adapted from COURAGE. Symptom based estimates were prepared
for the following conditions: arthritis, stroke, angina, lung disease, asthma
and hypertension. In such cases, the algorithm reflects a specific response
pattern or a combination of answers indicating whether someone has a disease
(see explanations below) in the previous 12 months. All other response patterns
(i.e. one or more no's) will result in 'no disease'. Additionally, self-reported
diagnoses (ever diagnosed) were included in all chronic conditions.

Important note: In "Edad con Salud" we have created an option (i.e. var_a =997)
for those individuals who reported a disease in first assessment, but in second
assessment they said that they have never had such disease. There are ≈10-20
individuals per disease that have reported this situation. We have considered
this circumstance in the self-reported algorithms presented below (i.e. those
who said 997 have been considered as not having the disease). This condition
applies for all the diagnoses.

## Arthritis
Rheumatoid Arthritis is a chronic systemic disease that affects the joints, 
connective tissues, muscle, tendons, and fibrous tissue. It tends to strike 
during the most productive years of adulthood, between the ages of 20 and 40, 
and is a chronic disabling condition often causing pain and deformity. 
Osteoarthritis is a degenerative joint disease, which mainly affects the 
articular cartilage. It is associated with ageing and will most likely affect 
the joints that have been continually stressed throughout the years including 
the knees, hips, fingers, and lower spine region. 

## Stroke
A stroke is caused by the interruption of the blood supply to the brain, 
usually because a blood vessel bursts or is blocked by a clot. This cuts off 
the supply of oxygen and nutrients, causing damage to the brain tissue.

## Angina
Angina is a chest pain that occurs when the blood supply to the muscles of 
the heart is restricted. It usually happens because the arteries supplying the 
heart become hardened and narrowed.

## Diabetes
Diabetes is a chronic, metabolic disease characterized by elevated levels of 
blood glucose (or blood sugar), which leads over time to serious damage to the 
heart, blood vessels, eyes, kidneys, and nerves. The most common is type 2 
diabetes, usually in adults, which occurs when the body becomes resistant to 
insulin or doesn't make enough insulin.

## Lung Disease
Lung disease is any problem in the lungs that prevents the lungs from 
working properly. There are three main types of lung disease: a) airway 
diseases, that affect the tubes (airways) that carry oxygen and other gases 
into and out of the lungs; b) Lung tissue diseases, affecting the structure of 
the lung tissue; and c) Lung circulation diseases, that affect the blood vessels 
in the lungs. 

## Asthma 
Asthma is a chronic disease characterized by recurrent attacks of breathlessness 
and wheezing, which vary in severity and frequency from person to person. During 
an asthma attack, the lining of the bronchial tubes swells, causing the airways 
to narrow and reducing the flow of air into and out of the lungs.

## Hypertension
Hypertension, also known as high or raised blood pressure, is a condition in 
which the blood vessels have persistently raised pressure. Blood is carried from 
the heart to all parts of the body in the vessels. Each time the heart beats, it 
pumps blood into the vessels. Blood pressure is created by the force of blood 
pushing against the walls of blood vessels (arteries) as it is pumped by the 
heart. The higher the pressure the harder the heart has to pump.

## Cataracts
A cataract is clouding of the lens of the eye, which impedes the passage of 
light. Most cataracts are related to ageing, although occasionally children may 
be born with the condition, or cataract may develop after an injury, 
inflammation or disease. 

## Oral Health
Oral health is essential to general health and quality of life. It is a state of 
being free from mouth and facial pain, oral and throat cancer, oral infection 
and sores, periodontal (gum) disease, tooth decay, tooth loss, and other 
diseases and disorders that limit an individual’s capacity in biting, chewing, 
smiling, speaking, and psychosocial wellbeing. Risk factors for oral diseases 
include an unhealthy diet, tobacco use and harmful alcohol use.

## Kidney Disease
Chronic kidney disease includes complications like high blood pressure, 
anaemia, weak bones, poor nutritional health and nerve damage. It may be caused 
by diabetes, high blood pressure and other disorders.

## Digestive Disease
A digestive disease is any health problem that occurs in the digestive tract, 
and may range from mild to serious conditions. Some common problems include 
cancer, irritable bowel syndrome, and lactose intolerance.

## Cancer
Cancer is a generic term for a large group of diseases that can affect any part 
of the body. One defining feature of cancer is the rapid creation of abnormal 
cells that grow beyond their usual boundaries, and which can then invade 
adjoining parts of the body and spread to other organs, the latter process is 
referred to as metastasizing. 

## Liver Disease
There are diverse kinds of liver diseases that difficult the process of 
digestion, store energy, or poisons removal. Viruses cause some of them 
(i.e. hepatitis A, hepatitis B and hepatitis C), while others can be the result 
of drugs, poisons or drinking too much alcohol.

## Osteoporosis
Osteoporosis is a disease in which the density and quality of bone are reduced. 
As bones become more porous and fragile, the risk of fracture is greatly 
increased. The loss of bone occurs silently and progressively. 

## Anaemia
Anaemia is a condition in which the number of red blood cells or their 
oxygen-carrying capacity is insufficient to meet physiologic needs, which vary 
by age, sex, altitude, smoking, and pregnancy status.

## Thyroid
The thyroid gland is an endocrine gland in the neck. It makes two hormones that 
are secreted into the blood: thyroxine (T4) and triodothyronine (T3). These 
hormones are necessary for all the cells in the body to work normally. Thyroid 
disorders are very common and tend mainly to occur in women. If too much of the 
thyroid hormones are secreted (hyperthyroid), the body cells work faster than 
normal, and this increased activity may lead to a quickening of your heart rate 
or increased activity of the intestine. If too little of the thyroid hormones 
are produced (hypothyroidism), the cells and organs of your body slow down. 
In that case, the heart rate, for example, may be slower than normal and the 
intestines work sluggishly.

## Glaucoma
Glaucoma is the general term for a group of similar diseases. In primary open 
angle glaucoma, the channels that drain fluid within the eye become blocked, 
causing the pressure within the eye to rise. It causes gradual loss of vision. 
There are few symptoms so that people may not notice for a long time that they 
are losing their sight.

## Heart Disease
Cardiovascular diseases are disorders of the heart and blood vessels and include 
coronary heart disease, cerebrovascular disease, rheumatic heart disease and 
other conditions. Four out of five CVD deaths are due to heart attacks and 
strokes. Individuals at risk of CVD may demonstrate raised blood pressure, 
glucose, and lipids as well as overweight and obesity.

# Explanation of how the variables were created

## Arthritis
A diagnosis of arthritis was made if the participant fulfilled any of the 
available criteria: (1) self-reported diagnosis of arthritis (q4001_arthritis_s3); (2) self-reported symptoms of arthritis (q4003_symptoms_s3, q4004_symptoms_s3,); (3) self-reported as under treatment for arthritis (q4002b_treated_s3) and characteristics like duration (q4005_duration_s3 and q4006_outcome_s3)
As for the self-reported diagnosis, arthritis was endorsed if the participant 
answered "yes" (1) to self-reported diagnosis of arthritis, but if they affirmed they had never had the disease (not "997" to q4001a_s3).

## Stroke
A diagnosis of stroke was made if the 
participant (1) self-reported the diagnosis of a stroke (q4010_stroke_s3) or (2) suffer from two symptoms: a sudden paralysis or weakness in arms or legs on one side of the body and a sudden loss of sensation on one side of the body (q4012_parals_s3, q4013_feeling_s3).

## Angina
A diagnosis of angina was made if the participant self reported suffering from angina (q4014_angina_s3) or if he presents several symptoms such as chest pain when walking downhill, chest pain when walking normally, etc. (q4016_symptoms_s3, q4017_symptoms_s3, q4018_r, q4019_s, q4020_where_6_s3, q4020_where_7_s3, q4020_where_8_s3, q4020_where_11_s3) 
Diagnosis of angina was endorsed if the participant answered "yes" (1) to self-reported diagnosis , but not if they affirmed they had never had the disease (not "997" to q4014_angina_s3).


## Diabetes
Original variables include the self-reported diagnosis of diabetes (q4022_diabetes_s3), the self-reported treatment of diabetes (q4023b_treated_s3) and the following of a special diet (q4024_diet_s3). Diabetes was endorsed when the participant answered "yes" to 
self-reported diagnose of diabetes, but not 997 in q4022a_s3 (they affirmed they had never had the disease). 

## Lung Disease
A diagnosis of lung disease was made if the participant fulfilled any of the available criteria 
(i.e. self-reported diagnosis or symptoms). Those participants who answered "yes" to q4027_symptoms_s3, q4028_symptoms_s3 and q4029_symptoms_s3, were defined as having had lung disease. As for the self-reported diagnosis, lung disease was endorsed if the participant answered "yes" to q4025_lungs_s3, but not 997 in q4025a_s3 (they affirmed they had never had the disease). Please, consider that this algorithm does not include asthma as a part of lung diseases.

## Asthma 
A diagnosis of asthma was made if the participant has self-reported having the diases (q4033_asthma_s3) or if he is treated for asthma (q4034b_treated_s3)  or several symptoms such as (1) wheezing attacks (q4035_symptoms_s3), (2) wheezing attack immediately after physical activity (q4036_symptoms_s3), (3) feeling of tightness in the chest (q4037_symptoms_s3), (4) woken up with a tight feeling in chest (q4038_symptoms_s3), (5) attack of shortness of breath, which occurs without an obvious cause (q4039_symptoms_s3).
Those participants who answered "yes" to all five symptoms were defined as having had asthma. As for the self-reported diagnosis, 
asthma was endorsed if the participant answered "yes" to q4033_asthma_s3, but not 997 in q4033a_s3 (they affirmed they had never had the disease). 
Note: it is possible to create the symptom algorithm in a less conservative way. 
That is: if q4035_symptoms_s3=1  and q4036_symptoms_s3=1  or q4037_symptoms_s3=1  
or q4038_symptoms_s3=1  or q4039_symptoms_s3=1, asthma_test=1. 
There is, apparently, no wrong or right option.

## Hypertension
Original variables include q2501_syst_s3, q2502_syst_s3, q2501_diast_s3, 
q2502_diast_s3, q4061b_treated_s3 and q4060_high_bp_s3. A diagnosis of asthma 
was made if the participant fulfilled any of the available criteria 
(i.e. self-reported diagnosis or symptoms).  
Those participants with systolic blood pressure ≥140 mmHg and/or diastolic blood 
pressure ≥90 mmHg were considered to have hypertension 
(blood pressure measurement). As for the self-reported diagnosis, 
asthma was endorsed if the participant answered "yes" to q4060_high_bp_s3, 
but not 997 in q4060a_s3 (they affirmed they had never had the disease). 

## Cataracts
Original variables include q4062_catar_s3, q4063_surgery_s3, q4064_vision_s3 
and q4065_light_s3. A diagnosis of cataracts was made if the participant 
fulfilled any of the available criteria 
(i.e. self-reported diagnosis or symptoms). 
Those participants who answered “yes” to q4064_vision_s3 and q4065_light_s3 
were defined as having had cataracts. We also included a treatment code 
(cataracts_self=2) in those individuals who, apart from having reported a 
cataract diagnosis, have had surgery for cataracts in the previous 12 months.

## Oral Health
Original variables include q4066_teeth_s3, q4067_symptoms_s3 
and q4068b_treated_s3. Oral problems were endorsed if the participant fulfilled 
any of the available criteria (i.e. total tooth loss or self-reported problems). 
Those participants who answered “yes” to q4066_teeth_s3, 
but not 997 in q4066a_s3 (they affirmed they had never lost their tooth) were 
defined as having had total tooth loss. As for the self-reported symptoms, 
oral problems were endorsed if the participant answered "yes" 
to q4067_symptoms_s3. 

## Kidney Disease
Original variables include q4085_kidney_s3 and q4086_s3. 
A diagnosis of kidney disease was endorsed if the participant 
answered "yes" to q4085_kidney_s3. 

## Digestive Disease
Original variables include q4087_s3 and q4088_s3. 
A diagnosis of digestive disease was endorsed if the participant 
answered "yes" to q4087_s3. 

## Cancer
Original variables include q4089_s3. A diagnosis of cancer disease was endorsed 
if the participant answered "yes" to q4089_s3. 

## Liver Disease
Original variables include q4091_s3 and q4092_s3. A diagnosis of liver disease 
was endorsed if the participant answered "yes" to q4091_s3. 

## Osteoporosis
Original variables include q4093_s3 and q4094_s3. A diagnosis of osteoporosis 
was endorsed if the participant answered "yes" to q4093_s3. 

## Anaemia
Original variables include q4095_s3 and q4096_s3. A diagnosis of anaemia was 
endorsed if the participant answered "yes" to q4095_s3. 

## Thyroid
Original variables include q4097_s3 and q4098_s3. A diagnosis of thyroid disease 
was endorsed if the participant answered "yes" to q4097_s3. 

## Glaucoma
Original variables include q4099_s3 and q4100_s3. A diagnosis of glaucoma was 
endorsed if the participant answered "yes" to q4099_s3. 

## Heart Disease
Original variables include q4101_s2 and q4102_s2. A diagnosis of heart disease was endorsed if the participant answered "yes" to q4101_s2. 

# Syntax used to create the variable (Stata 15)

```{stata load-data}

* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
* TODO: Escribir aquí la sintaxis de generación de las variables
```



```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}

keep ID_ECS q0007a_result proxy <variable(s) generada(s)>

codebook <variable(s) generada(s)>
```



```{stata save-data-test, eval=TEST}
saveold "Outcome_chronicconditions.dta", replace version(13)
```

```{stata save-data, eval=!TEST}
* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_chronicconditions.dta", replace version(13)

```

<br>

# Contact person:

-   Name: Lea Francia
-   e-mail: [lea.francia\@inv.uam.es](mailto:lea.francia@inv.uam.es){.email}
-   Phone number: (+34) 914974601

<br>

\< Original version:

-   Name: Emese Verdes-Tennant
-   e-mail: [verdese\@who.int](mailto:nombre.buzon@institucion.org){.email}
-   Phone number: +41 22 791 1428

<br>

-   Name: Beatriz Olaya
-   e-mail: [beatriz.olaya\@pssjd.org](mailto:beatriz.olaya@pssjd.org){.email}
-   Phone number: (+34)936406350 ext(1)2540


# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
