---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'CAREGIVER'S BURDEN'" 
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
# bibliography: "`r ecs.data::get_proj_bib_file()`" ## TODO: Borrar esta línea, O BIEN deescomentar esta línea y borrar siguiente UNA VEZ completo el documento
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib ## TODO: Descomentar esta línea SI se utiliza el gestor bibliográfico, en otro caso BORRAR esta línea
csl: "`r ecs.data::get_proj_csl_file()`"
nocite: '@bedard_zarit_2001, @zarit_relatives_1980' ## TODO: Descomentar y añadir cita(s) o borrar campo
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

# Ruta al ejecutable de Stata (Ejecuta `sysdir` en Stata para encontrar la ruta)

## TODO: sustituir '<path\to\>' por la ubicación de Stata
##       (ejecutar `sysdir` en Stata) y descomentar línea 62,
##       o eliminar linea 62:
STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"


# Opciones de configuración de las salidas de R y Stata:

options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( # Para todos los 'chunks' de código:
  ## TODO: Descomentar si hace falta
  # engine.path = list(stata = STATA_PATH),
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information
+----------------------+------------------------------------------+------------+----------------------+
| Main variable name   | Label                                    | Format     | Code                 |
+:=====================+:=========================================+:===========+:=====================+
| caregivers_burden_s3 | Caregiver's burden, continuous (c2011_w4)| Continuous |  0 - 43              |
+----------------------+------------------------------------------+------------+----------------------+

# Variable description

The 22-item version of the Zarit Burden Interview (ZBI) is the most widely used 
caregiver self-report measure used by many aging agencies (@zarit_relatives_1980).
It has been used in a variety of research designs, to discriminate among study 
participants and to measure change over time resulting from progression of the 
care recipient's condition, or from interventions aimed at reducing burden. 
A significant advantage of the popularity of the ZBI is that results obtained in
studies can be easily compared and synthesized. We used a reduced version of 12 
items (@bedard_zarit_2001).

Each item on the interview is a statement which the caregiver is 
asked to endorse using a 5-point scale. Response options range from 0 (Never) 
to 4 (Nearly Always).

# Explanation of how the variables were created
The variable is created as the product of the sum of the reduced version' items:
`q6712_s3` (insufficient time), `q6713_s3` (stress), `q6714_s3` (irritated),
`q6715_s3` (negative impact on social networks), `q6716_s3` (exhausted), 
`q6717_s3` (health), `q6718_s3` (limited private life), `q6719_s3` (worse social
networks), `q6720_s3` (life out of control), `q6721_s3` (insecure), `q6722_s3` 
(could do more) and `q6723_s3` (could do better).

It is a scale from 0 to 48 (empirically 43). Higher indexes indicate higher burden. 

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
recode q6712_s3 q6713_s3 q6714_s3 q6715_s3 q6716_s3 q6717_s3 q6718_s3 q6719_s3 q6720_s3 q6721_s3 q6722_s3 q6723_s3 (1=0) (2=1) (3=2) (4=3) (5=4) (else=.)


gen caregivers_burden_s3= q6712_s3 + q6713_s3 + q6714_s3 + q6715_s3 + q6716_s3 + q6717_s3 + q6718_s3 + q6719_s3 + q6720_s3 + q6721_s3 + q6722_s3 + q6723_s3


label variable caregivers_burden_s3 "Caregiver's burden, continuous (c2011_w4)"
```

<br> 

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result_s3 proxy_s3 caregivers_burden_s3

codebook caregivers_burden_s3
```

```{stata save-data-test, eval=TEST}
saveold "Outcome_caregiversburden.dta", replace version(13)
```


```{stata save-data, eval=!TEST}
* Cohorte 2011, ola 4:
* saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_caregiversburden.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Cristina Rodríguez Prada
-   e-mail: [cristina.rodriguez@inv.uam.es](mailto:cristina.rodriguez@inv.uam.es){.email}


<br>


Original version:

-   Name: Joan Domènech
-   e-mail:
    [j.domenech@pssjd.org](mailto:j.domenech@pssjd.org){.email}

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
