---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'Mini Mental State Examination'" 
output:
  word_document:
   reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib 
csl: "`r ecs.data::get_proj_csl_file()`"
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"

options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( 
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

| Main variable name | Label                                             | Format    | Code |
|:------------------|:----------------------|:------------------|:------------------|
| sum_MMSEc_s3       | Rounded and corrected MMSE total score (c2011_w4) | Continous | 0-30 |

# Variable description

The Mini Mental State Examination (MMSE) [@folstein1983] is effective as a
screening tool for cognitive impairment with older, community dwelling,
hospitalized and institutionalized adults, being able to separate patients with
cognitive impairment from those without it, and to measure changes in cognitive
status that may benefit from intervention. It is an 11-question measure that
tests five areas of cognitive function: orientation, registration, attention and
calculation, memory, and language. The maximum score is 30. A score of 23 or
lower is indicative of cognitive impairment.

# Explanation of how the variables were created

The total score of MMSE was calculated as the sum of the following items:
q1038a-e_s3 (temporal orientation), q1039a-e_s3 (spacial orientation), q1040_s3
(fixing), q1041_s3 (attention and calculation), q1042_s3 (memory), q1043_s3
(denomination), q1044_s3 (repetition), q1045_s3 (comprehension), q1046_s3
(reading), q1047_s3 (writing), and q1048_s3 (drawing). The total score ranges
from 0 to 30 (higher scores, better cognition).

There are some items which can contain "Not Applicable" (NA) [999], if
respondent has no education or whether has some physical impairment preventing
them from answering (e.g., blindness). For these people, we first created a
variable indicating whether the respondent had some NA (i.e. q1041_s3, q1043_s3,
q1045_s3, q1046_s3, q1047_s3, q1048_s3. Then, we recoded NA answers (999) into
missing values. We summed up all the items for people with no NA and for people
with special conditions. Then we calculated the corrected score using with the
following formula: (sum_MMSE_v2\*30)/ (30+d1+d2+d3+d4+d5+d6), being d1=-5 if
q1041=NA d2=-2 if q1043_s3=NA d3=-3 if q1045_s3=NA d4=-1 if q1046_s3=NA d5=-1 if
q1047=NA d6=-1 if q1048_s3=NA Finally, we rounded these corrected scores to the
nearest unit.

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"

```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
* Filter: q1038 (incapacidad de hacer mini mental) 

***Cases excluding NA ***

* Generating a variable for people with NA in any of the MMSE items 
gen na_mmse=0
replace na_mmse=1 if (q1041_s3==999 | q1043_s3==999 | q1045_s3==999 | q1046_s3==999 | q1047_s3==999 | q1048_s3==999) | (q1041_s3==888 | q1043_s3==888 | q1045_s3==888 | q1046_s3==888 | q1047_s3==888 | q1048_s3==888)
replace na_mmse=. if (q1038_s3==1  | q0007a_result!=1) 
label variable na_mmse "Individuals with NA answers in MMSE"

* Converting 999 and 888 in missing values
recode q1041_s3 q1043_s3 q1045_s3 q1046_s3 q1047_s3 q1048_s3 (999=.) (888=.)

* Sum of MMSE items for individuals without NA answers (all items were applied) */
egen sum_MMSE_v1=rowtotal(q1038a_s3 q1038b_s3 q1038c_s3 q1038d_s3 q1038e_s3 q1039a_s3 q1039b_s3 q1039c_s3 q1039d_s3 q1039e_s3 q1040_s3 q1041_s3 q1042_s3 q1043_s3 q1044_s3 q1045_s3 q1046_s3 q1047_s3 q1048_s3) if na_mmse==0, missing
label variable sum_MMSE_v1 "MMSE total score for individuals with no NA answers"

* Sum of MMSE items for individuals with NA answers
egen sum_MMSE_v2=rowtotal(q1038a_s3 q1038b_s3 q1038c_s3 q1038d_s3 q1038e_s3 q1039a_s3 q1039b_s3 q1039c_s3 q1039d_s3 q1039e_s3 q1040_s3 q1041_s3 q1042_s3 q1043_s3 q1044_s3 q1045_s3 q1046_s3 q1047_s3 q1048_s3) if na_mmse==1, missing
label variable sum_MMSE_v2 "MMSE total score for individuals with NA answers"

* Variables to correct total scores in case of NA answers
gen d1=0
replace d1=-5 if (q1041_s3==999)
gen d2=0
replace d2=-2 if (q1043_s3==999)
gen d3=0
replace d3=-3 if (q1045_s3==999)
gen d4=0
replace d4=-1 if (q1046_s3==999)
gen d5=0
replace d5=-1 if (q1047_s3==999)
gen d6=0
replace d6=-1 if (q1048_s3==999)


* Corrected MMSE total score
gen sum_MMSE_v3=.
replace sum_MMSE_v3= sum_MMSE_v1
replace sum_MMSE_v3= (sum_MMSE_v2*30)/ (30+d1+d2+d3+d4+d5+d6) if na_mmse==1
label variable sum_MMSE_v3 "Crude and corrected MMSE total score "   

gen sum_MMSEc_s3=round(sum_MMSE_v3)
label variable sum_MMSEc_s3 "Rounded and corrected MMSE total score (c2011_w4)"   

```

<br>

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}

keep ID_ECS q0007a_result proxy sum_MMSEc_s3

codebook sum_MMSEc_s3

```

```{stata save-data-test, eval=TEST}
saveold "Outcome_MMSE.dta", replace version(13)
```

```{stata save-data, eval=!TEST}

* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_MMSE.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Lea Francia
-   e-mail: [lea.francia\@inv.uam.es](mailto:lea.francia@inv.uam.es){.email}
-   Phone number: (+34) 914974601

Original version:

-   Name: Dario Moreno Agostino
-   e-mail: [d.moreno\@ucl.ac.uk](mailto:d.moreno@ucl.ac.uk){.email}

<br>

-   Name: Elvira Lara
-   e-mail: [elvira.lara\@uam.es](mailto:elvira.lara@uam.es){.email}
-   Phone number: (+34) 914974601

<br>

-   Name: Daniel Morillo
-   e-mail:
    [daniel.morillo\@inv.uam.es](mailto:daniel.morillo@inv.uam.es){.email}
-   Phone number: (+34) 914975816

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
