---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'EASI'"
output:
  word_document:
   reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib 
csl: "`r ecs.data::get_proj_csl_file()`"
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"


options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( 
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

+-------------------+----------------+--------------------+---------------------+
| Main variable     | Label          | Format             | Code                |
| name              |                |                    |                     |
+:==================+:===============+:===================+:====================+
| dependency_s3     | Dependency     | Ordinal            | -   0 = No          |
|                   | status         |                    |     dependence      |
|                   | comparable     |                    |                     |
|                   | (c2011_w4)     |                    | -   1 = Dependency  |
|                   |                |                    |     served          |
|                   |                |                    |                     |
|                   |                |                    | -   2 = Problematic |
|                   |                |                    |     dependence      |
|                   |                |                    |                     |
|                   |                |                    | -   3 = Neglected   |
|                   |                |                    |     dependence      |
+-------------------+----------------+--------------------+---------------------+
| abuse_negli_s3    | Negligence     | Ordinal            | -   0 = No abuse    |
|                   | comparable     |                    |                     |
|                   | (c2011_w4)     |                    | -   1 = Occasional  |
|                   |                |                    |                     |
|                   |                |                    | -   2 = Sometimes   |
|                   |                |                    |                     |
|                   |                |                    | -   3 = Quite often |
|                   |                |                    |                     |
|                   |                |                    | -   4 = Often       |
+-------------------+----------------+--------------------+---------------------+
| abuse_psycho_s3   | Psychological  | Ordinal            | -   0 = No abuse    |
|                   | abuse          |                    |                     |
|                   | comparable     |                    | -   1 = Occasional  |
|                   | (c2011_w4)     |                    |                     |
|                   |                |                    | -   2 = Sometimes   |
|                   |                |                    |                     |
|                   |                |                    | -   3 = Quite often |
|                   |                |                    |                     |
|                   |                |                    | -   4 = Often       |
+-------------------+----------------+--------------------+---------------------+
| abuse_eco_s3      | Economical     | Ordinal            | -   0 = No abuse    |
|                   | abuse          |                    |                     |
|                   | comparable     |                    | -   1 = Occasional  |
|                   | (c2011_w4)     |                    |                     |
|                   |                |                    | -   2 = Sometimes   |
|                   |                |                    |                     |
|                   |                |                    | -   3 = Quite often |
|                   |                |                    |                     |
|                   |                |                    | -   4 = Often       |
+-------------------+----------------+--------------------+---------------------+
| abuse_phys_s3     | Physical and   | Ordinal            | -   0 = No abuse    |
|                   | sexual         |                    |                     |
|                   | comparable     |                    | -   1 = Occasional  |
|                   | (c2011_w4)     |                    |                     |
|                   |                |                    | -   2 = Sometimes   |
|                   |                |                    |                     |
|                   |                |                    | -   3 = Quite often |
|                   |                |                    |                     |
|                   |                |                    | -   4 = Often       |
+-------------------+----------------+--------------------+---------------------+
| abusesuspicion_s3 | Suspicion of   | Dichotomic         | -   0 = No          |
|                   | abuse          |                    |                     |
|                   | (c2011_w4)     |                    | -   1 = Yes         |
+-------------------+----------------+--------------------+---------------------+

# Variable description

The Spanish version of the Elder Abuse Suspicion Index [@pérez-rojo2008] is an
adaptation of the Elder Abuse Suspicion Index (EASI), a brief instrument with
direct questions aimed at detecting suspicion of abuse in older people (aged
+64) without cognitive impairment. It is made up of five questions addressed to
the elderly person and one to the professional, which are asked in the course of
an encounter between the elderly person and the professional in the consultation
room and are formulated by the professional using simple language. The first
question focuses on the assessment of the level of dependency of the elderly
person for some basic or instrumental activity of daily living. The remaining
questions refer to different types of abuse: physical and sexual, psychological,
financial, and neglect. The items are dichotomous (0=No; 1=Yes). A positive
response to any of the questions would indicate suspicion of maltreatment.

# Explanation of how the variables were create

To created the ´abusesuspicion_s3´ variable, we decide to keep only the 5 items
addressed to the participants and excluded the question asked to the
professional(s). These 5 items ask information about different types of
maltreatment: neglect (`q6701_s3, q6702_s3`), psychological (`q6703_s3`),
economic (`q6704_s3`) and physical-sexual (`q6705_s`). The total scale of
suspicion of abuse was obtain by summing the answer given to each 5 items. A
transformation of the scores was then made, for those participants who answer
yes to on of the five item a suspicion of abuse is consider. The scale ranges
from 0: no suspicion of abuse to 1: suspicion of abuse. The questions refer to
situations that occurred in the 12 months prior to the completion of the
questionnaire and only persons aged 65 and over are asked.

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "/Users/leafrancia/Library/CloudStorage/OneDrive-Bibliothèquespartagées-UAM/Marta Miret Garcia - Bases de datos maestras Edad con Salud/Ola_4/Cohorte 2011/rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
*Suspicion of abuse
recode q6701_s3 (1=1) (2=0) (else=.)
recode q6702_s3 (1=1) (2=0) (else=.)
recode q6703_s3 (1=1) (2=0) (else=.)
recode q6704_s3 (1=1) (2=0) (else=.)
recode q6705_s3 (1=1) (2=0) (else=.)
gen abuse_tot = (q6701_s3 + q6702_s3 + q6703_s3 + q6704_s3 + q6705_s3)
gen abusesuspicion_s3 =.
replace abusesuspicion_s3=0 if abuse_tot<1 
replace abusesuspicion_s3=1 if abuse_tot>0 & abuse_tot!=.
label define abusesuspicion_l 0 "No suspicion of abuse" 1 "Suspicion of abuse"
label values abusesuspicion_s3 abusesuspicion_l
label variable abusesuspicion_s3 "Suspicion of abuse (c2011_w4)"
```

<br>

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result proxy abusesuspicion_s3

codebook abusesuspicion_s3
```

```{stata save-data-test, eval=TEST}

saveold "Outcome_suspicionofabuse.dta", replace version(13)
```

```{stata save-data, eval=!TEST}

* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_suspicionofabuse.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Lea Francia
-   e-mail: [lea.francia\@inv.uam.es](mailto:lea.francia@inv.uam.es){.email}
-   Phone number: (+34) 914974601

Original version:

-   Name: Dario Moreno Agostino
-   e-mail: [d.moreno\@ucl.ac.uk](mailto:d.moreno@ucl.ac.uk){.email}

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
