---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'Material deprivation'" ## TODO: Añadir "cohort", "assessment" y "variable name(s)/domain"
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
#bibliography: "`r ecs.data::get_proj_bib_file()`"
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib ## TODO: Descomentar esta línea SI se utiliza el gestor bibliográfico, en otro caso BORRAR esta línea
csl: "`r ecs.data::get_proj_csl_file()`"
nocite: '@instituto_nacional_de_estadistica_ine_encuesta_2005' ## TODO: Descomentar y añadir cita(s) o borrar campo
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---


```{r setup, include=FALSE}
library(Statamarkdown)
STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"
options(width = 88) 
knitr::opts_chunk$set( 
  echo        = FALSE, 
  results     = 'hide',
  cleanlog    = TRUE,  
  collectcode = TRUE,  
  comment     = ''     
)
TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information
+--------------------+------------------------------------------+------------+----------------------+
| Main variable name | Label                                    | Format     | Code                 |
+:===================+:=========================================+:===========+:=====================+
| material_s3        | Household with material deprivation      | Categorical| 0 = No               |
|                    | (c2011_w4)                               |            | 1 = Yes              |
+--------------------+------------------------------------------+------------+----------------------+

# Variable description

The framework of community action in the fight against social exclusion has 
allowed the development of indicators that make possible to measure inequality 
in income distribution, poverty rates before and after considering social 
transfers, the persistence of poverty index and the rate of deprivation of
material goods (tangible items that consumers can purchase, sell or trade; 
usually associated with a minimum quality on living conditions). 
**Material deprivation** is an indicator used in the Spanish Life 
Conditions Survey (ECV, Encuesta de Condiciones de Vida) designed by the Spanish
National Institute of Statistics (INE, Instituto Nacional de Estadística). It is
based on 7 items that include non-monetary indicators of household deprivation, 
including problems making ends meet, level of indebtedness and lack of access to
basic necessities. In the questionnaire, material deprivation is defined when 
the household can’t afford at least 3 items out of 7.

# Explanation of how the variables were created
Households with material deprivation was assessed through 7 items, 
considering the threshold for material deprivation when 3 of the 7 were met.
Original variables include `q0411c_s3`(being able to afford to go on 
vacation at least one week a year), `q0411d_s3` (afford a meal of meat, 
poultry or fish at least every two days), `q0411e_s3` (being able to afford
to maintain the housing at an adequate temperature), `q0411f_s3` (having the
capacity to face unforeseen expenses (from 650 euros)), `q0411g_s3` (having had
delays in the payment of expenses related to the main residence (mortgage or 
rent, gas bills, community fees...) or installment purchases in the last 12
months), `q0411h_s3`(being able to afford a car) and `q0411i_s3` (being
able to afford a personal computer). (1 = yes; 2 = no)

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
label values q0411e_s3 q0411f_s3 q0411g_s3 q0411h_s3 q0411i_s3

recode q0411c_s3 (888=.) (999=.) (1=0) (2=1)
recode q0411d_s3 (888=.) (999=.) (1=0) (2=1)
recode q0411e_s3 (888=.) (999=.) (1=0) (2=1)
recode q0411f_s3 (888=.) (999=.) (1=0) (2=1)
recode q0411g_s3 (888=.) (999=.) (2=0)
recode q0411h_s3 (888=.) (999=.) (1=0) (2=1)
recode q0411i_s3 (888=.) (999=.) (1=0) (2=1)

gen mat_sum_s3 = q0411c_s3 + q0411d_s3 + q0411e_s3 + q0411f_s3 + q0411g_s3 + q0411h_s3 + q0411i_s3

gen material_s3=.
replace material_s3=0 if mat_sum_s3<=2
replace material_s3=1 if (mat_sum_s3>2 & mat_sum_s3<8)

label define labelsm_s3 0 "No" 1 "Yes"
label values material_s3 labelsm_s3
label variable material_s3 "Household with material deprivation (c2011_w4)"

```

<br> 

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result_s3 proxy_s3 material_s3
codebook material_s3
```

```{stata save-data-test, eval=TEST}
saveold "Outcome_materialdeprivation.dta", replace version(13)
```


```{stata save-data, eval=!TEST}
* Cohorte 2011, ola 4:
* saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_materialdeprivation.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Cristina Rodríguez Prada _(adapted algorithm for wave 4)_
-   e-mail:
    [cristina.rodriguez@inv.uam.es](cristina.rodriguez@inv.uam.es){.email}

<br>

<!--# TODO: Completar / eliminar "original version" según corresponda -->

Original version:

-   Name: Aina Gabarrell
-   e-mail:
    [a.gabarrell@pssjd.org ](mailto:a.gabarrell@pssjd.org ){.email}

# References
::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
