---
title: "Edad con Salud -- cohort 2011 -- w4 -- 'Sociodemographic variables'" 
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
# bibliography: "`r ecs.data::get_proj_bib_file()`" ## TODO: Borrar esta línea, O BIEN deescomentar esta línea y borrar siguiente UNA VEZ completo el documento
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib ## TODO: Descomentar esta línea SI se utiliza el gestor bibliográfico, en otro caso BORRAR esta línea
csl: "`r ecs.data::get_proj_csl_file()`"
# nocite: '@<cita_1>, @<cita_2>' ## TODO: Descomentar y añadir cita(s) o borrar campo
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"


# Opciones de configuración de las salidas de R y Stata:

options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( # Para todos los 'chunks' de código:
  ## TODO: Descomentar si hace falta
  # engine.path = list(stata = STATA_PATH),
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

+--------------------+------------------------------------------+------------+----------------------+
| Main variable name | Label                                    | Format     | Code                 |
+:===================+:=========================================+:===========+:=====================+
| sex_s3             | Sex (c2011w4)                            | Nominal    | 1 = Male;            |
|                    |                                          |            |                      |
|                    |                                          |            | 2 = Female           |
+--------------------+------------------------------------------+------------+----------------------+
| agegroups_s3       | Age, categorical (c2011_w4)              | Ordinal    | 1 = 18-29;           |
|                    |                                          |            |                      |
|                    |                                          |            | 2 = 30-49;           |
|                    |                                          |            |                      |
|                    |                                          |            | 3 = 50-64;           |
|                    |                                          |            |                      |
|                    |                                          |            | 4 = 65-75;           |
|                    |                                          |            |                      |
|                    |                                          |            | 5 = 80+              |
+--------------------+------------------------------------------+------------+----------------------+
| marital_s3         | Marital status (c2011_w4)                | Nominal    | 1 = Never married    |
|                    |                                          |            | and no cohabiting;   |
|                    |                                          |            |                      |
|                    |                                          |            | 2 = Currently married|
|                    |                                          |            | /law-partnered;      |
|                    |                                          |            |                      |
|                    |                                          |            | 3 = Cohabiting;      |
|                    |                                          |            |                      |
|                    |                                          |            |4 = Separated/divorced|   
|                    |                                          |            |                      |
|                    |                                          |            | 5 = Widowed          |
+--------------------+------------------------------------------+------------+----------------------+
| education_s3       |Education level, categorical (c2011_w4)   | Ordinal    | 1 = Less than primary|
|                    |                                          |            |;                     |
|                    |                                          |            |                      |
|                    |                                          |            | 2 = Primary;         |
|                    |                                          |            |                      |
|                    |                                          |            | 3 = Lower secondary; |
|                    |                                          |            |                      |
|                    |                                          |            | 4 = Upper secondary; |
|                    |                                          |            |                      |
|                    |                                          |            | 5 = Tertiary         |
+--------------------+------------------------------------------+------------+----------------------+
| employment_s3      | Employment status (c2011_w4)             | Nominal    | 1 = Never worked/    |
|                    |                                          |            | homemaker;           |
|                    |                                          |            |                      |
|                    |                                          |            | 2 = Working;         |
|                    |                                          |            |                      |
|                    |                                          |            | 3 = Retired/disabled;|
|                    |                                          |            |                      |
|                    |                                          |            | 4 = Unemployed/ not  |
|                    |                                          |            |     working;         |
+--------------------+------------------------------------------+------------+----------------------+
| occupation_s3      | Working occupation (c2011_w4)            | Nominal    | 1 = Legislators and  |
|                    |                                          |            | professionals;       |
|                    |                                          |            |                      |
|                    |                                          |            | 2 = Technicians and  |
|                    |                                          |            | clerks;              |
|                    |                                          |            |                      |
|                    |                                          |            | 3 = Skilled workers; |
|                    |                                          |            |                      |
|                    |                                          |            | 4 = Plant opperators |
|                    |                                          |            | and elementary       |
|                    |                                          |            | occupations          |
|                    |                                          |            |                      |
|                    |                                          |            | 5 = Never worked;    |
+--------------------+------------------------------------------+------------+----------------------+

# Variable description

This dataset includes sociodemographic variables related to sex, age (continuous
and categorical by groups), marital status, education level, employment status
and occupation.

A unique dataset is created for these variables because of the relevance that
sociodemographic factors have in our outcome variables of interest, as 
mental health issues happens in a context modulated by sociodemographic variables.

# Explanation of how the variables were created

- `sex_s3`. Based on `q1009_sex`, we have renamed this variable. 
  
- `agegroups_s3`. Based on `q1011_age`, a continuous variable from 29-99, it is
divided in 5 groups: 18-29, 30-49, 50-64, 65-79 and 80+.

- `marital_s3`. Based on `q1012_mar_stat`, label values have been modified for
suiting 'sentence case' format.

- `education_s3`. Based on `q1016_highest`, five categories are created to
sum up education level: less than primary, primary, upper secondary, lower
secondary and tertiary.

- `employment_s3`. Based on `q1011_age`, `q1501_ever_worked`, `q1503_work_now` and
`q1504_reason`, four categories are created, indicating the employment situation 
of the person.

  - __never worked__ (1), using `q1501_ever_worked.` It also includes “homemakers”
  and “care family” from `q1504_reason`.
  - __working__ (2), using `q1503_work_now`. It also includes ”seasonal work” and 
  “vacation” from `q1504_reason`.
  - __retired__ (3), for people that is older than 65 years that have ever
  ever worked and are not working, combining q1011_age, q1501_ever_worked and
  q1503_work_now. It also includes "health", "not permitted" and "retired" from
  q1504_reason.
  - __unemployed__ (4), for people who has answered "couldn't find a job",
  "unpaid", "studies", "no need", "laid off", "doesn't want for a while" and 
  "other" from `q1504_reason`, and are under 77 years old (`q1011_age`).
  
  
- `occupation_s3`. Based on `q1510_catoccu` and `q1501_ever_worked`, five
categories are established: legislators/professionals, technicians, skilled workers,
elementary occupations and "never worked". This follows the OIT categorization 
for international comparative purposes. 

*Note: in variables `marital_s3` and `education_s3` there's one missing value
due to participant ID= 3500. He/she refused to answer questions 'q1012' to 'q1017' 
when he was recontacted.

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
*SEX
gen sex_s3 = q1009_sex
label variable sex_s3 "Sex (c2011w4)"

*AGE BY GROUPS - agegroups_s3
recode q1011_age_s3 (18/29 = 1) (30/49 = 2) (50/64 = 3) (65/79 = 4) (80/120 = 5) 
gen agegroups_s3 = q1011_age_s3
label define agegroups 1 "18-29" 2 "30-49" 3 "50-64" 4 "65-79" 5 "80+"
label value agegroups_s3 agegroups 
label variable agegroups_s3 "Age, categorical (c2011_w4)"


* MARITAL STATUS - marital_s3
gen marital_s3 = q1012_mar_stat
label define marital 1 "Never married and no cohabiting" 2 "Currently married/law-partnered" 3 "Cohabiting" 4 "Separated/divorced" 5 "Widowed"
label value marital_s3 marital
label variable marital_s3 "Marital status (c2011w4)"

*EDUCATION LEVEL - education_s3
recode q1016_highest (0/1 = 1) (5/7 = 5)
gen education_s3 = q1016_highest
label define education 1 "Less than primary" 2 "Primary" 3 "Lower secondary" 4 "Upper secondary" 5 "Tertiary"
label value education_s3 education
label variable education_s3 "Education level, categorical (c2011_w4)"

*EMPLOYMENT STATUS - employment_s3
gen employment_s3 =.

**Never worked
replace employment_s3=1 if q1501_ever_worked_s3==2 | (q1501_ever_worked_s3 == 1 & (q1504_reason==1 | q1504_reason==6))

**Working 
replace employment_s3=2 if q1503_work_now==1 | (q1503_work_now==2 & (q1504_reason==11 | q1504_reason==13))

**Retired
replace employment_s3=3 if (q1504_reason==5 | q1504_reason==8 | q1504_reason==9) | (q1011_age>65 & q1501_ever_worked_s3!=2 & employment_s3!=2)

**Unemployed
replace employment_s3=4 if (q1504_reason==2 | q1504_reason==3 | q1504_reason==4 | q1504_reason==7 | q1504_reason==10 | q1504_reason==12 | q1504_reason==87) & (q1011_age<66)

label define employment 1 "Never worked/homemaker" 2 "Working" 3 "Retired/disabled" 4"Unemployed/not working"
label value employment_s3 employment
label variable employment_s3 "Employment status (c2011_w4)"

*OCCUPATION STATUS - occupation_s3
recode q1510_catoccu (1/2 = 1) (3/4 = 2) (5/7 = 3) (8/10 = 4), gen(occupation_s3)
replace occupation_s3 =5 if q1501_ever_worked == 2
recode occupation_s3 (888 = .) (999 = .)

label define occupation 1 "Legislators and professionals" 2 "Technicians and clerks" 3 "Skilled workers" 4 "Plant operators and elementary occupations" 5 "Never worked"
label value occupation_s3 occupation
label variable occupation_s3 "Working occupation (c2011_w4)"
```

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result_s3 proxy_s3 sex_s3 agegroups_s3 marital_s3 education_s3 employment_s3 occupation_s3

codebook sex_s3 agegroups_s3 marital_s3 education_s3 employment_s3 occupation_s3
```


```{stata save-data-test, eval=TEST}
saveold "Outcome_sociodemo.dta", replace version(13)
```


```{stata save-data, eval=!TEST}
* Cohorte 2011, ola 4:
 saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_sociodemo.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Cristina Rodríguez Prada (adapted algorithm for wave 4)
-   e-mail:
    [cristina.rodriguez@inv.uam.es](mailto:cristina.rodriguez@inv.uam.es){.email}

<br>


Original version:

-   Name: Joan Domènech
-   e-mail:
    [joan.domenech@sjd.es](mailto:joan.domenech@sjd.es){.email}


::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
