---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'Stress'"
output:
  word_document:
   reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib 
csl: "`r ecs.data::get_proj_csl_file()`"
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"


# Opciones de configuración de las salidas de R y Stata:

options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( 
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

| Main variable name  | Label                       | Format     | Code |
|:--------------------|:----------------------------|:-----------|:-----|
| perceived_stress_s3 | Perceived stress (c2011_w4) | Continuous | 0-6  |

# Variable description

Although once more related to the field of biology, today the term "stress" is
commonly used in different fields such as physiology and medicine. In
psychology, stress refers to a disturbance of the homeostasis system. It is
therefore referred to as stress when the emotional (or physical) balance of an
individual is modified by an external (e.g. aggressor) or internal (e.g. virus)
agent. This modification then has an impact on certain physical and biological
aspects of the individual. It is also important to note that stress can be
negative, for example when it is too intense and impacts on the normal
functioning of the individual, but it can also be positive, for example when it
motivates and pushes the individual to run away from a problem [@lu2021].

# Explanation of how the variables were created

The level of stress was measured using 3 items corresponding to the same
question (how stressed/anguished do you feel?) asked with reference to 3
different times of the day, (1) morning (q7033), (2) evening (q7073), (3) night
(q7118). The total perceived stress score was obtained by adding the responses
given to these 3 items. The stress scale ranges from 0 to 6, the higher the
number the higher the perceived stress level.

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"

```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
* Perceived stress 
egen perceived_stress_s3 = rowtotal (q7033_tense_s3 q7073_tense_s3 q7118_tense_s3), missing
label variable perceived_stress_s3 "Perceived stress (c2011_w4)"
```

<br>

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result proxy perceived_stress
codebook perceived_stress
```

```{stata save-data-test, eval=TEST}
saveold "Outcome_stress.dta", replace version(13)
```

```{stata save-data, eval=!TEST}
* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_stress.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Lea Francia
-   e-mail: [lea.francia\@inv.uam.es](mailto:lea.francia@inv.uam.es){.email}
-   Phone number: (+34) 914975492

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
