---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'panic disorder'" 
output:
  word_document:
   reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib 
csl: "`r ecs.data::get_proj_csl_file()`"
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"


options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( 
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

+---------------+--------------------------+---------------+---------------+
| Main variable | Label                    | Format        | Code          |
| name          |                          |               |               |
+:==============+:=========================+:==============+:==============+
| panic_12m_s3  | Panic disorder during    | Categorical   | 0 = no        |
|               | last 12 months           |               |               |
|               | (c2011_w4)               |               | 1 = yes       |
+---------------+--------------------------+---------------+---------------+
| panic_3y_s3   | Inter-survey panic       | Categorical   | 0 = no        |
|               | disorder (c2011_w4)      |               |               |
|               |                          |               | 1 = yes       |
+---------------+--------------------------+---------------+---------------+

# Variable description

Panic disorder has been officially speared from Anxiety Disorder in the 1960's.
Panic disorder characterized by panic attacks (sudden feelings of terror without
rational reason). It is among the most prevalent mental disorders and it is
quite unpredictable. The trouble is associated with chronic medical conditions,
a higher burden on the health care system, and lower health-related quality of
life [@dick1994].

# Explanation of how the variables were created

An adapted version of the Composite International Diagnostic Interview (CIDI
3.0) was used to assess the presence of panic disorder [@haro2006]. An algorithm
based on the fifth edition of the Diagnostic and Statistical Manual of Mental
Disorders was used [@americanpsychiatricassociation2013] (APA, 1994).
Participants that reported crises of fear or anxiety in which they suddenly had
feelings such as fear, anxiety, anguished, restlessness, discomfort, etc., with
more than three physical symptoms such as pounding or very fast heartbeat,
feeling short of breath, nausea, or upset stomach, dizziness, sweating, shaking,
dry mouth, etc., and with two or more crises during their lifetime that had
started suddenly and got worse during the first 10 minutes, were considered to
be suffering from panic disorder.

To generate the inter-survey variable (panic disorder occurred since the last
interview, panic_3y_s3), the following criteria related to the DSM were
considered and all of them must be present:

1.  Crisis of fear or anxiety in which appears a sudden feeling such as fear,
    anxiety, anguished, restlessness, discomfort, etc. (items q4501_s3 and
    q4502_s3);
2.  More than three physical symptoms such as pounding or very fast heartbeat,
    feeling short of breath, nausea or upset stomach, dizziness, sweating,
    shaking, dry mouth, feeling of choking, chest discomfort, fear of losing
    control, feeling distant from the situation, feeling that things are not
    real, fear of losing consciousness, fear of dying, chills or feelings of
    numbness (items from q4503_s3 to q4518_s3) have to be present;
3.  Crisis started suddenly and got worse during the first 10 minutes (item
    q4519_s3);
4.  Having had two or more crises in the lifetime (item q4520_s3). Items are all
    dichotomous (1 = the symptom is present and 2 = the symptom is not present).

To generate the variable panic_12m_s3, the item q4524_s3 has been considered
(Have you had any of these crises in the last 12 months?).

# Syntax used to create the variable (Stata 15)

```{stata load-data}

* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"

```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
*# "Value labels (defined for all the outcome variables in the dataset)":
label define yesno    0 "No" 1 "Yes"

*# "Criterion 1": "Crisis of fear or anxiety in which appears a sudden feeling"
recode q4501_anxlife_s3 q4502_unclife_s3 (888=.) (999=.) (2=0)
gen     d_crit1 = .
replace d_crit1 = 0 if (q4501_anxlife_s3!=1 & q4501_anxlife_s3!=.) & (q4502_unclife_s3!=1 & q4502_unclife_s3!=.)
replace d_crit1 = 1 if  q4501_anxlife_s3==1             |  q4502_unclife_s3==1
  
*# "Criterion 2": "More than three physical symptoms"

  *#s1: "Pounding or very fast heartbeat"
recode q4503_beat_s3 (888=.) (999=.) (2=0), gen     (d_s1)

  *#s2: "Feeling short of breath"
recode q4504_air_s3 (888=.) (999=.) (2=0), gen     (d_s2)

  *#s3: "Nausea or upset stomach"
recode q4505_nausea_s3 (888=.) (999=.) (2=0), gen     (d_s3)
 
  *#s4: "Dizziness"
recode q4506_dizziness_s3 (888=.) (999=.) (2=0), gen     (d_s4)

  *#s5: "Sweating"
recode q4507_sweat_s3 (888=.) (999=.) (2=0), gen     (d_s5)

  *#s6: "Shaking/trembling" 
recode q4508_trembling_s3 (888=.) (999=.) (2=0), gen     (d_s6)

  *#s7: "Dry mouth"
recode q4509_mouth_s3 (888=.) (999=.) (2=0), gen     (d_s7)

  *#s8: "Feeling of choking" 
recode q4510_choke_s3 (888=.) (999=.) (2=0), gen     (d_s8)

  *#s9: "Chest discomfort"
recode q4511_chest_s3 (888=.) (999=.) (2=0), gen     (d_s9)

  *#s10: "Fear of losing control"
recode q4512_control_s3 (888=.) (999=.) (2=0), gen     (d_s10)
     
  *#s11: "Feeling distant from the situation"
recode q4513_deper_s3 (888=.) (999=.) (2=0), gen     (d_s11)

  *#s12: "Feeling that things are not real"
recode q4514_disreal_s3 (888=.) (999=.) (2=0), gen     (d_s12)

  *#s13: "Fear of losing consciousness"
recode q4515_cons_s3 (888=.) (999=.) (2=0), gen     (d_s13)

  *#s14: "Fear of dying"
recode q4516_dying_s3 (888=.) (999=.) (2=0), gen     (d_s14) 

  *#s15: "Suffocation"
recode q4517_suffoc_s3 (888=.) (999=.) (2=0), gen     (d_s15)

  *#s16: "Tingling"
recode q4518_tingling_s3 (888=.) (999=.) (2=0), gen     (d_s16)


*#"Sum of physical symptoms"
gen  d_symptot = d_s1 + d_s2 + d_s3 + d_s4 + d_s5 + d_s6 + d_s7 + d_s8 + d_s9 + d_s10 + d_s11 + d_s12 + d_s13 + d_s14 + d_s15 + d_s16

*#"To fulfill Criterion 2, three or more physical symptoms should be present".
gen     d_crit2 = .
replace d_crit2 = 1 if d_symptot>3 & d_symptot<17
replace d_crit2 = 0 if d_symptot<4 


*# "Criterion 3": "Crisis started suddenly and got worse during the first 10 minutes"
recode q4519_sudden_s3 (888=.) (999=.) (3=1) (2=0)
gen     d_crit3 = .
replace d_crit3 = 0 if (q4519_sudden_s3!=1 & q4519_sudden_s3!=.)
replace d_crit3 = 1 if  q4519_sudden_s3==1
  
*# "Criterion 4": "Having had two or more crises in the lifetime"
recode q4520_times_s3 (888=.) (999=.) 
gen     d_crit4 = .
replace d_crit4 = 0 if (q4520_times_s3<2 & q4520_times_s3!=.)
replace d_crit4 = 1 if  q4520_times_s3>=2 & q4520_times_s3<996

***#"Inter-survey panic disorder diagnostic variable"
gen     panic_3y_s3 = .
replace panic_3y_s3 = 0 if  d_crit1==0 | d_crit2==0 | d_crit3==0 | d_crit4==0
replace panic_3y_s3 = 1 if d_crit1==1 & d_crit2==1 & d_crit3==1 & d_crit4==1
label values   panic_3y_s3 yesno
label variable panic_3y_s3 "Inter-survey panic disorder (c2011_w4)"

***#"12 months panic disorder diagnostic variable"
recode q4524_crisis12_s3 (888=.) (999=.) (2=0)
gen panic_12m_s3 = .
replace panic_12m_s3 = 0 if panic_3y_s3 ==0 | (q4524_crisis12_s3!=1 & q4524_crisis12_s3!=.)
replace panic_12m_s3 = 1 if panic_3y_s3 ==1 & q4524_crisis12_s3==1
label values   panic_12m_s3 yesno
label variable panic_12m_s3 "Panic disorder during last 12 months (c2011_w4)"

```

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result proxy panic_12m_s3 panic_3y_s3

codebook panic_12m_s3 panic_3y_s3
```

```{stata save-data-test, eval=TEST}

saveold "Outcome_panicdisorder.dta", replace version(13)
```

```{stata save-data, eval=!TEST}

* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_panicdisorder.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Lea Francia
-   e-mail: [lea.francia\@inv.uam.es](mailto:lea.francia@inv.uam.es){.email}
-   Phone number: (+34) 914974601

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
