---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'EASI'"
output:
  word_document:
   reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib 
csl: "`r ecs.data::get_proj_csl_file()`"
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"


options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( 
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

+-----------------+----------------------+-----------------+-----------------+
| Main variable   | Label                | Format          | Code            |
| name            |                      |                 |                 |
+:================+:=====================+:================+:================+
| dependency_s3   | Dependency status    | Ordinal         | 0 = no          |
|                 | (c2011_w4)           |                 | dependence      |
|                 |                      |                 |                 |
|                 |                      |                 | 1 = dependency  |
|                 |                      |                 | served          |
|                 |                      |                 |                 |
|                 |                      |                 | 2 = problematic |
|                 |                      |                 | dependence      |
|                 |                      |                 |                 |
|                 |                      |                 | 3 = neglected   |
|                 |                      |                 | dependence      |
+-----------------+----------------------+-----------------+-----------------+
| abuse_negli_s3  | Negligence           | Ordinal         | 0 = no abuse    |
|                 | (c2011_w4)           |                 |                 |
|                 |                      |                 | 1 = occasional  |
|                 |                      |                 |                 |
|                 |                      |                 | 2 = sometimes   |
|                 |                      |                 |                 |
|                 |                      |                 | 3 = quite often |
|                 |                      |                 |                 |
|                 |                      |                 | 4 = often       |
+-----------------+----------------------+-----------------+-----------------+
| a               | Psychological abuse  | Ordinal         | 0 = no abuse    |
| buse_psycho_s3  | (c2011_w4)           |                 |                 |
|                 |                      |                 | 1 = occasional  |
|                 |                      |                 |                 |
|                 |                      |                 | 2 = sometimes   |
|                 |                      |                 |                 |
|                 |                      |                 | 3 = quite ofen  |
|                 |                      |                 |                 |
|                 |                      |                 | 4 = often       |
+-----------------+----------------------+-----------------+-----------------+
| abuse_eco_s3    | Economical abuse     | Ordinal         | 0 = no abuse    |
|                 | (c2011_w4)           |                 |                 |
|                 |                      |                 | 1 = occasional  |
|                 |                      |                 |                 |
|                 |                      |                 | 2 = sometimes   |
|                 |                      |                 |                 |
|                 |                      |                 | 3 = quite often |
|                 |                      |                 |                 |
|                 |                      |                 | 4 = often       |
+-----------------+----------------------+-----------------+-----------------+
| abuse_phys_s3   | Physical and sexual  | Ordinal         | 0 = no abuse    |
|                 | abuse (c2011_w4)     |                 |                 |
|                 |                      |                 | 1 = occasional  |
|                 |                      |                 |                 |
|                 |                      |                 | 2 = sometimes   |
|                 |                      |                 |                 |
|                 |                      |                 | 3 = quite often |
|                 |                      |                 |                 |
|                 |                      |                 | 4 = often       |
+-----------------+----------------------+-----------------+-----------------+

# Variable description

The Spanish version of the Elder Abuse Suspicion Index [@pérez-rojo2008] is an
adaptation of the Elder Abuse Suspicion Index (EASI), a brief instrument with
direct questions aimed at detecting suspicion of abuse in older people without
cognitive impairment. It is made up of five questions addressed to the elderly
person and one to the professional, which are asked in the course of an
encounter between the elderly person and the professional in the consultation
room and are formulated by the professional using simple language. The first
question focuses on the assessment of the level of dependency of the elderly
person for some basic or instrumental activity of daily living. The remaining
questions refer to different types of abuse: physical and sexual, psychological,
financial, and neglect. The items are assessed on a Likert-type scale with
response options between 0 and 4 (0 = no, 1 = occasional, 2 = sometimes, 3 =
quite often, and 4 = often). A positive response to any of the questions would
indicate suspicion of maltreatment. This instrument has to be used for
participants aged +64.

# Explanation of how the variables were created

Ten items was used to created the EASI scale. These items ask information about
different types of maltreatment: neglect (items q6702_s3, q6702a_s3),
psychological (items q6703_s3, q6703a_s3), economic (q6704_s3, q6704a_s3) and
physical-sexual (q6705_s3, q6705a_s3). For the four type of maltreatment a total
scale was obtain by summing the answer given to each items. The scale range from
0 no abused to 4 often abused. The dependency scale was obtain using tree items
(q6701_s3, q6701b_s3, and q6701d_s3) and ranges from 0 no dependency to 3
neglected dependency. The questions refer to situations that occurred in the 12
months prior to the completion of the questionnaire and only persons aged 65 and
over are asked.

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
*Dependency

recode q6701_s3 (1=1) (2=0) (else=.)
recode q6701b_s3 (1=1) (2=0) (else=.)
recode q6701d_s3 (1=1) (2=0) (else=.)
gen dependency_s3=0 if q6701_s3==0
replace dependency_s3=1 if q6701_s3==1
replace dependency_s3=2 if q6701b_s3==1
replace dependency_s3=3 if q6701_s3==0 & q6701d_s3==1
label define dependency_l 0 "no dependency" 1 "dependency served" 2 "problematic dependency served" 3 "neglected dependency"
label values dependency_s3 dependency_l
label variable dependency_s3 "Dependency status (c2011_w4)"

* Negligence 

recode q6702_s3 (1=1) (2=0) (else=.)
recode q6702a_s3 (1=1) (2=2) (3=3) (4=4)  (else=.)
gen abuse_negli_s3=q6702a_s3
replace abuse_negli_s3=0 if q6702_s3==0
label define abuse 0 "no abuse" 1 "occasionnal" 2 "sometimes" 3 "quite often" 4 "often"
label values abuse_negli_s3 abuse
label variable abuse_negli_s3 "Negligence (c2011_w4)"

* Psychological abuse 

recode q6703_s3 (1=1) (2=0) (else=.)
recode q6703a_s3 (1=1) (2=2) (3=3) (4=4)  (else=.)
gen abuse_psycho_s3 =q6703a_s3
replace abuse_psycho_s3=0 if q6703_s3==0
label values abuse_psycho_s3 abuse
label variable abuse_psycho_s3 "Psychological abuse (c2011_w4)"

* Economical abuse

recode q6704_s3 (1=1) (2=0) (else=.)
recode q6704a_s3 (1=1) (2=2) (3=3) (4=4)  (else=.)
gen abuse_eco_s3 =q6704a_s3
replace abuse_eco_s3=0 if q6704_s3==0
label values abuse_eco_s3 abuse
label variable abuse_eco_s3 "Economical abuse (c2011_w4)"

* Physical and sexual abuse

recode q6705_s3 (1=1) (2=0) (else=.)
recode q6705a_s3 (1=1) (2=2) (3=3) (4=4)  (else=.)
gen abuse_phys_s3=q6705a_s3
replace abuse_phys_s3=0 if q6705_s3==0
label values abuse_phys_s3 abuse
label variable abuse_phys_s3 "Physical and sexual abuse (c2011_w4)"

```

<br>

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result proxy dependency_s3 abuse_negli_s3 abuse_psycho_s3 abuse_eco_s3 abuse_phys_s3

codebook dependency_s3 abuse_negli_s3 abuse_psycho_s3 abuse_eco_s3 abuse_phys_s3
```

```{stata save-data-test, eval=TEST}

saveold "Outcome_EASI.dta", replace version(13)
```

```{stata save-data, eval=!TEST}

* Cohorte 2011, ola 4:
* saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_EASI.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Lea Francia
-   e-mail: [lea.francia\@inv.uam.es](mailto:lea.francia@inv.uam.es){.email}
-   Phone number: (+34) 914974601

Original version:

-   Name: Dario Moreno Agostino
-   e-mail: [d.moreno\@ucl.ac.uk](mailto:d.moreno@ucl.ac.uk){.email}

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
