---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'Suspicion of abuse'"
output:
  word_document:
   reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib 
csl: "`r ecs.data::get_proj_csl_file()`"
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"


options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( 
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

+-------------------+----------------+--------------------+---------------------+
| Main variable     | Label          | Format             | Code                |
| name              |                |                    |                     |
+:==================+:===============+:===================+:====================+
| dependency_s3     | Dependency     | Nominal            | -   0 = No          |
|                   | status         |                    |     dependence      |
|                   | comparable     |                    |                     |
|                   | (c2011_w4)     |                    | -   1 = Dependency  |
|                   |                |                    |     served          |
|                   |                |                    |                     |
|                   |                |                    | -   2 = Problematic |
|                   |                |                    |     dependence      |
|                   |                |                    |                     |
|                   |                |                    | -   3 = Neglected   |
|                   |                |                    |     dependence      |
+-------------------+----------------+--------------------+---------------------+
| abuse_negli_s3    | Negligence     | Ordinal            | -   0 = No abuse    |
|                   | abuse          |                    |                     |
|                   | comparable     |                    | -   1 = Occasional  |
|                   | (c2011_w4)     |                    |                     |
|                   |                |                    | -   2 = Sometimes   |
|                   |                |                    |                     |
|                   |                |                    | -   3 = Quite often |
|                   |                |                    |                     |
|                   |                |                    | -   4 = Often       |
+-------------------+----------------+--------------------+---------------------+
| abuse_psycho_s3   | Psychological  | Ordinal            | -   0 = No abuse    |
|                   | abuse          |                    |                     |
|                   | comparable     |                    | -   1 = Occasional  |
|                   | (c2011_w4)     |                    |                     |
|                   |                |                    | -   2 = Sometimes   |
|                   |                |                    |                     |
|                   |                |                    | -   3 = Quite often |
|                   |                |                    |                     |
|                   |                |                    | -   4 = Often       |
+-------------------+----------------+--------------------+---------------------+
| abuse_eco_s3      | Economical     | Ordinal            | -   0 = No abuse    |
|                   | abuse          |                    |                     |
|                   | comparable     |                    | -   1 = Occasional  |
|                   | (c2011_w4)     |                    |                     |
|                   |                |                    | -   2 = Sometimes   |
|                   |                |                    |                     |
|                   |                |                    | -   3 = Quite often |
|                   |                |                    |                     |
|                   |                |                    | -   4 = Often       |
+-------------------+----------------+--------------------+---------------------+
| abuse_phys_s3     | Physical and   | Ordinal            | -   0 = No abuse    |
|                   | sexual abuse   |                    |                     |
|                   | comparable     |                    | -   1 = Occasional  |
|                   | (c2011_w4)     |                    |                     |
|                   |                |                    | -   2 = Sometimes   |
|                   |                |                    |                     |
|                   |                |                    | -   3 = Quite often |
|                   |                |                    |                     |
|                   |                |                    | -   4 = Often       |
+-------------------+----------------+--------------------+---------------------+
| abusesuspicion_s3 | Suspicion of   | Dichotomic         | -   0 = No          |
|                   | abuse          |                    |                     |
|                   | (c2011_w4)     |                    | -   1 = Yes         |
+-------------------+----------------+--------------------+---------------------+

# Variable description

The Spanish version of the Elder Abuse Suspicion Index [@pérez-rojo2008] is an
adaptation of the Elder Abuse Suspicion Index (EASI), a brief instrument with
direct questions aimed at detecting suspicion of abuse in older people (aged
+64) without cognitive impairment. It is made up of five questions addressed to
the elderly person and one to the professional, which are asked in the course of
an encounter between the elderly person and the professional in the consultation
room and are formulated by the professional using simple language. The first
question focuses on the assessment of the level of dependency of the elderly
person for some basic or instrumental activity of daily living. The remaining
questions refer to different types of abuse: physical and sexual, psychological,
financial, and neglect. The items are dichotomous (0=No; 1=Yes). A positive
response to any of the questions would indicate suspicion of maltreatment.

# Explanation of how the variables were created

At wave 4 we realize that the material cited by EASI does not indicate the
instructions that have been followed in previous waves to create these outcomes.
However, for the sake of consistency and comparability between waves, we decided
to follow the same scheme:

-   Create a variable indicating the **degree of dependency** (`dependency_s3`),
    based on the block of questions `q6701_s3` (help from someone), `q6701b_s3`
    (problems with them) and `q6701d_s3` (need help). In this way, we found four
    categories:

    -   *No dependence*: has neither received help nor needed it.
    -   *Dependency served*: had help but did not know if he/she needed it, did
        not have help but needed it
    -   *Problematic dependence*: had help, had problems with the person who
        helped them and they needed that help;
    -   *Neglected dependence*: did not have help, had problems with who was
        supposed to help them and they needed that help.

-   Create four variables for **each specific type of abuse**. For these
    variables, the first of each pair indicated below is selected, which is a
    dichotomous item on the presence or absence of that type of abuse, and is
    complemented with the second item of the pair, which indicates the
    frequency.

    -   `abuse_negli_s3`, negligence abuse: `q6702_s3`, `q6702a_s3`
    -   `abuse_psycho_s3`, psychological abuse: `q6703_s3`, `q6703a_s3`
    -   `abuse_eco_s3`, economical abuse: `q6704_s3`, `q6704a_s3`
    -   `abuse_phys_s3`, physical and sexual abuse: `q6705_s3`, `q6705a_s3`

To create the abusesuspicion_s3 variable, we keep only the 5 items addressed to
the participants and exclude the question asked to the professional(s). These 4
items ask for information about different types of maltreatment: neglect
(`q6702_s3`), psychological (`q6703_s3`), economic (`q6704_s3`) and
physical-sexual (`q6705_s`). The item `q6701_s3` indicates if the person has
needed help, so in the case of responding 'yes' it could be some suspicion of
abuse. The total scale of suspicion of abuse was obtained by summing the answer
given to each 4 items. A transformation of the scores was then made, for those
participants who answer yes to one of the five items a suspicion of abuse is
considered. The scale ranges from 0: no suspicion of abuse to 1: suspicion of
abuse. The questions refer to situations that occurred in the 12 months before
the completion of the questionnaire and only persons aged 65 and over were
asked.

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~/UAM/Marta Miret Garcia - Bases de datos maestras Edad con Salud/Ola_4/Cohorte 2011/rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
*DEPENDENCY
*Dependency (comparable)
recode q6701_s3 q6701d_s3 q6701b_s3 (2=0) (999=.) 
gen dependency_s3=. 

**0 = No dependence; has not had any help and has not needed it:
replace dependency_s3=0 if q6701_s3==0 & q6701d_s3==0 

**1 = Dependency served; had any help but NA in had needed it OR NA in had any help and had needed it
replace dependency_s3=1 if (q6701_s3==1 & q6701d_s3==0) | (q6701_s3==0 & q6701d_s3==1) 

**2 = Problematic dependence; had any help, had problems with people who helped them, and needed help
replace dependency_s3=2 if (q6701_s3==1 & q6701b_s3==1 & q6701d_s3==0) |  (q6701_s3==1 & q6701b_s3==0 & q6701d_s3==1)

**3 = Neglected dependence; had no help, had problems with people who helped them, and needed help
replace dependency_s3=3 if q6701_s3==1 & q6701b_s3==1 & q6701d_s3==1 

label define dep 0 "No dependence" 1 "Dependency served" 2 "Problematic dependence" 3 "Neglected dependence"  
label values dependency_s3 dep 
label variable dependency_s3 "Dependency status comparable (c2011_w4)" 


*NEGLIGENCE
*abuse_negli_s3 "Negligence (c2011_w4)"	 - q6702_s3, q6702a_s3 (comparable)

recode q6702_s3 (2=0) (999=.) 
recode q6702a_s3 (999=.) 

gen abuse_negli_s3=. 
**0 = No abuse; (dichotomic item)
replace abuse_negli_s3=0 if q6702_s3==0

**1 = Occasional abuse; 1 in likert scale item
replace abuse_negli_s3=1 if q6702a_s3==1 
**2 = Sometimes; 2 in likert scale item
replace abuse_negli_s3=2 if q6702a_s3==2 
**3 = Quite often; 3 in likert scale item
replace abuse_negli_s3=3 if q6702a_s3==3 
**4 = Often; 4 in likert scale item
replace abuse_negli_s3=4 if q6702a_s3==4 

 
label define ab 0 "No abuse" 1 "Occasional" 2 "Sometimes" 3 "Quite often" 4 "Often" 
label values abuse_negli_s3 ab 
label variable abuse_negli_s3 "Negligence comparable (c2011_w4)" 

*PSYCHOLOGICAL ABUSE
*abuse_psycho_s3 Psychological abuse (c2011_w4) - q6703_s3, q6703a_s3 

recode q6703_s3 (2=0) (999=.) 
recode q6703a_s3 (999=.) 

gen abuse_psycho_s3=. 

**0 = No abuse; (dichotomic item)
replace abuse_psycho_s3=0 if q6703_s3==0 

**1 = Occasional abuse; 1 in likert scale item
replace abuse_psycho_s3=1 if q6703a_s3==1 
**2 = Sometimes; 2 in likert scale item
replace abuse_psycho_s3=2 if q6703a_s3==2 
**3 = Quite often; 3 in likert scale item
replace abuse_psycho_s3=3 if q6703a_s3==3 
**4 = Often; 4 in likert scale item
replace abuse_psycho_s3=4 if q6703a_s3==4 

label values abuse_psycho_s3 ab 
label variable abuse_psycho_s3 "Psychological abuse comparable (c2011_w4)" 

 

*ECONOMICAL ABUSE
*abuse_eco_s3 Economical abuse (c2011_w4)- q6704_s3, q6704a_s3 
recode q6704_s3 (2=0) (999=.) 
recode q6704a_s3 (999=.) 

gen abuse_eco_s3=. 
**0 = No abuse; (dichotomic item)
replace abuse_eco_s3=0 if q6704_s3==0

**1 = Occasional abuse; 1 in likert scale item
replace abuse_eco_s3=1 if q6704a_s3==1 
**2 = Sometimes; 2 in likert scale item
replace abuse_eco_s3=2 if q6704a_s3==2 
**3 = Quite often; 3 in likert scale item
replace abuse_eco_s3=3 if q6704a_s3==3 
**4 = Often; 4 in likert scale item
replace abuse_eco_s3=4 if q6704a_s3==4 

label values abuse_eco_s3 ab 
label variable abuse_eco_s3 "Economical abuse comparable (c2011_w4)" 

 

**abuse_phys_s3	Physical and sexual abuse (c2011_w4) - q6705_s3, q6705a_s3 

recode q6705_s3 (2=0) (999=.) 
recode q6705a_s3 (999=.) 

gen abuse_phys_s3=. 
**0 = No abuse; (dichotomic item)
replace abuse_phys_s3=0 if q6705_s3==0 
**1 = Occasional abuse; 1 in likert scale item
replace abuse_phys_s3=1 if q6705a_s3==1 
**2 = Sometimes; 2 in likert scale item
replace abuse_phys_s3=2 if q6705a_s3==2 
**3 = Quite often; 3 in likert scale item
replace abuse_phys_s3=3 if q6705a_s3==3 
**4 = Often; 4 in likert scale item
replace abuse_phys_s3=4 if q6705a_s3==4 

 
label values abuse_phys_s3 ab 
label variable abuse_phys_s3 "Physical and sexual abuse comparable (c2011_w4)" 



*SUSPICION OF ABUSE (new, correct version)
gen abuse_tot = (q6701_s3 + q6702_s3 + q6703_s3 + q6704_s3 + q6705_s3)

gen abusesuspicion_s3 =.
replace abusesuspicion_s3=0 if abuse_tot<1 
replace abusesuspicion_s3=1 if abuse_tot>0 & abuse_tot!=.

label define abusesuspicion_l 0 "No suspicion of abuse" 1 "Suspicion of abuse"
label values abusesuspicion_s3 abusesuspicion_l
label variable abusesuspicion_s3 "Suspicion of abuse (c2011_w4)"
```

<br>

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result proxy dependency_s3 abuse_negli_s3 abuse_psycho_s3 abuse_eco_s3 abuse_phys_s3 abusesuspicion_s3
codebook dependency_s3 abuse_negli_s3 abuse_psycho_s3 abuse_eco_s3 abuse_phys_s3 abusesuspicion_s3
```

```{stata save-data-test, eval=TEST}

saveold "Outcome_suspicionofabuse.dta", replace version(13)
```

```{stata save-data, eval=!TEST}

* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_suspicionofabuse.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Lea Francia

-   e-mail: [lea.francia\@inv.uam.es](mailto:lea.francia@inv.uam.es){.email}

-   Phone number: (+34) 914974601

-   Name: Cristina Rodríguez Prada

-   e-mail:
    [cristina.rodriguez\@inv.uam.es](mailto:cristina.rodriguez@inv.uam.es){.email}

Original version:

-   Name: Dario Moreno Agostino
-   e-mail: [d.moreno\@ucl.ac.uk](mailto:d.moreno@ucl.ac.uk){.email}

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
