---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'grip strength'" 
output:
  word_document:
   reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib 
csl: "`r ecs.data::get_proj_csl_file()`"
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"


options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( 
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

+----------------+-------------------------+----------------+----------------+
| Main variable  | Label                   | Format         | Code           |
| name           |                         |                |                |
+:===============+:========================+:===============+:===============+
| grip_max_s3    | Maximum grip strength   | Quantitative   | 0,5 - 99,0     |
|                | [kg] (c2011_w4)         |                |                |
+----------------+-------------------------+----------------+----------------+
| grip_s3        | Grip strength           | Categorical    | 0 = Normal     |
|                | (c2011_w4)              |                |                |
|                |                         |                | 1 = Weak       |
+----------------+-------------------------+----------------+----------------+

# Variable description

This variable classifies the respondents in a subject with a normal grip
strength and a subject with a weak grip strength. Hand-grip strength has been
measured using the Smedley Hand Dynamometer (Stoelting Company, Wood Dale,
Illinois,United States). While each participant was sitting, he/she extended
his/her arm in front of him/her and gripped the dynamometer with his/her
dominant hand. The width of the handle was adjusted, so that, when the subject
held the dynamometer, the second joint of the forefinger was in 90 degree
flexion. Two trials, with brief pauses, were allowed unless the difference
between the two measures was more than 10%; in that case, the test was repeated
once more after a rest period. Participants were encouraged to exert their
maximal grip. The best result will be chosen for analyses.

# Explanation of how the variables were created

A dichotomous variable categorising grip strength in *Normal* and *Weak* was
generated, based on the information about the trials collected in the items
q2521_first_s3, q2522_second_s3 and q2522a_third_s3. The best grip strength
value was taken and the following algorithm in Stata was used to create the two
categories (0 = Normal, 1 = Weak), depending on sex and body
mass index (expressed as ratio).
Note: To create this variable it is necessary to combine the whole raw dataset 
with the "anthotropometric measures" dataset to be able to use the previously 
created variable relating to BMI (bmi_s3). For participants who decide not to 
answer (999) the answer was consider as missing (.).

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"

```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
/* First, we would need to take the BMI variable from the anthropometrics measures outcomes dataset */
merge 1:1 ID_ECS using "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_anthropometricmeasures.dta"

recode q2521_first_s3 q2522_second_s3 q2522a_third_s3 (999 =.)
generate grip_max_s3=q2521_first_s3 if q2521_first_s3>= q2522_second_s3 | q2522_second_s3==.
replace grip_max_s3=q2522_second_s3 if q2522_second_s3> q2521_first_s3 | q2521_first_s3==.
replace grip_max_s3= q2522a_third_s3 if q2521_first_s3>100

generate grip_s3=1 if (q1009_sex==1 & bmi_s3>0 & bmi_s3<=24 & grip_max_s3<=29)|( q1009_sex==1 & bmi_s3>24 & bmi_s3<=26 & grip_max_s3<=30)| (q1009_sex==1 & bmi_s3>26 & bmi_s3<=28 & grip_max_s3<=30) | (q1009_sex==1 & bmi_s3>28 & bmi_s3<=40 & grip_max_s3<=32)

replace grip_s3=1 if (q1009_sex==2 & bmi_s3>0 & bmi_s3<=23 & grip_max_s3<=17)|( q1009_sex==2 & bmi_s3>23 & bmi_s3<=26 & grip_max_s3<=17.3)| (q1009_sex==2 & bmi_s3>26 & bmi_s3<=29 & grip_max_s3<=18) | (q1009_sex==2 & bmi_s3>29 & bmi_s3<=40 & grip_max_s3<=21)

replace grip_s3=0 if grip_s3==.

replace grip_s3=. if (q2521_first_s3==.| q2521_first_s3<0 | q2522_second_s3==. | q2522_second_s3 < 0 | bmi_s3>=50 | bmi_s3==.) | (q2521_first_s3>100 & q2522_second_s3>100)

label define grip 0 "Normal" 1 "Weak"
label values grip_s3 grip
label variable grip_s3 "Grip strength at wave 4 (c2011_w4)"

```

<br>

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}

keep ID_ECS q0007a_result proxy grip_max_s3 grip_s3

codebook grip_max_s3 grip_s3
```

```{stata save-data-test, eval=TEST}

saveold "Outcome_gripstrength.dta", replace version(13)
```

```{stata save-data, eval=!TEST}
* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_gripstrength.dta", replace version(13)

```

<br>

# Contact person:

-   Name: Lea Francia
-   e-mail:
    [\<lea.francia\@inv.uam.es\>](mailto:nombre.buzon@institucion.org){.email}
-   Phone number: (+34) 914974601

<br>

Original version:

-   Name: Félix Caballero
-   e-mail:
    [\<felix.caballero\@uam.es\>](mailto:nombre.buzon@institucion.org){.email}
-   Phone number: (+34)914975088

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
