---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'ALCOHOL CONSUMPTION'" 
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
# bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib ## TODO: Descomentar esta línea SI se utiliza el gestor bibliográfico, en otro caso BORRAR esta línea
csl: "`r ecs.data::get_proj_csl_file()`"
# nocite: '@<cita_1>, @<cita_2>' ## TODO: Descomentar y añadir cita(s) o borrar campo
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

# Ruta al ejecutable de Stata (Ejecuta `sysdir` en Stata para encontrar la ruta)

## TODO: sustituir '<path\to\>' por la ubicación de Stata
##       (ejecutar `sysdir` en Stata) y descomentar línea 62,
##       o eliminar linea 62:
STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"


# Opciones de configuración de las salidas de R y Stata:

options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( # Para todos los 'chunks' de código:
  ## TODO: Descomentar si hace falta
  # engine.path = list(stata = STATA_PATH),
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

+--------------------+---------------------------------------------+-------------+--------------------------------+
| Main variable name | Label                                       | Format      | Code                           |
+:===================+:============================================+:============+:===============================+
| alcohol_s3         | Alcohol consumption, categorical (c2011_w4) | Ordinal     | 0 = Lifetime abstainers;       |
|                    |                                             |             |                                |
|                    |                                             |             | 1 = Occasional drinkers;       |
|                    |                                             |             |                                |
|                    |                                             |             | 2 = Non-heavy drinkers;        |
|                    |                                             |             |                                |
|                    |                                             |             | 3 = Infrequent heavy drinkers; |
|                    |                                             |             |                                |
|                    |                                             |             | 4 = Frequent heavy drinkers;   |
+--------------------+---------------------------------------------+-------------+--------------------------------+

# Variable description

This variable describes the alcohol consumption habit of the respondents.

# Explanation of how the variables were created

Five categories have been created, as follows:

-   **Life-time abstainers** (never consumed alcohol) 
- **Occasional drinkers** (did not consume alcohol in the last 30 days or in the
last 7 days);
-   **Non-heavy drinkers** (did consume alcohol in the last 30 days and in 
the last 7 days);
-   **Infrequent heavy drinkers** (did consume alcohol 1-2 days per week, with 
5 or more standard drinks in last 7 days for men and 4 or more for women);
-   **Frequent heavy drinkers** (did consume alcohol 3 or more days per week 
with 5 or more standard drinks in last 7 days for men and 4 or more for women).

Items used: `q3007_everalc_s3` _(ever consumed alcohol?)_; `q3008_last30_s3` 
_(alcohol last 30 days?)_; `q3008b_last7_s3` _(alcohol last 7 days?)_ and in which
day it was (monday-sunday): `q3009a_s3` `q3009b_s3` `q3009c_s3` `q3009d_s3` 
`q3009e_s3` `q3009f_s3` `q3009g_s3`.


# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
**Variables used: q3007_everalc_s3 q3008_last30_s3 q3008b_last7_s3 q3009a_s3 q3009b_s3 q3009c_s3 q3009d_s3 q3009e_s3 q3009f_s3 q3009g_s3**
recode q3007_everalc_s3 (888=.), gen (q3007)
recode q3008b_last7_s3 (888=.), gen (q3008b)
recode q3009a_s3 (888=.) (999=.), gen (q3009a)
recode q3009b_s3 (888=.) (999=.), gen (q3009b)
recode q3009c_s3 (888=.) (999=.), gen (q3009c)
recode q3009d_s3 (888=.) (999=.), gen (q3009d)
recode q3009e_s3 (888=.) (999=.), gen (q3009e)
recode q3009f_s3 (888=.) (999=.), gen (q3009f)
recode q3009g_s3 (888=.) (999=.), gen (q3009g)


**Count the number of missings in variables q3009a-q3009f

egen missq3009= rowmiss(q3009a q3009b q3009c q3009d q3009e q3009f q3009g)


**Exclude cases that have missing in 5 or more days
**Days of the week on which they drink 5 or more units of alcohol (men) and 4 or more units of alcohol (women)

egen daysalc5= anycount (q3009a q3009b q3009c q3009d q3009e q3009f q3009g), v(5/40)
egen daysalc4= anycount (q3009a q3009b q3009c q3009d q3009e q3009f q3009g), v(4/40)

**the answer for q3007 (ha bebido alguna vez alcohol ("have you ever drunk alcohol")) it will be "no" if in q3008 they answer that they have "never drank" (“nunca han bebido”)
recode q3007 (1=2) if q3008_last30_s3==3
recode q3008_last30_s3 (3=.), gen (q3008)

**Create separate variables for men and women and then combine them.
**men
gen alcoholm_s3=. 
replace alcoholm_s3=0 if q3007==2 & q1009_sex==1
replace alcoholm_s3=1 if q3007==1 & (q3008==2 | q3008b==2) & q1009_sex==1
replace alcoholm_s3=2 if q3007==1 & (q3008==1 & q3008b==1) & (daysalc5<1) & q1009_sex==1
replace alcoholm_s3=3 if q3007==1 & (q3008==1 & q3008b==1) & missq3009<5 & (daysalc5>0 & daysalc5<3)  & q1009_sex==1
replace alcoholm_s3=4 if q3007==1 & (q3008==1 & q3008b==1) & missq3009<5 & (daysalc5>2) & q1009_sex==1

**women
gen alcoholw_s3=. 
replace alcoholw_s3=0 if q3007==2 & q1009_sex==2
replace alcoholw_s3=1 if q3007==1 & (q3008==2 | q3008b==2) & q1009_sex==2
replace alcoholw_s3=2 if q3007==1 & (q3008==1 & q3008b==1) & (daysalc5<1) & q1009_sex==2
replace alcoholw_s3=3 if q3007==1 & (q3008==1 & q3008b==1) & missq3009<5 & (daysalc4>0 & daysalc4<3)  & q1009_sex==2
replace alcoholw_s3=4 if q3007==1 & (q3008==1 & q3008b==1) & missq3009<5 & (daysalc4>2) & q1009_sex==2

gen alcohol_s3= alcoholm_s3 if q1009_sex==1
replace alcohol_s3=alcoholw_s3 if q1009_sex==2

label define alc_labels_value 0 "Lifetime abstainers" 1 "Occasional drinkers" 2 "Non-heavy drinkers" 3 "Infrequent heavy drinkers" 4 "Frequent heavy drinkers"
label values alcohol_s3 alc_labels_value
label variable alcohol_s3 "Alcohol consumption, categorical (c2011_w4)"
```

<br> 

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result_s3 proxy_s3 alcohol_s3
codebook alcohol_s3
```

```{stata save-data-test, eval=TEST}
saveold "Outcome_alcoholconsumption.dta", replace version(13)
```

```{stata save-data, eval=!TEST}
* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_alcoholconsumption.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Cristina Rodríguez Prada (_adapted algorithm for Wave 4_)
-   e-mail: [cristina.rodriguez@inv.uam.es](mailto:cristina.rodriguez@inv.uam.es){.email}

<br>


Original version:

-   Name: Beatriz Olaya
-   e-mail: [beatriz.olaya@pssjd.org](mailto:beatriz.olaya@pssjd.org){.email}


\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
