---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'anxiety_DSM5'" ## TODO: Añadir "cohort", "assessment" y "variable name(s)/domain"
output:
  word_document:
   reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib 
csl: "`r ecs.data::get_proj_csl_file()`"
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"

options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( 
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

+--------------------+--------------------------+-------------+-------+
| Main variable name | Label                    | Format      | Code  |
+:===================+:=========================+:============+:======+
| gad_12m_s3         | Generalized anxiety      | Catehorical | 0 no  |
|                    | disorder during last 12  |             |       |
|                    | months (c2011_w4)        |             | 1 yes |
+--------------------+--------------------------+-------------+-------+
| gad_s3             | Lifetime generalized     | Categorical | 0 no  |
|                    | anxiety disorder         |             |       |
|                    | (c2011_w4)               |             | 1 yes |
+--------------------+--------------------------+-------------+-------+

# Variable description

Generalized Anxiety Disorder is among the most prevalent mental disorders and is
associated with chronic medical conditions, a higher burden on the health care
system, and lower health-related quality of life.

# Explanation of how the variables were created

An adapted version of the Composite International Diagnostic Interview (CIDI
3.0) was used to assess the presence of generalized anxiety disorders
[@haro2006]. An algorithm based on the fifth edition of the Diagnostic and
Statistical Manual of Mental Disorders was used
[@americanpsychiatricassociation2013]. Participants that reported feelings of
worry or anxiety, with excessive concerns for more than one aspect of their life
during 6 months or over, and with further physical and mental problems related
to anxiety -- of which at least one may be nerves on edge, tiredness,
irritability, concentration difficulty, muscle pain, or insomnia -- were
considered to be suffering from Generalized Anxiety Disorder.

More exactly the criteria used to created the syntaxe was the following ones: A.
Excessive anxiety and worry for at least six months (q4552_s3).

B. Difficulty to control the worry (q4556_s3).

C. At least three of the following six symptoms: -Restlessness or feeling keyed
up or on edge (q4560a_S3) -Being easily fatigued (q4560b_S3) -Irritability
(q4560c_S3) -Difficulty concentrating or mind going blank (q4560d_S3) -Muscle
tension (q4560e_S3) -Sleep disturbance (difficulty falling or staying asleep, or
restless unsatisfying sleep) (q4560e_S3)

D. The worries are related to diffuse concerns such as personal problems, phobic
and obsessive-compulsive situations, social problems (q4554_s3)

E. The symptoms are not attributable to psychoactive substance use or to any
organic mental disorder

NB: There is no major difference between the DSM-4 and DSM-5 critera. See
<https://www.ncbi.nlm.nih.gov/books/NBK519704/table/ch3.t15/> for more
information.

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
*# "Value labels (defined for all the outcome variables in the dataset)":
label define yesno    0 "No" 1 "Yes"

*#Criterion A: "Excessive anxiety and worry for at least six months"
gen gad_crit_a = .
replace gad_crit_a = 0 if (q4552_6months_s3!=1 & q4552_6months_s3!=.) 
replace gad_crit_a = 1 if  q4552_6months_s3==1   

*#Criterion B: "Difficulty to control the worry"
gen gad_crit_b = .
replace gad_crit_b = 0 if (q4556_control_s3!=1 & q4556_control_s3!=.) 
replace gad_crit_b = 1 if  q4556_control_s3==1   

*#Criterion C: "At least three of the following six symptoms":
*#s1: "Restlessness or feeling keyed up or on edge" 
  gen gad_s1 = .
replace gad_s1 = 0 if (q4560a_restless_S3!=1 & q4560a_restless_S3!=.) 
replace gad_s1 = 1 if  q4560a_restless_S3==1  
*#s2: "Being easily fatigued"
  gen gad_s2 = .
replace gad_s2 = 0 if (q4560b_tiredness_S3!=1 & q4560b_tiredness_S3!=.) 
replace gad_s2 = 1 if  q4560b_tiredness_S3==1  
*#s3: "Irritability"
  gen gad_s3 = .
replace gad_s3 = 0 if (q4560c_irritable_S3!=1 & q4560c_irritable_S3!=.) 
replace gad_s3 = 1 if  q4560c_irritable_S3==1  
*#s4: "Difficulty concentrating or mind going blank" 
  gen gad_s4 = .
replace gad_s4 = 0 if (q4560d_concentration_S3!=1 & q4560d_concentration_S3!=.) 
replace gad_s4 = 1 if  q4560d_concentration_S3==1  
*#s5: "Muscle tension"
  gen gad_s5 = .
replace gad_s5 = 0 if (q4560e_tight_S3!=1 & q4560e_tight_S3!=.) 
replace gad_s5 = 1 if  q4560e_tight_S3==1  
*#s6: "Sleep disturbance"
  gen gad_s6 = .
replace gad_s6 = 0 if (q4560f_sleep_S3!=1 & q4560f_sleep_S3!=.) 
replace gad_s6 = 1 if  q4560f_sleep_S3==1  

*#"Sum of depressive symptoms"
egen d_symptot = rowtotal (gad_s1 gad_s2 gad_s3 gad_s4 gad_s5 gad_s6), missing

*#"To fulfill Criterion C, at least three symptoms should be present".
gen     gad_crit_c = .
replace gad_crit_c= 0 if d_symptot<3 
replace gad_crit_c = 1 if d_symptot>=3 & d_symptot!=.

*#Criterion D: "Worries are related to diffuse concerns"
  gen w1 = .
replace w1 = 0 if (q4554_worry1_s3!=1 & q4554_worry1_s3!=.) 
replace w1 = 1 if  q4554_worry1_s3==1  

  
*#Criterion E: "The episode is not attributable to psychoactive substance use or
*#             to any organic mental disorder"
gen     gad_crit_e = .
replace gad_crit_e = 0 if                            q4224b_s3==1              | (q4224_s3==1 & (q4225_causemed12_s3==1 | q4225_causemed23_s3==1 | q4225_causemed24_s3==1 | q4225_causemed25_s3==1 | q4225_causemed30_s3==1 | q4225_causemed31_s3==1))
replace gad_crit_e = 1 if (q4224a_s3!=1 & q4224a_s3!=.) | (q4224b_s3!=1 & q4224b_s3!=.) | (q4224_s3==1 & (q4225_causemed12_s3!=1 & q4225_causemed23_s3!=1 & q4225_causemed24_s3!=1 & q4225_causemed25_s3!=1 & q4225_causemed30_s3!=1 & q4225_causemed31_s3!=1))
*#"NOTE: When `q4224b==1` it is necessarily the case that `q4224a==1`,
*#       so only `q4224b` out of these two needs to be checked for considering
*#       the criterion negative."

***#"Lifetime generalized anxiety disorder diagnostic variable":
gen     gad_s3  = .
replace gad_s3  = 0 if gad_crit_a==0 | gad_crit_b==0 | gad_crit_c==0 | gad_crit_d==0 | gad_crit_e==0
replace gad_s3  = 1 if gad_crit_a==1 & gad_crit_b==1 & gad_crit_c==1 & gad_crit_d==1 & gad_crit_e==0
label values   gad_s3  yesno
label variable gad_s3  "Lifetime generalized anxiety disorder (c2011_w4)"

***#"12 months generalized anxiety disorder diagnostic variable":
  gen gad_12m_s3 = .
replace gad_12m_s3 = 0 if (q4573_s3!=1 & q4573_s3!=.) 
replace gad_12m_s3 = 1 if  q4573_s3==1 
label values   gad_12m_s3  yesno
label variable gad_12m_s3  "Generalized anxiety disorder during last 12 months (c2011_w4)"
  
```

<br>

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result proxy gad_s3 gad_12m_s3

codebook gad_s3 gad_12m_s3
```

```{stata save-data-test, eval=TEST}
saveold "Outcome_anxiety_DSM5.dta", replace version(13)
```

```{stata save-data, eval=!TEST}
* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_anxiety_DSM5.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Lea Francia
-   e-mail: [lea.francia\@inv.uam.es](mailto:lea.francia@inv.uam.es){.email}
-   Phone number: (+34) 914974601

Original version:

-   Name: Joan Domènech Abella
-   e-mail: [joan.domenech\@sjd.es](mailto:joan.domenech@sjd.es)){.email}

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
