---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'Anthropometric measures'" 
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib
csl: "`r ecs.data::get_proj_csl_file()`"
nocite: '@worldhealthorganization, @cdc2022'
editor_options: 
  markdown:
    wrap: 80 ; 
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"

options(width = 88) 

knitr::opts_chunk$set( 
  # engine.path = list(stata = STATA_PATH),
  echo        = FALSE,
  results     = 'hide',
  cleanlog    = TRUE,  
  collectcode = TRUE,  
  comment     = ''     
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

+----------------+-------------------------+----------------+----------------+
| Main variable  | Label                   | Format         | Code           |
| name           |                         |                |                |
+:===============+:========================+:===============+:===============+
| height_s3      | Height, in cms          | Continuous     | 50 - 200       |
|                | (c2011_w4)              |                |                |
+----------------+-------------------------+----------------+----------------+
| weight_s3      | Weight, in kgs          | Continuous     | 20 - 200       |
|                | (c2011_w4)              |                |                |
+----------------+-------------------------+----------------+----------------+
| waist_s3       | Waist circumference, in | Continuous     | 40 - 300       |
|                | cms (c2011_w4)          |                |                |
+----------------+-------------------------+----------------+----------------+
| bmi_s3         | Body mass index,        | Continuous     | 14.86 - 47.60  |
|                | continuous (c2011_w4)   |                |                |
+----------------+-------------------------+----------------+----------------+
| bmi_cat_s3     | Body mass index,        | Categorical    | 1 =            |
|                | categorical (c2011_w4)  |                | Underweight;   |
|                |                         |                |                |
|                |                         |                | 2 = Normal     |
|                |                         |                | weight;        |
|                |                         |                |                |
|                |                         |                | 3 =            |
|                |                         |                | Overweight;    |
|                |                         |                |                |
|                |                         |                | 4 = Obese      |
+----------------+-------------------------+----------------+----------------+

# Variable description

-   **Height** refers to the height in centimeters.

-   **Weight** refers to the weight in kilograms.

-   **Waist circumference** refers to waist circumference in cms.

-   **BMI**, body mass index, provides a fairly reliable indictor of body
    fatness for most people and is used to screen for weight categories that may
    lead to health problems. BMI does not measure body fat directly, but
    research has shown that BMI correlates to direct measures of body fat, such
    as underwater weighing and dual energy x-ray absorptiometry (DXA).

# Explanation of how the variables were created

-   **Height** is renamed from `q2506_height_s3` to `height_s3`. Is the height
    of the participant (cms). It is obtained by using a measuring tape. 
    Constraints are set to allow only valid values.

-   **Weight** is renamed from `q2507_weight` to `weight_s3`. Is the weight of 
    the participant in kilograms. It is obtained using a weight scale without
    shoes. Constraints are set to allow only valid values.

-   **Waist circumference** is renamed from `q2508_waist_s3` to `waist_s3`. It
    is obtained by identifying the midpoint between the last rib and the illiac
    crest, ensuring that the tape measure around the body is parallel to the
    ground. It is intended that the measurement is taken directly from the skin,
    if not, with light clothing. Constraints are set to allow only valid values.

-   **BMI**: Defined as the weight in kilograms divided by the square of
    the height in meters (kg/m2) [@worldhealthorganization; @cdc2022]). Four
    categories are created using the WHO BMI cut-off points:

+--------------------+---------------------------+----------------------------+
| Classification of  |                           |                            |
| BMI                |                           |                            |
+====================+===========================+============================+
|                    | **Principal cut-off       | **Additional cut-off       |
|                    | points**                  | points**                   |
+--------------------+---------------------------+----------------------------+
| **Underweight**    | **\<18.50**               | \<18.50                    |
+--------------------+---------------------------+----------------------------+
| Severe thinness    | \<16.00                   | \<16.00                    |
+--------------------+---------------------------+----------------------------+
| Moderate thinness  | 16.00 - 16.99             | 16.00 - 16.99              |
+--------------------+---------------------------+----------------------------+
| Mild thiness       | 17.00 - 18.49             | 17.00 - 18.49              |
+--------------------+---------------------------+----------------------------+
| **Normal range**   | 18.50 - 24.99             | 18.50 - 22.99              |
+--------------------+---------------------------+----------------------------+
| Normal range       | 18.50 - 24.99             | 23.00 - 24.99              |
+--------------------+---------------------------+----------------------------+
| **Overweight**     | \>= 25.00                 | \>= 25.00                  |
+--------------------+---------------------------+----------------------------+
| Pre-obese          | 25.00 - 29.99             | 27.50 - 29.99              |
+--------------------+---------------------------+----------------------------+
| **Obese**          | \>= 30.00                 | \>= 30.00                  |
+--------------------+---------------------------+----------------------------+
| Obese class I      | 30.00 - 34.99             | 30.00 - 32.49 / 32.50 -    |
|                    |                           | 34.99                      |
+--------------------+---------------------------+----------------------------+
| Obese class II     | 35.00 - 39.99             | 35.00 - 37.49 / 37.50 -    |
|                    |                           | 39.99                      |
+--------------------+---------------------------+----------------------------+
| Obese class III    | \>= 40.00                 | \>= 40.00                  |
+--------------------+---------------------------+----------------------------+

: BMI cut-offs classification

# Syntax used to create the variable (Stata 15)

```{stata load-data}

* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"

```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
recode q2506_height_s3 (997=.)(998=.), gen (height_s3)
label variable height_s3 "Height in cms (c2011w4)"

recode q2507_weight_s3 (997=.)(998=.), gen (weight_s3)
label variable weight_s3 "Weight in kgs (c2011w4)"

recode q2508_waist_s3 (997=.)(998=.), gen (waist_s3)
label variable waist_s3 "Waist circumference in cms (c2011w4)"

gen height=.
replace height=height_s3/100 if height_s3>=120 & height_s3<=200

gen bmi_s3=.
replace bmi_s3= weight_s3/(height*height) if weight_s3>=27.3 & weight_s3<=200
label variable bmi_s3 "Body mass index, continuous (c2011w4)"

gen bmi_cat_s3=.
	replace bmi_cat_s3=1 if bmi_s3<18.5
	replace bmi_cat_s3=0 if bmi_s3>=18.5 & bmi_s3<25
	replace bmi_cat_s3=2 if bmi_s3>=25 & bmi_s3<30
	replace bmi_cat_s3=3 if bmi_s3>=30 & bmi_s3<50
	
label define obese1 0 "Normal weight" 1 "Underweight" 2 "Overweight" 3 "Obese"
label values bmi_cat_s3 obese1
label variable bmi_cat_s3 "Body mass index, categorical (c2011w4)"

```

<br>

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result proxy height_s3 weight_s3 waist_s3 bmi_s3 bmi_cat_s3

codebook height_s3 weight_s3 waist_s3 bmi_s3 bmi_cat_s3
```

```{stata save-data-test, eval=TEST}
saveold "Outcome_anthropometric_measures.dta", replace version(13)
```

```{stata save-data, eval=!TEST}

* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_anthropometric_measures.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Cristina Rodríguez Prada *(adapted the algorithm for wave 4)*
-   e-mail:
    [cristina.rodriguez\@inv.uam.es](mailto:cristina.rodriguez@inv.uam.es){.email}
-   Institution: Universidad Autónoma de Madrid / FIB-HU La Princesa
-   Phone number: (+34) 914975492 <br>

Original version:

-   Name: Beatriz Olaya
-   e-mail: [beatriz.olaya\@pssjd.org](mailto:beatriz.olaya@pssjd.org){.email}
-   Phone number: (+34) 936406350 ext (1) 2540

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
