---
title: "Edad con Salud -- cohort 2011 -- wave 4-- 'social network'" 
output:
  word_document:
   reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib 
csl: "`r ecs.data::get_proj_csl_file()`"
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"

# Opciones de configuración de las salidas de R y Stata:

options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set(
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

| Main variable name | Label                                                                  | Format     | Code |
|:-----------------|:--------------------------|:-----------------|:-----------------|
| size_s3            | Size of the social network (c2011_w4)                                  | Continuous | 0-99 |
| quality_s3         | Quality of contact with the members of the social network (c2011_w4)   | Continuous | 0-8  |
| frequency_s3       | Frequency of contact with the members of the social network (c2011_w4) | Continuous | 0-8  |

# Variable description

The structural dimension of the Berkman-Syme Social Network Index considers the
number of the social ties, frequency and quality (closeness) of contact with the
members of the network as components of the social network [@berkman1979].

# Explanation of how the variables were created

The variable related to the size of the social network (**size3**) corresponds
to the item q6101_s3 (how many people composed the social network). This
variable ranges from 0 to 99, with the datum indicating the number of social
ties.

The variable related to the quality of the social network (**quality3**) was
generated from informations about close relationship with husband/wife (item
q6102_s3), parents (q6111_s3), sons (q6112_s3), grandchildren (q6113_s3), other
relatives (q6114_s3), work colleagues (q6115_s3), friends (q6116_s3), and
neighbour (q6117_s3). The original variables were dichotomized into 0 (whenever
the respondent considered his/her relationship with the corresponding
person/group as *not very close,* *not close at all*, or when he/she informed
not having a close relationship with any member of the corresponding group), or
1 (in the rest of cases). The final score is the summation of these dichotomized
variables, thus ranging from 0 (lower quality of contact with the members of the
social network) to 9 (higher quality).

The variable about the frequency of social contact with the social network
(**frequency3**) was generated from the items q6211_s1 , q6212_s1, q6213_s1,
q6214_s1, q6215_s1, q6216_s1, q6217_s1, q6218_s1, q6219_s1. Those items ask
information about the frecuency of the social contact in the last 12 months,
respectively, with wife/husband, parents, sons, grandchildren, other relatives,
friends, neighbours, and other people. The original variables were dichotomized
into 0 (whenever the respondent informed having contact with the corresponding
person/group either never or once a year), or 1 (in the rest of cases). The
final score is the summation of these dichotomized variables, thus ranging from
0 (lower frequency of contact with the members of the social network) to 1
(higher frequency).

# Syntax used to create the variable (Stata 15)

```{stata load-data}

* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"

```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
* FREQUENCY

generate frequency_spouse = q6211_spouse_s3
recode frequency_spouse (1/2=0)(3/5=1)(999=.)
replace frequency_spouse=0 if q6010a_spouse_s3==2

generate frequency_parents = q6212_parents_s3
recode frequency_parents (1/2=0)(3/5=1)(999=.)
replace frequency_parents=0 if q6010b_parents_s3==2

generate frequency_children = q6213_children_s3
recode frequency_children (1/2=0)(3/5=1)(999=.)
replace frequency_children=0 if q6010c_children_s3 ==2

generate frequency_grandchildren = q6214_grandch_s3
recode frequency_grandchildren (1/2=0)(3/5=1)(999=.)
replace frequency_grandchildren=0 if q6010d_grandch_s3==2

generate frequency_relatives = q6215_other_s3
recode frequency_relatives (1/2=0)(3/5=1)(999=.)
replace frequency_relatives=0 if q6010e_other_s3==2

generate frequency_coworkers = q6216_cowork_s3 
recode frequency_coworkers (1/2=0)(3/5=1)(999=.)
replace frequency_coworkers=0 if q6010f_cowork_s3==2

generate frequency_friends = q6217_friends_s3 
recode frequency_friends (1/2=0)(3/5=1)(999=.)
replace frequency_friends=0 if q6010g_friends_s3==2

generate frequency_neighbours = q6218_neighbours_s3 
recode frequency_neighbours (1/2=0)(3/5=1)(999=.)
replace frequency_neighbours=0 if q6010h_neighbours_s3==2

gen frequency_s3 = frequency_spouse + frequency_parents + frequency_children + frequency_grandchildren + frequency_relatives + frequency_coworkers + frequency_friends + frequency_neighbours

drop frequency_spouse frequency_parents frequency_children frequency_grandchildren frequency_relatives frequency_coworkers frequency_friends frequency_neighbours

label variable frequency_s3 "Frequency of contact with the members of the social network (c2011_w4)"


* SIZE

gen size_s3 = q6101_close_s3
label variable size_s3 "Size of the social network (c2011_w4)"


* QUALITY

gen quality_a = q6102_spouse_s3
recode quality_a (1/2=1) (3/4=0) 
replace quality_a = 0 if q6010a_spouse_s3 == 2

gen quality_b = q6111_parents_s3
gen quality_c = q6112_children_s3
gen quality_d = q6113_grandch_s3
gen quality_e = q6114_other_s3
gen quality_f = q6115_cowork_s3
gen quality_g = q6116_friends_s3
gen quality_h = q6117_neighbours_s3

recode quality_b-quality_h (1=0) (2/3=1)

replace quality_b = 0 if q6010b_parents_s3 == 2
replace quality_c = 0 if q6010c_children_s3 == 2
replace quality_d = 0 if q6010d_grandch_s3 == 2
replace quality_e = 0 if q6010e_other_s3 == 2
replace quality_f = 0 if q6010f_cowork_s3 == 2
replace quality_g = 0 if q6010g_friends_s3 == 2
replace quality_h = 0 if q6010h_neighbours_s3 == 2

gen quality_s3 = quality_a + quality_b + quality_c + quality_d + quality_e + quality_f + quality_g + quality_h

drop quality_a quality_b quality_c quality_d quality_e quality_f quality_g quality_h

label variable quality_s3 "Quality of contact with the members of the social network (c2011_w4)"

```

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}

keep ID_ECS q0007a_result proxy size_s3 frequency_s3 quality_s3

codebook size_s3 frequency_s3 quality_s3
```

```{stata save-data-test, eval=TEST}
saveold "Outcome_socialnetwork.dta", replace version(13)
```

```{stata save-data, eval=!TEST}
* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_socialnetwork.dta", replace version(13)

```

<br>

# Contact person:

-   Name: Lea Francia
-   e-mail: [lea.francia\@inv.uam.es](mailto:lea.francia@inv.uam.es){.email}
-   Phone number: (+34) 914974601

Original version:

-   Name: Dario Moreno Agostino
-   e-mail: [d.moreno\@ucl.ac.uk](mailto:d.moreno@ucl.ac.uk){.email}

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
