---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'Feelings of guilt (for being a burden)'" 
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: "`r ecs.data::get_proj_bib_file()`" 
csl: "`r ecs.data::get_proj_csl_file()`"
nocite: '@pedroso-chaparro_validation_2021, @cousineau_measuring_2003' 
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"

options(width = 88) 
knitr::opts_chunk$set( 
  echo        = FALSE, 
  results     = 'hide',
  cleanlog    = TRUE,  
  collectcode = TRUE,  
  comment     = ''     
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

+--------------------+------------------------------------------+------------+----------------------+
| Main variable name | Label                                    | Format     | Code                 |
+:===================+:=========================================+:===========+:=====================+
| guilt_burden_s3    | Feelings of guilt for being a burden in  | Continuous | 4-18                 |
|                    | direct scores (c2011_w4)                 |            |                      |
+--------------------+------------------------------------------+------------+----------------------+
| guilt_burden_100_s3| Feelings of guilt for being a burden     | Continuous | 0-100                |
|                    | (transformed to 0-100) (c2011_w4)        |            |                      |
+--------------------+------------------------------------------+------------+----------------------+
| burden_to_family_s3| Self-perception of being a burden        | Continuous | 1-5                  |
|                    | to the family (c2011_w4)                 |            |                      |
+--------------------+------------------------------------------+------------+----------------------+


# Variable description

The feeling of guilt about being a burden to the caregiver is one of the main 
discomforts experienced by the elderly. It is one of the most common 
health-related concerns as they become dependent and take a toll on the 
resources of their family members. 

The original scale proposed by Cousineau et al. (2003) "Self-Perceived Burden
Scale" assesses the feeling of guilt for perceiving oneself as a burden to the 
family in older adults without any functional or cognitive impairment.  
It consists of 25 Likert-type scale items, from 1 ("never or almost never") 
to 5 ("almost always"). 

The version adapted by Pedroso-Chaparro et al. (2021) uses an abbreviated 
version of 16 items. In Wave 4 in Age with Health a reduced version of this 
abbreviated version of only 5 items has been used.

Mainly two facets are distinguished:
- __Guilt associated with the perception of being a burden.__ In Edad con Salud
these items would be composed of items `q6724_guilt1_s3` (I feel guilty for 
overburdening my family for taking care of me); 
`q6727_extra_s3` (I feel good receiving care from my family); 
`q6725_guilt2_s3` (I feel guilty because my family has lost control of their
life in order to take care of me); `q6726_guilt3_s3` (I feel guilty because my 
family is helping me beyond their capacity). 

- __Self-perception of being a burden to the family.__ Item `q6728_extra_s3` 
(I think I make things easy for my family). 

It is assumed that there is a relationship between both variables, 
since they would be part of the same phenomenon. 

# Explanation of how the variables were created

- __Guilt associated with the perception of being a burden.__ In Edad con Salud
these items would be composed of items `q6724_guilt1_s3` (I feel guilty for 
overburdening my family for taking care of me); 
`q6727_extra_s3` (I feel good receiving care from my family); 
`q6725_guilt2_s3` (I feel guilty because my family has lost control of their
life in order to take care of me); `q6726_guilt3_s3` (I feel guilty because my 
family is helping me beyond their capacity). 
  - __guilt_burden_s3__: Sum of the aforementioned items. Item `q6727_extra_s3`
    is an inverse item, so it is necessary to recode the values for a correct
    interpretation of the score. This is a direct score.
  - __guilt_burden_100_s3__: Same as above, but in 0 to 100 scale. This 
  transformation is done following the formula `newvar = ((var - min)/(max-min))*100`,
  which first subtracts the minimum value of the original range (4) from each 
  value of var. It then divides the result by the range of the original variable
  (18-4=14) to get a proportion between 0 and 1. Finally, it multiplies this 
  proportion by the range of the new variable (100) to get the rescaled value.

- __Self-perception of being a burden to the family.__ Item `q6728_extra_s3` 
(I think I make things easy for my family). 
   - __burden_to_family_s3__: Direct score of the item, which has to be reversed
    for score to maintain its correct interpretation.

Rising scores indicating increasing burden. 

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
* Guilt associated with the perception of being a burden
recode q6724_guilt1_s3 q6727_extra_s3 q6725_guilt2_s3 q6726_guilt3_s3 (888=.) (999=.)

recode q6727_extra_s3 (1 = 5) (2 = 4) (3 = 3) (4 = 2) (5 = 1)

gen guilt_burden_s3 = (q6724_guilt1_s3 + q6727_extra_s3 + q6725_guilt2_s3 + q6726_guilt3_s3)
label variable guilt_burden_s3 "Feelings of guilt for being a burden in direct scores (c2011_w4)"

gen guilt_burden_100_s3 = ((guilt_burden_s3-4) / (18-4))*100
label variable guilt_burden_100_s3 "Feelings of guilt for being a burden (transformed to 0-100) (c2011_w4)"

*Self-perception of being a burden to the family
recode q6728_extra_s3 (1 = 5) (2 = 4) (3 = 3) (4 = 2) (5 = 1) (888=.) (999=.)
gen burden_to_family_s3 = q6728_extra_s3
label variable burden_to_family_s3 "Self-perception of being a burden to the family (c2011_w4)"
```

<br> 

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result_s3 proxy_s3 guilt_burden_s3 guilt_burden_100_s3 burden_to_family_s3

codebook guilt_burden_s3 guilt_burden_100_s3 burden_to_family_s3
```


```{stata save-data-test, eval=TEST}
saveold "Outcome_burdenguilt.dta", replace version(13)
```


```{stata save-data, eval=!TEST}
* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_burdenguilt.dta", replace version(13)

```

<br>

# Contact person:

-   Name: Cristina Rodríguez Prada
-   e-mail:
    [cristina.rodriguez@inv.uam.es](cristina.rodriguez@inv.uam.es){.email}

<br>


# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
