---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'TOBACCO CONSUMPTION'" 
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
# bibliography: "`r ecs.data::get_proj_bib_file()`" ## TODO: Borrar esta línea, O BIEN deescomentar esta línea y borrar siguiente UNA VEZ completo el documento
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib ## TODO: Descomentar esta línea SI se utiliza el gestor bibliográfico, en otro caso BORRAR esta línea
csl: "`r ecs.data::get_proj_csl_file()`"
nocite: '@world_health_organization_tobacco_2023' ## TODO: Descomentar y añadir cita(s) o borrar campo
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

# Ruta al ejecutable de Stata (Ejecuta `sysdir` en Stata para encontrar la ruta)

## TODO: sustituir '<path\to\>' por la ubicación de Stata
##       (ejecutar `sysdir` en Stata) y descomentar línea 62,
##       o eliminar linea 62:
STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"


# Opciones de configuración de las salidas de R y Stata:

options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( # Para todos los 'chunks' de código:
  ## TODO: Descomentar si hace falta
  # engine.path = list(stata = STATA_PATH),
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

+--------------------+---------------------------------------------+-------------+------------------------+
| Main variable name | Label                                       | Format      | Code                   |
+:===================+:============================================+:============+:=======================+
| tobacco_s3         | Tobacco consumption, categorical (c2011_w4) | Ordinal     | 0 = Never smoked;      |
|                    |                                             |             |                        |
|                    |                                             |             | 1 = Daily smoker;      |
|                    |                                             |             |                        |
|                    |                                             |             | 2 = Smoker, not daily; |
|                    |                                             |             |                        |
|                    |                                             |             | 3 = Not current smoker |
+--------------------+---------------------------------------------+-------------+------------------------+

# Variable description

**Tobacco products** are made entirely or partly of leaf tobacco as raw material, which are intended to be smoked, sucked, chewed or snuffed. All contain the highly addictive psychoactive ingredient, nicotine.

**Tobacco use** is one of the main risk factors for a number of chronic diseases, including cancer, lung diseases, and cardiovascular diseases.

# Explanation of how the variables were created

This variable describes the smoking habit of the participants, including the use of smokeless tobacco. Original variables include `q3001_used_tobac_s3` *(Have you used tobaco?* *(1 = Yes, 2 = No))* and `q3002_use_tobaco_s3` *(Do you use tobaco?)* *(1 = Yes, daily; 2 = Yes, not daily; 3 = No; 4 = The participant said that he/* *she has never smoked before)*

Our final variable has four levels, product of the combination of both variables:

-   0 = Never smoked;if `q3001_used_tobac_s3` == 2 and`q3002_use_tobaco_s3` == 1 & 4
-   1 = Daily smoker; if `q3001_used_tobac_s3`==1 & `q3002_use_tobaco_s3`==1     
-   2 = Smoker, not daily; if `q3001_used_tobac_s3`==1 & `q3002_use_tobaco_s3`==2
-   3 = Not current smoker if `q3001_used_tobac_s3`==1 & `q3002_use_tobaco_s3`==3

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
gen tobacco_s3=.
replace tobacco_s3=0 if q3001_used_tobac_s3==2
replace tobacco_s3=0 if q3001_used_tobac_s3==1 & q3002_use_tobaco_s3==4
replace tobacco_s3=1 if q3001_used_tobac_s3==1 & q3002_use_tobaco_s3==1
replace tobacco_s3=2 if q3001_used_tobac_s3==1 & q3002_use_tobaco_s3==2
replace tobacco_s3=3 if q3001_used_tobac_s3==1 & q3002_use_tobaco_s3==3

label define dTobacco 0 "Never smoked" 1 "Daily smoker" 2 "Smoker, not daily" 3 "Not current smoker"
label values tobacco_s3 dTobacco
label variable tobacco_s3 "Tobacco consumption, categorical (c2011_w4)"
```

<br> 

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result proxy tobacco_s3
codebook tobacco_s3
```

```{stata save-data-test, eval=TEST}
saveold "Outcome_tobaccoconsumption.dta", replace version(13)
```

```{stata save-data, eval=!TEST}
* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_<variable>_<qualifier>.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Cristina Rodríguez Prada
-   e-mail: [cristina.rodriguez@inv.uam.es](mailto:cristina.rodriguez@inv.uam.es){.email}

<br>



Original version:

-   Name: Beatriz Olaya
-   e-mail: [beatriz.olaya@pssjd.org](mailto:beatriz.olaya@pssjd.org){.email}

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
