---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'Quality of life (WHOQOL-AGE)'" 
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
# bibliography: "`r ecs.data::get_proj_bib_file()`" ## TODO: Borrar esta línea, O BIEN deescomentar esta línea y borrar siguiente UNA VEZ completo el documento
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib ## TODO: Descomentar esta línea SI se utiliza el gestor bibliográfico, en otro caso BORRAR esta línea
csl: "`r ecs.data::get_proj_csl_file()`"
nocite: '@caballero_validation_2013' ## TODO: Descomentar y añadir cita(s) o borrar campo
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
library(Statamarkdown)
STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"
options(width = 88) 
knitr::opts_chunk$set( 
  echo        = FALSE, 
  results     = 'hide',
  cleanlog    = TRUE,  
  collectcode = TRUE,  
  comment     = ''     
)
TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

| Main variable name | Label                      | Format     | Code    |
|:-------------------|:---------------------------|:-----------|:--------|
| WHOQOL_AGE_s3      | Quality of life (c2011_w4) | Continuous | 0 - 100 |

# Variable description

A global score in quality of life was obtained by means of WHOQOL-AGE
(World Health Organization Quality Of Life - AGE), which is composed by 13 
positive/direct items measured on a 1-5 Likert scale. Two factors were 
identified after conducting an Exploratory Factor Analysis. As a previous step, 
a Graded Response Model was run and then scores were obtained for each factor 
based on these models. Finally, the factor scores were transformed into a 0-100 
scale and the WHOQOL-AGE global score was obtained as the average in the two
factor scores.

# Explanation of how the variables were created

**Quality of Life** was assessed by means of 13 items: `q7001`, `q7002`, 
`q7003`, `q7004`, `q7005`, `q7006`, `q7007`, `q7008`, `q7009a`, `q7009b`, 
`q7009c`, `q7009d`, and `q7010b.`.

Details of the validation process were explained in @caballero_validation_2013.

Two factors were obtained, F1 and F2, and the following formula to calculate
factor scores was used:

-   `F1=  q7001/2+q7002+(q7003+q7005)/2+q7004+q7006+q7007+q7008`

-   `F2=  q7001/2+(q7009a+q7009b+q7009c)/3+q7009d+q7010b`

Factor scores were transformed into a 0-100 scale and their sum was divided by
two in order to obtain the global score in quality of life.

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
recode q7001_life_s3 q7002_senses_s3 q7003_health_s3 q7004_yourself_s3 q7005_daily_s3 q7006_relation_s3 q7007_home_s3 q7008_time_s3 q7009a_energy_s3 q7009b_control_s3 q7009c_achieve_s3 q7009d_money_s3 q7010b_intimate_s3 (888=.) (999=.)

egen missing_values_s3 = rowmiss(q7001_life_s3 q7002_senses_s3 q7003_health_s3 q7004_yourself_s3 q7005_daily_s3 q7006_relation_s3 q7007_home_s3 q7008_time_s3 q7009a_energy_s3 q7009b_control_s3 q7009c_achieve_s3 q7009d_money_s3 q7010b_intimate_s3)

generate qol_factor1_s3=(q7001_life_s3/2)+q7002_senses_s3+((q7003_health_s3+q7005_daily_s3)/2)+q7004_yourself_s3+q7006_relation_s3+q7007_home_s3+q7008_time_s3
replace qol_factor1_s3=((qol_factor1_s3-6.5)/26)*100

generate qol_factor2_s3=(q7001_life_s3/2)+((q7009a_energy_s3+q7009b_control_s3+q7009c_achieve_s3)/3)+q7009d_money_s3+q7010b_intimate_s3
replace qol_factor2_s3=((qol_factor2_s3-3.5)/14)*100

replace qol_factor1_s3=(((q7002_senses_s3+((q7003_health_s3+q7005_daily_s3)/2)+q7004_yourself_s3+q7006_relation_s3+q7007_home_s3+q7008_time_s3)-6)/24)*100 if (missing_values_s3==1)&(missing(q7001_life_s3))

replace qol_factor1_s3=((((q7001_life_s3/2)+((q7003_health_s3+q7005_daily_s3)/2)+q7004_yourself_s3+q7006_relation_s3+q7007_home_s3+q7008_time_s3)-5.5)/22)*100 if (missing_values_s3==1)&(missing(q7002_senses_s3))

replace qol_factor1_s3=((((q7001_life_s3/2)+q7002_senses_s3+q7005_daily_s3+q7004_yourself_s3+q7006_relation_s3+q7007_home_s3+q7008_time_s3)-6.5)/26)*100 if (missing_values_s3==1)&(missing(q7003_health_s3))

replace qol_factor1_s3=((((q7001_life_s3/2)+q7002_senses_s3+((q7003_health_s3+q7005_daily_s3)/2)+q7006_relation_s3+q7007_home_s3+q7008_time_s3)-5.5)/22)*100 if (missing_values_s3==1)&(missing(q7004_yourself_s3))

replace qol_factor1_s3=((((q7001_life_s3/2)+q7002_senses_s3+q7003_health_s3+q7004_yourself_s3+q7006_relation_s3+q7007_home_s3+q7008_time_s3)-6.5)/26)*100 if (missing_values_s3==1)&(missing(q7005_daily_s3))

replace qol_factor1_s3=((((q7001_life_s3/2)+q7002_senses_s3+((q7003_health_s3+q7005_daily_s3)/2)+q7004_yourself_s3+q7007_home_s3+q7008_time_s3)-5.5)/22)*100 if (missing_values_s3==1)&(missing(q7006_relation_s3))

replace qol_factor1_s3=((((q7001_life_s3/2)+q7002_senses_s3+((q7003_health_s3+q7005_daily_s3)/2)+q7004_yourself_s3+q7006_relation_s3+q7008_time_s3)-5.5)/22)*100 if (missing_values_s3==1)&(missing(q7007_home_s3))

replace qol_factor1_s3=((((q7001_life_s3/2)+q7002_senses_s3+((q7003_health_s3+q7005_daily_s3)/2)+q7004_yourself_s3+q7006_relation_s3+q7007_home_s3)-5.5)/22)*100 if (missing_values_s3==1)&(missing(q7008_time_s3))

replace qol_factor2_s3=(((((q7009a_energy_s3+q7009b_control_s3+q7009c_achieve_s3)/3)+q7009d_money_s3+q7010b_intimate_s3)-3)/12)*100 if (missing_values_s3==1)&(missing(q7001_life_s3))

replace qol_factor2_s3=((((q7001_life_s3/2)+((q7009b_control_s3+q7009c_achieve_s3)/2)+q7009d_money_s3+q7010b_intimate_s3)-3.5)/14)*100 if (missing_values_s3==1)&(missing(q7009a_energy_s3))

replace qol_factor2_s3=((((q7001_life_s3/2)+((q7009a_energy_s3+q7009c_achieve_s3)/2)+q7009d_money_s3+q7010b_intimate_s3)-3.5)/14)*100 if (missing_values_s3==1)&(missing(q7009b_control_s3))

replace qol_factor2_s3=((((q7001_life_s3/2)+((q7009a_energy_s3+q7009b_control_s3)/2)+q7009d_money_s3+q7010b_intimate_s3)-3.5)/14)*100 if (missing_values_s3==1)&(missing(q7009c_achieve_s3))

replace qol_factor2_s3=((((q7001_life_s3/2)+((q7009a_energy_s3+q7009b_control_s3+q7009c_achieve_s3)/3)+q7010b_intimate_s3)-2.5)/10)*100 if (missing_values_s3==1)&(missing(q7009d_money_s3))

replace qol_factor2_s3=((((q7001_life_s3/2)+((q7009a_energy_s3+q7009b_control_s3+q7009c_achieve_s3)/3)+q7009d_money_s3)-2.5)/10)*100 if (missing_values_s3==1)&(missing(q7010b_intimate_s3))

replace qol_factor1_s3=. if missing_values_s3>1
replace qol_factor2_s3=. if missing_values_s3>1
generate WHOQOL_AGE_s3=(qol_factor1_s3+qol_factor2_s3)/2
label variable WHOQOL_AGE_s3 "Quality of life (c2011_w4)"

```

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result_s3 proxy_s3 WHOQOL_AGE_s3
codebook WHOQOL_AGE_s3
```

```{stata save-data-test, eval=TEST}
saveold "Outcome_WHOQOL.dta", replace version(13)
```

```{stata save-data, eval=!TEST}
* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_WHOQOL.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Cristina Rodríguez Prada *(adapted algorithm for wave 4)*
-   e-mail: [cristina.rodriguez\@inv.uam.es](cristina.rodriguez@inv.uam.es){.email}

<br>

<!--# TODO: Completar / eliminar "original version" según corresponda -->

Original version:

-   Name: Elvira Lara *(adapted algorithm for wave 3)*
-   e-mail: [elvilara\@ucm.es](elvilara@ucm.es){.email}

<br>

-   Name: Marta Miret
-   e-mail: [marta.miret\@uam.es](marta.miret@uam.es){.email}

<br>

-   Name: Félix Caballero
-   e-mail: [mailto:felix.caballero\@uam.es](mailto:felix.caballero@uam.es){.email}

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
