---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'mediterranean diet adherence'" 
output:
  word_document:
   reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib 
csl: "`r ecs.data::get_proj_csl_file()`"
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"


options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( 
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

| Main variable name | Label                                         | Format     | Code |
|:------------------|:------------------------|:------------------|:------------------|
| medas_s3           | Mediterranean Diet Adherence score (c2011_w4) | Continuous | 0-14 |

# Variable description

Mediterranean Diet (MD) is based on high consumption of fruits, vegetables,
cereals, pulses, nuts and seeds. In this diet, there is a moderate consumption
of dairy products, fish, poultry, and eggs and little use of red meat.
Additionally, there is usually a low to moderate amount of wine. Olive oil is
the main cooking and dressing oil. This dietary pattern has been shown to be
beneficial to cardiac functions. The Mediterranean Adherence Screener (MEDAS) is
a 14-item screening questionnaire with questions on food intake and cooking and
eating habits considered characteristic of a Mediterranean dietary pattern
[@martínez-gonzález2012]. MEDAS score ranges from 0 to 14, with higher scores
indicating higher adherence to MD.

# Explanation of how the variables were created

Original variables include informations about: -fruits consumption (q3012_s3);
-vegetables consumption (q3013_s3); -consumption of olive oil (q3033_s3 and
q3034_s3); -consumption of red meat, hamburger, or meat products per day
(q3035_s3); -consumption of butter, margarine, or cream per day (q3036_s3);
-consumption of sweet or carbonated beverages per day (q3037_s3); -consumption
of wine per week (q3038_s3); -consumption of legumes per week (q3039_s3);
-consumption of fish or shellfish per week (q3040_s3); -consumption of
commercial sweets or pastries per week (q3041_s3); -consumption of nuts per week
(q3042_s3); -preferential consumption of chicken, turkey, or rabbit meat instead
of veal, pork, hamburger, or sausage (q3043_s3) -consumption of sofrito per week
(q3044_s3).

Fruit and vegetable consumption items were dichotomized following the previously
stated scoring criteria. The MEDAS score was calculated as the sum of the 14
dichotomized items above. For those participants that chose the "Not applicable"
option in the q3043_poultry variable, a correction formula ([SCORE / 13] \* 14)
was applied to account for their adherence despite the absence of meat
consumption.

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"

```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
*# "Value labels (defined for all the outcome variables in the dataset)":
label define yesno    0 "No" 1 "Yes"

***#"VARIABLE `Mediterranean Diet Adherance total score`"
recode q3013_s3 q3012_s3 (888 =.) 
recode q3043_s3 (999=.)
recode q3013_s3 (min/1.9=0) (2/50=1)
recode q3012_s3 (min/2.9=0) (3/50=1)

egen medas_s3 = rowtotal (q3033_s3 q3034_s3 q3013_s3 q3012_s3 q3035_s3 q3036_s3 q3037_s3 q3038_s3 q3039_s3 q3040_s3 q3041_s3 q3042_s3 q3043_s3 q3044_s3), missing
replace medas_s3 = medas_s3 / 13 * 14 if q3043_s3 == 999
label variable medas_s3 "Mediterranean Diet Adherence score (c2011_w4)"
```

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result proxy medas_s3

codebook medas_s3
```

```{stata save-data-test, eval=TEST}
saveold "Outcome_mediterraneandiet.dta", replace version(13)
```

```{stata save-data, eval=!TEST}

* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_mediterraneandiet.dta", replace version(13)
```

<br>

# Contact person:

Contact:

-   Name: Lea Francia
-   e-mail: [lea.francia\@inv.uam.es](mailto:lea.francia@inv.uam.es){.email}
-   Phone number: (+34) 914974601

Original version:

-   Name: Dario Moreno Agostino
-   e-mail: [d.moreno\@ucl.ac.uk](mailto:d.moreno@ucl.ac.uk){.email}

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
