---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'Wellbeing'" 
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
# bibliography: "`r ecs.data::get_proj_bib_file()`" ## TODO: Borrar esta línea, O BIEN deescomentar esta línea y borrar siguiente UNA VEZ completo el documento
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib ## TODO: Descomentar esta línea SI se utiliza el gestor bibliográfico, en otro caso BORRAR esta línea
csl: "`r ecs.data::get_proj_csl_file()`"
nocite: '@<cita_1>, @<cita_2>' ## TODO: Descomentar y añadir cita(s) o borrar campo
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"

options(width = 88) 
knitr::opts_chunk$set( 
  echo        = FALSE, 
  results     = 'hide',
  cleanlog    = TRUE,  
  collectcode = TRUE,  
  comment     = ''     
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

+--------------------+------------------------------------------+------------+----------------------+
| Main variable name | Label                                    | Format     | Code                 |
+:===================+:=========================================+:===========+:=====================+
| cantril_present_s3 | Evaluative wellbeing. Satisfaction with  |            |                      |
|                    | life (present) (c2011_w4)                |Quantitative| 0-10                 |
+--------------------+------------------------------------------+------------+----------------------+
| cantril_future_s3  | Evaluative wellbeing. Satisfaction with  |            |                      |
|                    | life (future, in 5 years) (c2011_w4)     |Quantitative| 0-10                 |
+--------------------+------------------------------------------+------------+----------------------+
|flourishing_scale_s3| Eudaimonic wellbeing. Flourishing Scale  |Quantitative| 8 (lowest wellbeing)-|
|                    | Score (c2011_w4)                         |            |56 (highest wellbeing)|
+--------------------+------------------------------------------+------------+----------------------+
| positiveaffect_s3  | Experiential wellbeing. Positive affect  |Quantitative| 0-6                  |
|                    | (DRM) (c2011w4)                          |            |                      |                  
+--------------------+------------------------------------------+------------+----------------------+
| negativeaffect_s3  | Experiential wellbeing. Negative affect  |Quantitative| 0-6                  |
|                    | (DRM) (c2011w4)                          |            |                      |  
+--------------------+------------------------------------------+------------+----------------------+
| netaffect_s3       | Experiential wellbeing. Net affect       |Quantitative| -6 - +6              |
|                    | (DRM) (c2011w4)                          |            |                      | 
+--------------------+------------------------------------------+------------+----------------------+

# Variable description

__Subjective wellbeing__ is a psychological phenomenon that can be divided into
three dimensions:

- __Evaluative wellbeing__ _(Cantril Self-Anchoring Striving Scale,_ @cantril_pattern_1966)
(`cantril_present_s3` and `cantril_future_s3`). In words
of Cantril, “A person is asked to define on the basis of his own assumptions,
perceptions, goals and values the two extremes or anchoring points of the 
spectrum on which some scale measurement is desired – for example he may
be asked to define the ‘top’ and ‘bottom,’ the ‘good’ and ‘bad,’ the ‘best’ 
and the ‘worst.’ This well-defined continuum is then used as our measuring
device”. The Cantril self-anchoring striving scale or in short the 
Cantril-ladder is an instrument to measure people’s attitudes towards their 
life and its components in various respects (_evaluative wellbeing_). 
The original scale contains a ladder with 11 steps, and the end points of
the scale are defined by the respondents in terms of their best and their worst 
life experience.


- __Eudaimonic wellbeing__ _(Flourishing Scale_, @diener_new_2009) 
(`flourishing_scale_s3`). Several authors, based on earlier humanistic 
psychology theories, suggest that there are several universal human psychological
needs, such as the need for competence, relatedness, and self-acceptance. 
Additionally, other approaches have proposed the importance of interest and 
engagement, and purpose and meaning, as basic components of human wellbeing. 
These aspects of wellbeing have been conceptualized as “eudaimonic wellbeing”. 
The _Flourishing Scale_ allows assessing them in a brief and reliable way. 


- __Experiential wellbeing__. The _Day Reconstruction Method (DRM)_ (@kahneman_survey_2004,
@kahneman_developments_2006) measures how people spend their time and how they
emotionally experience the various activities of life. In the extended version,
it contains 12 items, including 4 positive and 8 negative. An example of a DRM
item follows: \_"Indicate to what extent you\_ *felt happy during specific
activities of your day"*. Response options are conformed by a 7-point Likert
scale ranging from 0,"not at all" to 6, "very much".

In Edad con Salud, respondents were asked to complete an abbreviated version
of the DRM (@miret_validation_2012, @ayuso-mateos_multi-country_2013,
@caballero_evaluation_2014, @mellor-marsa_measurement_2016). Each participant
provided information about the activities conducted the previous day and rated
the emotions experienced during these activities. By means of Computer-Assisted
Personal Interviewing, one activity was randomly selected for each time frame
(morning, afternoon and evening).

-   **Positive affect** (`positiveaffect_s3`) refers to emotions of happiness, 
    warminess, friendliness and enjoyment with myself.
-   **Negative affect** (`negativeaffect_s3`) is defined by emotions of 
    frustration, annoyment, depression, feeling blue, hassled/pushed around, 
    angry/hostile, worried/anxious, criticized/put down.
-   **Net affect** (`netaffect_s3`) is defined as the average of two positive 
    emotions (calm/relaxed and enjoying) minus the average of the five negative
    ones (worried, rushed, irritated/angry, depressed, and tense/stressed), 
    i.e., positive affect minus negative affect.


# Explanation of how the variables were created

Variables of __"evaluative wellbeing"__ (`cantril_present_s3` and `cantril_future_s3`)
are conformed by the independent score of two items of the Cantril 
Self-Anchoring Striving Scale. Original items were`q7008d_cantril_s3` for the 
present life satisfaction, and `cantril_future_s3` for the life satisfaction 
respondants expect to have in 5 years. These variables are not modified, 
just have been recoded their values for missing cases (999 and 888). 
They have been renamed too. Theoretically, the score ranges from 0 to 10, 
highest scores indicating higher evaluative wellbeing.

<br>

From __"eudaimonic wellbeing"__, variable `flourishing_scale_s3` is created 
based on items `q7528_s3:q7535_s3`.
First, we recoded 888 and 999 values as missing values. Second, all variables are
added together. Theoretically, the score ranges from 8 (lowest  wellbeing possible) 
to 56 (highest wellbeing possible).

<br>

In the case of __"experiential wellbeing"__:

Three measures were obtained from the Day Reconstruction Method (DRM): *positive*
*affect*, *negative affect* and *net affect*.

The Day Reconstruction Method records the activities that a person has performed
in the morning (until 12 noon), in the afternoon (12 noon - 6 p.m.) and in the 
evening (6 p.m. - bedtime). In our database, activities are recorded as 
dichotomous variables and three of them are randomly chosen, 
one for each time of day. 

Subsequently, the duration of these activities is asked with the variables
`q7026_s3`, `q7066_s3`and `q7111_s3.` The format is currently in seconds and
is used as a way of weighting the duration of the emotion.

Items `q7029_worr_s3:q7033_tense_s3`,`q7069_worr_s3:q7073_tense_s3`, 
`q7114_worr_s3:q7118_tense_s3` ask about different negative emotions associated
with these tasks: worry, rush, irritation, depresion, and tension. 

Items `q7034_calm_s3 q7035_enjoy_s3`, `q7074_calm_s3 q7075_enjoy_s3` and
`q7119_calm_s3 q7120_enjoy_s3` ask about the positive emotions of calm/relaxation
and enjoying. 

The first step consists of constructing the mean of positive emotions per 
activity and the mean of negative emotions per activity while maintaining 
the missings. 

Then, the mean of negative and positive affect is created by multiplying by 
each time variable for affect in the three activities separately and weighting 
them by the final duration. Since duration is given as a fraction of the time of
day, if the value 1 is exceeded (duration_DRM4 >1), it can be understood that 
the duration estimate is unreliable. Therefore, to avoid losing these values, 
an intermediate solution is chosen: to obtain a score for these participants 
based on the average of the emotions, but not by weighting on the duration of 
the activities.

-   **Positive affect** (`positiveaffect_s3`: the average of the two positive 
    emotions (calm/relaxed and enjoying).

-   **Negative affect** (`negativeaffect_s3`): the average of the five negative 
    ones (worried, rushed, irritated/angry, depressed, and tense/stressed).

-   **Net affect** (`netaffect_s3`): the average of the two positive emotions 
    (calm/relaxed and enjoying) minus the average of the five negative ones 
    (worried, rushed, irritated/angry, depressed, and tense/stressed), i.e., 
    positive affect minus negative affect (Kahneman & Krueger, 2006). 
    Individual net affect was calculated adding up the net-affect in each 
    activity performed weighted by its duration. Net affect scores ranged from
    -6 to 6, with higher scores representing a better affective state.


# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
*EVALUATIVE WELLBEING

recode q7008d_cantril_s3 (999 =.) (888=.), gen(cantril_present_s3)
label variable cantril_present_s3 "Evaluative wellbeing. Satisfaction with life (present) (c2011_w4)"  

recode q7008e_cantril_future_s3 (999 =.) (888=.), gen(cantril_future_s3)
label variable cantril_future_s3 "Evaluative wellbeing. Satisfaction with life (future, in 5 years) (c2011_w4)"  


*EUDAIMONIC WELLBEING
recode q7528_s3 q7529_s3 q7530_s3 q7531_s3 q7532_s3 q7533_s3 q7534_s3 q7535_s3 (888 999 = .)
gen flourishing_scale_s3  = q7528_s3 + q7529_s3 + q7530_s3 + q7531_s3 + q7532_s3 + q7533_s3 + q7534_s3 + q7535_s3
label variable flourishing_scale_s3 "Eudaimonic wellbeing. Flourishing Scale Score (c2011_w4)"


*EXPERIENTIAL WELLBEING
* Duration of the three activities randomly chosen in the DRM at wave 4

gen duration_DRM4=q7026_s3 + q7066_s3 + q7111_s3

* Negative affect per activity
egen neg_activity1=rowmean(q7029_worr_s3-q7033_tense_s3)
egen neg_activity2=rowmean(q7069_worr_s3-q7073_tense_s3)
egen neg_activity3=rowmean(q7114_worr_s3-q7118_tense_s3)

* Positive affect per activity
egen pos_activity1=rowmean(q7034_calm_s3 q7035_enjoy_s3)
egen pos_activity2=rowmean(q7074_calm_s3 q7075_enjoy_s3)
egen pos_activity3=rowmean(q7119_calm_s3 q7120_enjoy_s3)

* Mean of negative and positive affect
gen negativeaffect_s3=((q7026_s3*neg_activity1) + (q7066_s3*neg_activity2) + (q7111_s3*neg_activity3))/duration_DRM4
gen positiveaffect_s3=((q7026_s3*pos_activity1) + (q7066_s3*pos_activity2) + (q7111_s3*pos_activity3))/duration_DRM4
replace negativeaffect_s3=((neg_activity1+neg_activity2+neg_activity3)/3) if duration_DRM4>1
replace positiveaffect_s3=((pos_activity1+pos_activity2+pos_activity3)/3) if duration_DRM4>1

*Net affect
gen netaffect_s3=positiveaffect_s3-negativeaffect_s3

label variable positiveaffect_s3 "Experiential wellbeing. Positive affect (DRM) (c2011w4)"
label variable negativeaffect_s3 "Experiential wellbeing. Negative affect (DRM) (c2011w4)"
label variable netaffect_s3 "Experiential wellbeing. Net affect (DRM) (c2011w4)"

```

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result_s3 proxy_s3 cantril_present_s3  cantril_future_s3  flourishing_scale_s3 positiveaffect_s3 negativeaffect_s3  netaffect_s3   
codebook cantril_present_s3  cantril_future_s3  flourishing_scale_s3 positiveaffect_s3 negativeaffect_s3  netaffect_s3
```

```{stata save-data-test, eval=TEST}
saveold "Outcome_wellbeing.dta", replace version(13)
```

```{stata save-data, eval=!TEST}
* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_wellbeing.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Cristina Rodríguez Prada (adapted algorithms for wave 4 & creates
    originally cantril variables)
-   e-mail:
    [cristina.rodriguez\@inv.uam.es](mailto:cristina.rodriguez@inv.uam.es){.email}

<br>

**Original version (DRM):**

-   Name: Félix Caballero
-   e-mail: [felix.caballero\@uam.es](mailto:felix.caballero@uam.es){.email}
-   Institution: Universidad Autónoma de Madrid
-   Phone number: (+34) 91 497 50 88

<br>

-   Name: Marta Miret
-   e-mail: [marta.miret\@uam.es](mailto:marta.miret@uam.es){.email}
-   Institution: Universidad Autónoma de Madrid
-   Phone number: (+34) 91 497 27 16

<br>

-   Name: Elvira Lara *(adapted the algorithm for wave 3)*
-   e-mail: [elvira.lara\@uam.es](mailto:elvira.lara@uam.es){.email}
-   Institution: Universidad Autónoma de Madrid / Universidad Complutense de Ma
    drid

<br>

**Original version (Flourishing Scale):**

-   Name: Darío Moreno Agostino
-   e-mail: [dario.moreno\@uam.es](mailto:dario.moreno@uam.es)

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
