---
title: "Edad con Salud -- cohort 2011 -- wave 4-- 'income'"
output:
  word_document:
   reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib 
csl: "`r ecs.data::get_proj_csl_file()`"
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"


# Opciones de configuración de las salidas de R y Stata:

options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( 
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

| Main variable name | Label                                                | Format      | Code                                    |
|:------------------|:----------------------|:------------------|:------------------|
| hhincome_s3        | Annual household income level (bracket) (c2011_w4)   | Continuous  | 0 - 1.000.000                           |
| hhquintile_s3      | Quintile of annual household income level (c2011_w4) | Categorical | 1 = First quintile - 5 = Fifth quintile |
| indincome_s3       | Annual working income level (bracket) (c2011_w4)     | Categorical | 0 - 1.000.000                           |
| indquintile_s3     | Quintile of annual working income level (c2011_w4)   | Categorical | 1 = First quintile - 5 = Fifth quintile |

# Variable description

The annual household income level is the participant's estimation of the overall
household level of income during the last 12 months taking into account all
income sources. The annual working income level is the participant's estimation
of his/her working earnings during the last 12 months. Both variables are
reported by means of a category representing a bracket of income in euros. The
"continuous" variables replace the value of the category with the midpoint of
the corresponding bracket. The quintile variables reorder that information into
five groups.

# Explanation of how the variables were created

The annual household income level (hhincome_s3 and indincome_s3) was calculated
using items q0410g and q1516_total giving respectively score through a bracket
and the bracket's midpoint. These two items have been completed by three items
that give information about gross income of the family (q0410i-k) after
excluding "Less than 0€", "Don't know" and "Refuse" categories. The same process
has been used to created the Annual working income level variables. Quintile of
annual household income level and Quintile of annual working income level
variables have also been created using the same items to obtain categorical
variables related to participant's income. In order to assess the level of
income the middle of the bracket has been taking into acccount (e.g. income
between 1.000 and 1.999 == 1.500).

To obtain an amount ranging from 0 to 1.000.000 the category items
(q1516_total_s3 q0410g_total_s3 q0410i_ss_s3 q0410j_government_s3
q0410k_other_s3) ranging from 0 to 35 have been recoded using the middle of each
bracket; (1=0; 2=0; 3=500; 4=1500; 5=2500; 6=3500; 7=4500; 8=5500; 9=6500; 10=7500;
11=8500; 12=9500; 13=10500; 14=11500; 15=12500; 16=13500; 17=14500; 18=15500;
19=16500; 20=17500; 21=18500; 22=19500; 23=22500; 24=27500; 25=32500; 26=37500;
27=42500; 28=47500; 29=62500; 30=87500; 31=125000; 32=175500; 33=250000;
34=400000; 35=750000; 36=1000000)

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"

```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
*# "Value labels (defined for all the outcome variables in the dataset)":
label define quintile  1 "First quintile" 2 "Second quintile" 3 "Third quintile" 4 "Fourth quintile" 5 "Fifth quintile"

*# Mid Brackets of income level
recode q1516_total_s3 q0410g_total_s3 q0410i_ss_s3 q0410j_government_s3 q0410k_other_s3 (1=0) (2=0) (3=500) (4=1500) (5=2500) (6=3500) (7=4500) (8=5500) (9=6500) (10=7500) (11=8500) (12=9500) (13=10500) (14=11500) (15=12500) (16=13500) (17=14500) (18=15500) (19=16500) (20=17500) (21=18500) (22=19500) (23=22500) (24=27500) (25=32500) (26=37500) (27=42500) (28=47500) (29=62500) (30=87500) (31=125000) (32=175500) (33=250000) (34=400000) (35=750000) (36=1000000) (else=.)

*# Annual household income level (bracket)
gen hhincome_s3=q0410g_total_s3
replace hhincome_s3 = q0410i_ss_s3 if q0410i_ss_s3 > hhincome_s3 & q0410i_ss_s3!=.
replace hhincome_s3 = q0410j_government_s3 if q0410j_government_s3 > hhincome_s3 & q0410j_government_s3!=.
replace hhincome_s3 = q0410k_other_s3 if q0410k_other_s3 > hhincome_s3 & q0410k_other_s3!=.
label variable hhincome_s3 "Annual household income level (bracket) (c2011_w4)"

*# Annual working income level (bracket) (c2011_w4) 
gen indincome_s3 = q1516_total_s3
replace indincome_s3 = q0410k_other_s3 if indincome_s3==0
replace indincome_s3 = q0410i_ss_s3 if indincome_s3==0
replace indincome_s3 = q0410j_government_s3 if indincome_s3==0 
label variable indincome_s3 "Annual working income level (bracket) (c2011_w4)"


*# Quintile of annual and working income level
xtile hhquintile_s3 = hhincome_s3, nq(5)
xtile indquintile_s3 = indincome_s3, nq(5)

label variable hhquintile_s3 "Quintile of annual household income level (c2011_w4)"
label variable indquintile_s3 "Quintile of annual working income level (c2011_w4)"
label values   hhquintile_s3 quintile 
label values   indquintile_s3 quintile 

```

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result proxy hhincome_s3 indincome_s3 hhquintile_s3 indquintile_s3

codebook hhincome_s3 indincome_s3 hhquintile_s3 indquintile_s3
```

```{stata save-data-test, eval=TEST}
saveold "Outcome_income.dta", replace version(13)
```

```{stata save-data, eval=!TEST}

* Cohorte 2011, ola 4:
* saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_income.dta", replace version(13)

```

<br>

# Contact person:

-   Name: Lea Francia
-   e-mail: [lea.francia\@inv.uam.es](mailto:lea.francia@inv.uam.es){.email}
-   Phone number: (+34) 914974601

Original version:

-   Name: Dario Moreno Agostino
-   e-mail: [d.moreno\@ucl.ac.uk](mailto:d.moreno@ucl.ac.uk){.email}

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
