---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'Frailty'" ## TODO: Añadir "cohort", "assessment" y "variable name(s)/domain"
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
# bibliography: "`r ecs.data::get_proj_bib_file()`" ## TODO: Borrar esta línea, O BIEN deescomentar esta línea y borrar siguiente UNA VEZ completo el documento
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib ## TODO: Descomentar esta línea SI se utiliza el gestor bibliográfico, en otro caso BORRAR esta línea
csl: "`r ecs.data::get_proj_csl_file()`"
# nocite: '@<cita_1>, @<cita_2>' ## TODO: Descomentar y añadir cita(s) o borrar campo
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}
library(Statamarkdown)
STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"
options(width = 88) 
knitr::opts_chunk$set( 
  echo        = FALSE, 
  results     = 'hide',
  cleanlog    = TRUE,  
  collectcode = TRUE,  
  comment     = ''     
)
TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

+--------------------+------------------------------------------+------------+----------------------+
| Main variable name | Label                                    | Format     | Code                 |
+:===================+:=========================================+:===========+:=====================+
| frailty_index_s3   | Frailty (c2011_w4)                       | Dichotomous| 0 non frail;         |
|                    |                                          |            | 1 frail              |
+--------------------+------------------------------------------+------------+----------------------+
| frailty_score_s3   | Frailty score (c2011_w4)                 |Quantitative| 0-1                  |
+--------------------+------------------------------------------+------------+----------------------+

# Variable description

_Frailty_ is an index that considers the various biopsychosocial factors that can
influence the presence of frailty in the elderly. This index originally includes
a total of 31 ratings, covering the three dimensions assessed: “Chronic Health 
Conditions,” “Cognitive functions” and “Disability and daily-life functioning.” 
Each of these ratings receive a specific score (Tyrovolas et al., 2018). 
Note though that presence of injury was not assessed in this wave, so 30 
out of 31 scales were considered here.

- __Chronic Health Conditions__: Considers the variables Obesity, Hypertension, 
  Stroke, Lung disease, Angina, Asthma, Arthritis, Oral impairment, Diabetes,
  Cataracts, Use of technical aids and Grip Strength.
-	__Cognitive Functions__: Comprises the assessment of _Learning and Short-term_
  _Memory_ (Delayed Recall task), _Learning and Short-term Memory_ (Immediate Recall
  task), _Working memory_ (Forwards and Backwards Digit tasks), and _Verbal fluency_
  (Animal Naming task).
	Verbal Fluency and the Immediate Recall task was assessed using the Consortium
	to Establish a Registry for Alzheimer Disease Word List Memory 
	(Moms et al., 1989).
	The Delayed Recall task and the Forwards and Backwards Digit tasks were based 
	on the WAIS III (Wechsler, 2002).
-	__Disability and daily-life functioning__: Assesses the participant’s 
  functionality based on the WHODAS version 2.0 scale, and adds questions that 
  cover the self-perception of the health status. Items such as Mobility, 
  Interpersonal activities, Activities of domestic life, Sleep, Rest, Hearing, 
  and Pain are considered @wilson2006, @measurin2010.
	Participants were asked to report the level of difficulty they had in 
	performing some activities over the past 30 days using a five-point scale 
	(none = 1, mild = 2, moderate = 3, severe = 4, extreme / can’t do = 5). 
	However, in the case of the scales for Attention, Mobility, and Interpersonal 
	activities, there are different number of items, so the additive scores have 
	ranges 13-65, 5-25 and 6-30, respectively.
	In the case of mobility, items 2026 and 2041 were omitted (Tyrovolas et al., 2018).
	Quintiles of the aforementioned scores were used to classify participants: 
	without difficulty, mild, moderate, severe, and extreme.

<br>

For further details on the frailty index, please see Tyrovolas et al. (2018) 
(section 2.3.1 in p. 139, and Table 1).


# Explanation of how the variables were created

The scores are given according to the following components assessed:

•	**Dichotomous rating**: Assigned to the first 12 items (obesity, hypertension,
  stroke, lung disease, angina, asthma, arthritis, oral impairment, diabetes, 
  cataracts, use of aid machines, grip strength). 
  Integer values 0 or 1 are given depending on whether the participant has (1) 
  or does not have (0) the condition assessed.
  
•	**3-point scale rating**: Given to the four components of cognitive function 
  and the physical activity outcome. Values of 0, 0.5 and, 1 are assigned, where
  0 means “no alterations,” 0.5 implies intermediate function, and 1 a low 
  function for the “cognitive function” components, and for physical activity 0
  means “high,” 0.5 “moderate, and 1 ”low”.
  
•	**4-point scale rating**: Values of 0, 0.33, 0.66 and 1 are given to the
  vision deficit assessment. A higher value represents a greater visual 
  alteration.
  
•	**5-point scale rating**: Used for the daily life activities domain; 
  scores of 0, 0.25, 0.5, 0.75 and 1 are assigned, where the higher the assigned 
  score, the greater the limitation in the activities.
  All the subscales are then averaged, and a threshold of 0.25 is applied; 
  if the total score is below 0.25, the Non-frail category is assigned (0), 
  while a value of 0.25 or higher indicates frailty (1).


# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"
```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
* TODO: Escribir aquí la sintaxis de generación de las variables
```

<br> 

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
* TODO: Sustituir '<variable(s) generada(s)>' por las variables del dataset
* TODO: Añadir `subsample_pre interview_pre` si procede (sólo para Cohorte 2019, Ola 1)
keep ID_ECS q0007a_result proxy <variable(s) generada(s)>

* TODO: Sustituir '<variable(s) generada(s)>' por las variables del dataset
codebook <variable(s) generada(s)>
```


```{stata save-data-test, eval=TEST}
saveold "Outcome_frailty.dta", replace version(13)
```


```{stata save-data, eval=!TEST}
* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_frailty.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Cristina Rodríguez Prada _(adapted algorithms for wave 4)_
-   e-mail:
    [cristina.rodriguez\@inv.uam.es](mailto:cristina.rodriguez@inv.uam.es){.email}

<br>



Original version:

-   Name: Viviana Contreras Pizarro
-   e-mail:
    [vcontrer@ucm.es](mailto:vcontrer@ucm.es){.email}

<br>

-   Name: Daniel Morillo Cuadrado
-   e-mail:
    [daniel.morillo@cibersam.es](mailto:daniel.morillo@cibersam.es){.email}

<br>

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
