---
title: "Edad con Salud -- cohort 2011 -- wave 4 -- 'cognitive reserve'"
output:
  word_document:
   reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: ../../../../www/Edad_con_salud_outcome_vars.bib 
csl: "`r ecs.data::get_proj_csl_file()`"
editor_options: 
  markdown:
    wrap: 80
params:
  test: true
  double-coding: false
---

```{r setup, include=FALSE}

library(Statamarkdown)

options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( 
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)

TEST          <- params$test
DOUBLE_CODING <- params$`double-coding` & TEST
```

# Variable information

| Main variable name | Label                                    | Format    | Code |
|:-------------------|:-----------------------------------------|:----------|:-----|
| t_CR_s3            | cognitive reserve total score (c2011_w4) | Continous | 0-25 |

# Variable description

Cognitive reserve is the ability of the cognitive paradigm underlying a task to
sustain disruption and still operate effectively. Alternately, this could
consist of the ability to use alternate paradigms to approach a problem when the
more standard approach is no longer operational. The concept of cognitive
reserve provides a ready explanation for why many studies have demonstrated that
higher levels of intelligence and of educational and occupational attainment are
good predictors of which individuals can sustain greater brain damage before
demonstrating functional deficit. Rather than positing that these individuals'
brains are grossly anatomically different than those with less reserve (e.g.,
they have more synapses), the cognitive reserve hypothesis posits that they
process tasks in a more efficient manner.

# Explanation of how the variables were created

We have included the cognitive questionnaire from [@ramigonzález2011], which is
composed by the following items: q1016_highest_s3 (education level), q1052_s3
(training course), q1501a_s3 (languages perfectly spoken), q1053_s3 (musical
instrument), q1051_s3 (languages), q1054_s3 (number of books read per year),
q1055_s3 (intellectual games), q1050_s3 (education level of parents). Item q1052
ranges from 0 (none) to 3 (more than 5), items q1053 from 0 (none) to 2 (good
formation), items q1054 from 0 (none) to 4 (more than 10) and item q1055 ranges
from 0 (never or sometimes) to 2 (more than 5 times per month). Transformations
were set to allow only valid values for q1016_highest_s3 (0 "No education" 1
"Basic or primary" 2 "Secondary") and q1050_s3 (0 "Not qualified" 1 "Manual
qualification" 2 "Not manual qualification" 3 "Professional" 4 "Manager"). We
have created a composite score by adding the score obtained for each item, with
higher scores indicating better cognitive reserve. Values range from 0 to 25.

Note: we have not included the value 1 in level_edu, because we don´t have this
category (Lee y escribe de manera autodidacta) in our variable q1016_highest_s3.
Instead, we have "no formal education" or "primary studies incompleted."

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 4:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_4\Cohorte 2011\rawdata_c2011w4.dta"

```

```{stata generate-outcomes, echo=!DOUBLE_CODING}
/* Level of education */
recode q1016_highest_s3 (0=0) (1=2) (2=3) (3/4=4) (5/6=5) (else=.), gen (level_edu)
label variable level_edu "highest level of education at wave 1"
label define dEdu 0 "No education" 2 "Basic" 3 "Primary" 4 "Secondary" 5 "Tertiary"
label values level_edu dEdu

/* Parents´ level of education */
recode q1050_s3 (0=0) (1=1) (2=1) (3/6=2) (else=.), gen (edu_parent)

label define dEduP 0 "No education" 1 "Basic or primary" 2 "Secondary" 
label values edu_parent dEduP
label variable edu_parent "Highest education level of either parent"

/* Occupation */
gen occupation=.
replace occupation=0 if q1501b_s3==0 
replace occupation=1 if q1501b_s3==1 
replace occupation=2 if q1501b_s3==2 
replace occupation=3 if q1501b_s3==3 
replace occupation=4 if q1501b_s3==4 
label variable occupation "Occupational level"
label define dOccupation 0 "Not qualified" 1 "Manual qualification" 2 "Not manual qualification" 3 "Professional" 4 "Manager"
label values occupation dOccupation

/* cognitive reserve */
egen t_CR_s3 =rowtotal(level_edu edu_parent q1051_s3 q1052_s3 occupation q1053_s3 q1054_s3 q1055_s3)
label variable t_CR_s3 "cognitive reserve total score (c2011_w4)"
```

<br>

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup', include=!DOUBLE_CODING}
keep ID_ECS q0007a_result proxy t_CR_s3

sum t_CR_s3, detail
```

```{stata save-data-test, eval=TEST}
saveold "Outcome_cognitivereserve.dta", replace version(13)
```

```{stata save-data, eval=!TEST}

* Cohorte 2011, ola 4:
saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 4\Outcomes\Cohorte 2011\Outcome datasets\Outcome_cognitivereserve.dta", replace version(13)

```

<br>

# Contact person:

-   Name: Lea Francia
-   e-mail: [lea.francia\@inv.uam.es](mailto:lea.francia@inv.uam.es){.email}
-   Phone number: (+34) 914974601

Original version:

-   Name: Dario Moreno Agostino
-   e-mail: [d.moreno\@ucl.ac.uk](mailto:d.moreno@ucl.ac.uk){.email}

<br>

-   Name: Elvira Lara
-   e-mail: [elvira.lara\@uam.es](mailto:elvira.lara@uam.es){.email}
-   Phone number: (+34) 914974601

<br>

-   Name: Daniel Morillo
-   e-mail:
    [daniel.morillo\@inv.uam.es](mailto:daniel.morillo@inv.uam.es){.email}
-   Phone number: (+34) 914975816

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
