---
title: "Workload estimation"
author: "Daniel Morillo"
format:
  docx:
    table-of-contents: true
    self-contained:    true
    fig-width:          7.75
    fig-height:         4.25
    df-print:          kable
knitr:
  opts_chunk:
    echo:    false
    error:   false
    warning: false
    message: false
---

```{r setup}
library(knitr)
library(patchwork)
library(gtsummary)
library(scales)
```

```{r source}
read_chunk("../src/C2011_W4_Planning.R")
```

```{r main}
```

# Planning

The total number of outcomes assigned to each team member is shown in
@fig-num-outcomes-output. From left to right, the figure shows the datasets
assigned to be coded by each member, reviewed by each member, and the sum of
both. The same information can be seen in @fig-coding-date disaggregated by due
date.

```{r num-outcomes}
```

```{r num-outcomes-output}
#| label:   fig-num-outcomes-output
#| fig-cap: Number of outcome datasets by team member
coding_load <- coding_load + ggtitle("coded")
review_load <- review_load + ggtitle("reviewed")
total_load  <- total_load  + ggtitle("total (code + reviewed)")

coding_load + review_load + total_load
```

```{r coding-date}
#| label:   fig-coding-date
#| fig-cap: Number of outcome datasets by team member and due date
#| fig-height: 8.5
coding_date <- coding_date + ggtitle("coded")    + guides(color = guide_none())
review_date <- review_date + ggtitle("reviewed") + guides(color = guide_none())
total_date  <- total_date            +
  ggtitle("total (code + reviewed)") +
  guides(color = guide_legend(title = NULL, direction = "horizontal")) +
  theme(legend.position = "bottom")

coding_date / review_date / total_date
```

# Completion

```{r}
completion_data <- planning_data  |>
  select(
    SECTION:`VARIABLE (s)`,
    `Due date` = `DUE DATE...16`,
    Complete   = COMPLETE...17,
    Incidence  = REMARKS,
    starts_with("Errors")
  ) |>
  rename_with(str_to_sentence) |>
  mutate(Complete = Complete |> coalesce("No") |> str_to_sentence())

n_datasets_total    <- completion_data |> nrow()
n_datasets_complete <- completion_data |>
  filter(Complete == "Yes") |>
  nrow()

incomplete_datasets <- completion_data |>
  filter(Complete == "No") |>
  select(Section:`Variable (s)`, Incidence)
```

There were `r n_datasets_total` datasets planned to be coded. Out of them,
`r n_datasets_complete` have been successfully coded (see @tab-completion-out).
However, some datasets have been found to be dependent on non-existent outcomes
from previous waves. That means that some variables could not be created because
the variables they depend on have not been created in previous waves, either.
These are reported in @tbl-dependent-detail-out. Also, the details of the
datasets that have not been coded are shown in @tbl-incomplete-detail-out.

```{r dependent-detail-out}
#| label:   tbl-dependent-detail-out
#| tbl-cap: |
#|   Detail of complete datasets that have had incidences due to being dependent
#|   on non-existing variables in previous waves.
completion_data |>
  filter(Incidence == '"Lifetime" needs to be created') |>
  select(-`Due date`, -Complete)
```

```{r incomplete-detail-out}
#| label:   tbl-incomplete-detail-out
#| tbl-cap: Detail of the incomplete datasets.
incomplete_datasets
```

# Error detection and correction

```{r error-data}
error_data <- completion_data |> filter(Complete == "Yes")

errors_summary <- error_data |>
  pivot_longer(starts_with("Error"), names_to = "Variable") |>
  mutate(value = value |> coalesce("No")) |>
  group_by(Variable) |>
  count(value) |>
  mutate(percentage = n / sum(n)) |>
  filter(value == "Yes")

n_errors_dc <- errors_summary |>
  filter(Variable == "Errors detected (with double coding)") |>
  pull(n)
perc_errors_dc <- errors_summary |>
  filter(Variable == "Errors detected (with double coding)") |>
  pull(percentage) |>
  percent(accuracy = .1)

n_errors_pw <- errors_summary |>
  filter(Variable == "Errors in previous waves") |>
  pull(n)
perc_errors_pw <- errors_summary |>
  filter(Variable == "Errors in previous waves") |>
  pull(percentage) |>
  percent(accuracy = .1)

```

The double coding process implemented has helped detect and correct coding
errors in `r n_errors_dc` out of the `r n_datasets_complete` coded datasets;
this constitutes a `r perc_errors_dc` as represented in @fig-error-data-out.
It is worth noting that **these errors have only been detected and corrected by
the independent double coding, and not the review process**, which highlights
the need to implement this process in the coding of the outcome datasets of
future waves. Additionally, by using the data from previous waves as starting
point for the coding of the new data, we have detected errors in `r n_errors_pw`
of the equivalent datasets in previous waves, representing a `r perc_errors_pw`
of the coded datasets (see @fig-error-data-out).

```{r error-data-out}
#| label: fig-error-data-out
#| fig-cap: |
#|   Errors detected with double coding (left) and in the datasets in previous
#|   waves.

# Gauge adapted from https://pomvlad.blog/2018/05/03/gauges-ggplot2/
errors_summary |>
  ggplot(
    aes(fill = Variable, ymax = percentage, ymin = 0, xmax = 2, xmin = 1)
  ) +
  geom_rect(aes(ymax = 1, ymin = 0, xmax = 2, xmin = 1), fill ="#ece8bd") +
  geom_rect() + 
  coord_polar(theta = "y", start = -pi / 2) +
  xlim(c(0, 2)) +
  ylim(c(0, 2)) +
  geom_text(aes(x = 0, y = 0, label = n, colour = Variable), size = 6.5) +
  geom_text(aes(x = 1.5, y = 1.5, label = Variable), size = 4.2) +
  facet_wrap(~Variable) +
  theme_void() +
  scale_fill_manual(
    values = c(
      "Errors detected (with double coding)" = "red",
      "Errors in previous waves"             = "#DA9112"
    ),
    aesthetics = c("colour", "fill")
  ) +
  theme(strip.background = element_blank(), strip.text.x = element_blank()) +
  guides(fill = "none", colour = "none")
```

# Dedication

The total nº of hours dedicated by each team member is detailed in
@tbl-dedication-out. These hours are an estimate, and include the time dedicated
to coding, review, checking the original syntax with the double coding, internal
communication, incidence resolution, reporting, etc. Also, other members of the
Edad con Salud team have been involved in the approval of the final versions of
the datasets, which has included solving inconsistencies and doubts, reviewing
and suggesting changes, and attending meetings.

Information of the dedication of these members and finer-detail information of
the dedication of the members assigned to the task are not available.

```{r}
#| label: tbl-dedication-out
#| tbl-cap: Estimate of dedicated hours to the creation of the outcome datasets
tribble(
  ~Member,     ~`Estimate (hours)`,
   "Cristina",  453,
   "Daniel",    321,
   "Lea",       670
)
```

# Conclusions

* The coding of the 4-th wave outcome datasets of the 2011 Cohort of Edad con
Salud has taken a large amount of the working hours of the three team members
involved.

* The independent double coding procedure has proven to be valuable for
detecting and correcting potential errors in the datasets.
