---
title: "Edad con Salud -- cohort 2019 -- COVID substudy -- 'Health status'"
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
    df_print: kable
bibliography: "`r ecs.data::get_proj_bib_file()`"
csl: "`r ecs.data::get_proj_csl_file()`"
# nocite: '@<cita_1>, @<cita_2>' ## TODO: Descomentar y añadir cita(s) o borrar campo
---

```{r setup, include=FALSE}
library(knitr)
library(scales)

# Opciones de configuración de las salidas de R y Stata:

options(width = 88) # Ancho de salida de texto en R

opts_chunk$set( # Para todos los 'chunks' de código:
  echo    = TRUE,  #   - Muestra toda la sintaxis en el documento
  results = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  warning = FALSE, #   - Omite avisos
  message = FALSE, #   - Omite mensajes
  comment = ''     #   - No añade ningún prefijo a los resultados
)
```

# Variable information

+---------------------+------------------------------------------+------------+----------------------+
| Main variable name  | Label                                    | Format     | Code                 |
+:====================+:=========================================+:===========+:=====================+
| **health_abb**      | Abbreviated health status (c2019_wCOVID) | continuous | (standardized)       |
+---------------------+------------------------------------------+------------+----------------------+

# Variable description

`health_abb` is an abbreviated (standardized) health status score,
with higher values meaning better health.
Health status was assessed with a subset of items from the Whodas questionnaire.
For each question, the responses were recorded on a five-point scale
ranging from 1 (no difficulty/problem) to 5 (extreme difficulty/inability).
A composite health status score was derived using the Partial Credit Model
[@wilson_improving_2006].

# Explanation of how the variables were created

Items from Section SF, and item `DOLOR1`, with a five-categories response,
were used when calibrating (health) item difficulties
and person abilities (individual health levels) in a Partial Credit Model.
A total of 14 items have been included;
these were chosen because the corresponding ones are part of the total
pool of items used when computing the "Global health status" variable in the
main "Edad con Salud" study [@salomon_quantifying_2003].

The method employed to generate the global score was based on
a Partial Credit Model [@wilson_improving_2006],
using R package `mirt` v1.31.6 [@chalmers_mirt_2012] to estimate
the person ability estimate for each participant, based on maximum likelihood.

**Please note:** The metric of an IRT model latent scores is "arbitrary";
that is, the measurement scale is not identified.
Although the original variable this one is based on was "normalized" on a 0-100
scale, we have opted here to leave the scores as standardized values, in order
to avoid further distortions of the scores.
The most important takeaway is that a pre-post contrast of difference-of-means
cannot be performed directly from these score estimates.
However, a multiple linear regression, e.g. with the post- measure as response
and the pre- as a predictor may be performed without problems.

# Syntax used to create the variable (R 4.1.3)

```{r includes}
# Activate packages:

library(tidyverse)
library(haven)
library(ecs.data)
library(mirt)
library(gtsummary)
```

```{r constants}
HEALTH_VARS <- c(
  # General health status:
  "SF1",
  "SF2_1",  "SF2_2",  "SF2_3",
  "SF2_4",  "SF2_5",  "SF2_6",
  "SF2_7",  "SF2_8",  "SF2_9",
  "SF2_10", "SF2_11", "SF2_12",
  
  # Pain & discomfort:
  "DOLOR1"
)

COVID_POST_SURVEY_DATASET <- file.path(
  read_ecs_folder("DB"),
  "Subestudio_COVID",
  "Edad_con_salud_Fichero_Completo.dta"
)
```

```{r load-data, echo=TRUE}
survey_data <- read_dta(COVID_POST_SURVEY_DATASET)
```

```{r generate-outcomes, echo=TRUE}
survey_data <- survey_data                  |>
  mutate(
    across(!!HEALTH_VARS, na_if, 888),
    across(!!HEALTH_VARS, na_if,   9)
  ) |>
  # Filter cases where there are no responses to the Whodas items:
  filter(ESTADO_ENTREVISTA %in% c(1, 9))

health_status_data <- survey_data |> select(!!HEALTH_VARS)

hs_irt_model <- health_status_data |> mirt(model = 1, itemtype = "Rasch")

hs_scores <- hs_irt_model |>
  fscores()               |>
  scale()                 |>
  drop()

# Invert the scores so a higher score means a better health status
hs_scores <- -hs_scores

attr(hs_scores, "label") <- "Abbreviated health status (c2019_wCOVID)"

output <- survey_data               |>
  select(ID_ECS, ESTADO_ENTREVISTA) |>
  add_column(health_abb = hs_scores)
```

**Model fit statistics:**

```{r model-fit, results='asis'}
# Goodness of fit statistics:
hs_irt_model                                         |>
  M2(na.rm = TRUE)                                   |>
  pivot_longer(everything(), names_to = "statistic") |>
  mutate(value = value |> number(accuracy = 1e-3))   |>
  kable(row.names = FALSE, align = 'rl')
```

<br>

**Variable descriptive statistics:**

```{r descriptives, results='asis'}
output               |>
  select(health_abb) |>
  tbl_summary(
  type = health_abb ~ "continuous2",
  statistic = all_continuous2() ~ c(
      "{mean} ({sd})", 
      "{median} ({p25}, {p75})", 
      "{min}, {max}"
    )
  )
```

<br>

```{r save-data, echo=FALSE}
output |> write_dta(
  file.path(
    read_ecs_folder("DOC"),
    "Edad con salud - Subestudio COVID/Outcomes/Outcome datasets",
    "Outcome_healthstatus.dta"
  ),
  version = 13
)
```

<br>

# Contact person:

-   Name:         Daniel Morillo (adaptation for Cohort 2019, Wave 1)
-   e-mail:       daniel.morillo@uam.es
-   Phone number: (+34) 91 497 58 16

The creation of this outcome variable is based on the wave-2 equivalent
of the 2011 Cohort from the Edad con Salud study
(see document [Description_HealthStatus.doc][1]).
For the original document, please refer to:

[1]: https://dauam-my.sharepoint.com/:w:/r/personal/marta_miret_uam_es/Documents/Edad%20con%20Salud/Documentacion%20Edad%20con%20Salud/Edad%20con%20salud%20-%20Ola%202/Outcomes/Outcome%20descriptions/Description_HealthStatus.doc?d=wde2429dfc1bf4bc884d84b8e461fda79&csf=1&web=1&e=3tovyX

-   Name:         Félix Caballero
-   e-mail:       felix.caballero@uam.es
-   Phone number: (+34) 91 497 24 47

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
