---
title: "Edad con Salud -- cohort 2019 -- COVID substudy -- 'Health status'"
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
    df_print: kable
bibliography: "`r ecs.data::get_proj_bib_file()`"
csl: "`r ecs.data::get_proj_csl_file()`"
# nocite: '@<cita_1>, @<cita_2>' ## TODO: Descomentar y añadir cita(s) o borrar campo
---

```{r setup, include=FALSE}
library(knitr)
library(scales)

# Opciones de configuración de las salidas de R y Stata:

options(width = 88) # Ancho de salida de texto en R

opts_chunk$set( # Para todos los 'chunks' de código:
  echo    = TRUE,  #   - Muestra toda la sintaxis en el documento
  results = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  warning = FALSE, #   - Omite avisos
  message = FALSE, #   - Omite mensajes
  comment = ''     #   - No añade ningún prefijo a los resultados
)
```

# Variable information

+---------------------+------------------------------------------+------------+----------------------+
| Main variable name  | Label                                    | Format     | Code                 |
+:====================+:=========================================+:===========+:=====================+
| **health_abb_post** | Abbreviated health status (c2019_wCOVID) | continuous | (standardized)       |
+---------------------+------------------------------------------+------------+----------------------+

# Variable description

`health_abb_post` is an abbreviated (standardized) health status score,
with higher values meaning better health.
Health status was assessed with a subset of items from the Whodas questionnaire.
For each question, the responses were recorded on a five-point scale
ranging from 1 (no difficulty/problem) to 5 (extreme difficulty/inability).
A composite health status score was derived using the Partial Credit Model.

# Explanation of how the variables were created

Items from Section SF, and item `DOLOR1`, with a five-categories response,
were used when calibrating (health) item difficulties
and person abilities (individual health levels) in a Partial Credit Model.
A total of 14 items have been included;
these were chosen because the corresponding ones are part of the total
pool of items used when computing the "Global health status" variable in the
main "Edad con Salud" study.

The method employed to generate the global score was based on
a Partial Credit Model, using R package `mirt` v1.31.6 to estimate
the person ability estimate for each participant, based on maximum likelihood.

**Please note:** The metric of an IRT model latent scores is "arbitrary";
that is, the measurement scale is not identified.
Although the original variable this one is based on was "normalized" on a 0-100
scale, we have opted here to leave the scores as standardized values, in order
to avoid further distortions of the scores.
The most important takeaway is that a pre-post contrast of difference-of-means
cannot be performed directly from these score estimates.
However, a multiple linear regression, e.g. with the post- measure as response
and the pre- as a predictor may be performed without problems.

# Syntax used to create the variable (R 4.1.3)

```{r includes}
# Activate packages:

library(tidyverse)
library(haven)
library(ecs.data)
library(mirt)
library(gtsummary)
```

```{r constants}
HEALTH_VARS <- c(
  # General health status:
  "SF1",
  "SF2_1",  "SF2_2",  "SF2_3",
  "SF2_4",  "SF2_5",  "SF2_6",
  "SF2_7",  "SF2_8",  "SF2_9",
  "SF2_10", "SF2_11", "SF2_12",
  
  # Pain & discomfort:
  "DOLOR1"
)

COVID_POST_SURVEY_DATASET <- file.path(
  read_ecs_folder("DB"),
  "Subestudio_COVID",
  "Edad_con_salud_Fichero_Completo.dta"
)
```

```{r load-data, echo=TRUE}
survey_data <- read_dta(COVID_POST_SURVEY_DATASET)
```

```{r generate-outcomes, echo=TRUE}
survey_data <- survey_data                  |>
  mutate(
    across(!!HEALTH_VARS, na_if, 888),
    across(!!HEALTH_VARS, na_if,   9)
  ) |>
  # Filter cases where there are no responses to the Whodas items:
  filter(ESTADO_ENTREVISTA %in% c(1, 9))

health_status_data <- survey_data |> select(!!HEALTH_VARS)

hs_irt_model <- health_status_data |> mirt(model = 1, itemtype = "Rasch")

hs_scores <- hs_irt_model |>
  fscores()               |>
  scale()                 |>
  drop()

# Invert the scores so a higher score means a better health status
hs_scores <- -hs_scores

attr(hs_scores, "label") <- "Abbreviated health status (c2019_wCOVID)"

output <- survey_data               |>
  select(ID_ECS, ESTADO_ENTREVISTA) |>
  add_column(health_abb_post = hs_scores)
```

**Model fit statistics:**

```{r model-fit, results='asis', echo=FALSE}
# Goodness of fit statistics:
hs_irt_model                                         |>
  M2(na.rm = TRUE)                                   |>
  pivot_longer(everything(), names_to = "statistic") |>
  mutate(value = value |> number(accuracy = 1e-3))   |>
  kable(row.names = FALSE, align = 'rl')
```

<br>

**Variable descriptive statistics:**

```{r descriptives, results='asis', echo=FALSE}
output                    |>
  select(health_abb_post) |>
  tbl_summary(
  type = health_abb_post ~ "continuous2",
  statistic = all_continuous2() ~ c(
    "{mean} ({sd})", 
    "{median} ({p25}, {p75})", 
    "{min}, {max}"
  )
  )
```

<br>

```{r save-data}
output |> write_dta(
  file.path(
    read_ecs_folder("DOC"),
    "Edad con salud - Subestudio COVID/Outcomes/Outcome datasets",
    "Outcome_healthstatus.dta"
  ),
  version = 13
)
```

<br>

# Contact person:

-   Name:         Daniel Morillo (adaptation for Cohort 2019, Wave 1)
-   e-mail:       daniel.morillo@uam.es
-   Phone number: (+34) 91 497 58 16

The creation of this outcome variable is based on the wave-2 equivalent
of the 2011 Cohort from the Edad con Salud study
(see document [Description_HealthStatus.doc][1]).
For the original document, please refer to:

[1]: https://dauam-my.sharepoint.com/:w:/r/personal/marta_miret_uam_es/Documents/Edad%20con%20Salud/Documentacion%20Edad%20con%20Salud/Edad%20con%20salud%20-%20Ola%202/Outcomes/Outcome%20descriptions/Description_HealthStatus.doc?d=wde2429dfc1bf4bc884d84b8e461fda79&csf=1&web=1&e=3tovyX

-   Name:         Félix Caballero
-   e-mail:       felix.caballero@uam.es
-   Phone number: (+34) 91 497 24 47

# References

::: {#refs}
Chalmers, R. P. (2012).
mirt: A Multidimensional Item Response Theory Package for the R Environment.
Journal of Statistical Software, 48(1), 1-29.
https://doi.org/10.18637/jss.v048.i06

Salomon JA, Mathers CD, Chatterji S, Sadana R, Üstün TB, Murray CJL (2003).
Quantifying individual levels of health:
definitions, concepts, and measurement levels.
In: Murray CJL, Evans DB, editors.
Health systems performance assessment: debates, methods, empiricism.
Geneva: World Health Organization. pp: 301-318.

Wilson M, Allen DD, Li JC (2006).
Improving measurement in health education and health behavior research using
item response modeling: comparison with the clasical test theory approach.
Health Education Research, 21 (Suppl 1): i19-i32.
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
