---
title: "Edad con Salud -- cohort 2019 -- COVID substudy -- 'Ailment'"
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: "`r ecs.data::get_proj_bib_file()`"
csl: "`r ecs.data::get_proj_csl_file()`"
# nocite: '@<cita_1>, @<cita_2>' ## TODO: Descomentar y añadir cita(s) o borrar campo
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

# Ruta al ejecutable de Stata (Ejecuta `sysdir` en Stata para encontrar la ruta)

STATA_PATH <- r"(C:/Program Files (x86)/Stata15/StataSE-64.exe)"


# Opciones de configuración de las salidas de R y Stata:

options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( # Para todos los 'chunks' de código:
  engine.path = list(stata = STATA_PATH),
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)
```

# Variable information

+--------------------+---------------------------------------------------------------------+--------------+-----------------------------+
| Main variable name | Label                                                               | Format       | Code                        |
+:===================+:====================================================================+:=============+:============================+
| infected           | (Likely) infected with COVID-19 (c2019_wCOVID)                      | dichotomous  | 0 = No;                     |
|                    |                                                                     |              |                             |
|                    |                                                                     |              | 1 = Yes                     |
+--------------------+---------------------------------------------------------------------+--------------+-----------------------------+
| severity_raw       | Severity of COVID-19 infection (c2019_wCOVID)                       | ordinal      | 0 = Not infected;           |
|                    |                                                                     |              |                             |
|                    |                                                                     |              | 1 = Only infected;          |
|                    |                                                                     |              |                             |
|                    |                                                                     |              | 2 = Hospitalized;           |
|                    |                                                                     |              |                             |
|                    |                                                                     |              | 3 = Hospitalized with O2;   |
|                    |                                                                     |              |                             |
|                    |                                                                     |              | 4 = Admitted in ICU         |
+--------------------+---------------------------------------------------------------------+--------------+-----------------------------+
| severity           | Severity of COVID-19 infection, recoded (c2019_wCOVID)              | ordinal      | 0 = Not infected;           |
|                    |                                                                     |              |                             |
|                    |                                                                     |              | 1 = Infected;               |
|                    |                                                                     |              |                             |
|                    |                                                                     |              | 2 = Infected & hospitalized |
+--------------------+---------------------------------------------------------------------+--------------+-----------------------------+
| rel_isolated       | Co-habited/ing with relative isolated by COVID-19 (c2019_wCOVID)    | dichotomous  | 0 = No;                     |
|                    |                                                                     |              |                             |
|                    |                                                                     |              | 1 = Yes                     |
+--------------------+---------------------------------------------------------------------+--------------+-----------------------------+
| rel_concerned      | Concerned about relative/friend infected by COVID-19 (c2019_wCOVID) | dichotomous  | 0 = No;                     |
|                    |                                                                     |              |                             |
|                    |                                                                     |              | 1 = Yes                     |
+--------------------+---------------------------------------------------------------------+--------------+-----------------------------+
| rel_deceased       | Friend/familiar/acquantied deceased by COVID-19 (c2019_wCOVID)      | dichotomous  | 0 = No;                     |
|                    |                                                                     |              |                             |
|                    |                                                                     |              | 1 = Yes                     |
+--------------------+---------------------------------------------------------------------+--------------+-----------------------------+

# Variable description

The following variables are related to the ailment caused by the COVID-19 pandemic:

-   **infected**: Self-reported response of confirmed or likely infection by COVID-19, diagnosed by a physician. (**IMPORTANT:** Likely diagnosis of COVID-19, as may be estimated by the checklist of compatible symptoms in item section COVID2 is not contemplated in this variable.)

-   **severity_raw**: Original variable of severity of the COVID-19 infection, from 0 (no infection) to 4 (hospitalization with oxygen and ICU-admission and artificial breathing).

-   **severity**: Severity of the COVID-19 infection, from 0 (no infection) to 2 (hospitalization). This variable collapses categories 2 to 4 of variable `severity_raw` into a single category (2).

-   **rel_isolated**: Self-resported response of whether the participant lives or has lived with a person isolated at home due to COVID-19 infection.

-   **rel_concerned**: Self-resported response of whether the participant is or has been especially concerned about a relative or friend infected with COVID-19.

-   **rel_deceased**: Self-resported response of whether any participant's relative, friend, or acquaintance, has passed away due to COVID-19 infection.

# Explanation of how the variables were created

-   **infected**: Item `COVID1` recoded as `0` (*No*) or `1` (*Yes*).

-   **severity_raw**: Affirmative responses to items `COVID1`, `COVID1B`, `COVID1C_1`, and `COVID1C_2` are summed up.

-   **severity**: From `severity_raw`, response categories `2` to `4` are collapsed into category `2` (= *Hospitalized*).

-   **rel_isolated**: Item COVID3 recoded as `0` (*No*) or `1` (*Yes*).

-   **rel_concerned**: Item COVID4 recoded as `0` (*No*) or `1` (*Yes*).

-   **rel_deceased**: Item COVID5 recoded as `0` (*No*) or `1` (*Yes*).

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2019, ola post-confinamiento:
use "~\UAM\marta.miret@uam.es - Bases de datos maestras Edad con Salud\Subestudio_COVID\Edad_con_salud_Fichero_Completo.dta"
```

```{stata generate-outcomes, echo=TRUE}
* Recode COVID item responses ("No" responses as 0)
recode COVID1 COVID1B COVID1C_1 COVID1C_2 COVID3 COVID4 COVID5 (2=0)

* Label definition for dichotomous variables
label define BOOL_LABELS 0 "No" 1 "Yes"

* infected:
gen infected = COVID1
label values infected BOOL_LABELS
label variable infected "(Likely) infected with COVID-19 (c2019_wCOVID)"
replace infected=. if ESTADO_ENTREVISTA != 1 & ESTADO_ENTREVISTA != 9

* severity_raw:
egen severity_raw = rowtotal(COVID1 COVID1B COVID1C_1 COVID1C_2)
replace severity_raw=. if COVID1 == .
label define dSeverityRaw 0 "Not infected" 1 "Only infected" 2 "Hospitalized" 3 "Hospitalized with O2" 4 "Admitted in ICU"
label values severity_raw dSeverityRaw
label variable severity_raw "Severity of COVID-19 infection (c2019_wCOVID)"
replace severity_raw=. if ESTADO_ENTREVISTA != 1 & ESTADO_ENTREVISTA != 9

* severity:
gen severity = severity_raw
recode severity (3=2) (4=2)
label define dSeverity 0 "Not infected" 1 "Infected" 2 "Infected & hospitalized"
label values severity dSeverity
label variable severity "Severity of COVID-19 infection, recoded (c2019_wCOVID)"
replace severity=. if ESTADO_ENTREVISTA != 1 & ESTADO_ENTREVISTA != 9

* rel_isolated:
gen rel_isolated = COVID3
label values rel_isolated BOOL_LABELS
label variable rel_isolated "Co-habited/ing with relative isolated by COVID-19 (c2019_wCOVID)"
replace rel_isolated=. if ESTADO_ENTREVISTA != 1 & ESTADO_ENTREVISTA != 9

* rel_concerned:
gen rel_concerned = COVID4
label values rel_concerned BOOL_LABELS
label variable rel_concerned "Concerned about relative/friend infected by COVID-19 (c2019_wCOVID)"
replace rel_concerned=. if ESTADO_ENTREVISTA != 1 & ESTADO_ENTREVISTA != 9

* rel_deceased:
gen rel_deceased = COVID5
label values rel_deceased BOOL_LABELS
label variable rel_deceased "Friend/familiar/acquantied deceased by COVID-19 (c2019_wCOVID)"
replace rel_deceased=. if ESTADO_ENTREVISTA != 1 & ESTADO_ENTREVISTA != 9
```

<br> <!--# Salto de línea (i.e. párrafo vacío); añadir donde se necesite -->

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup'}
codebook infected severity_raw severity rel_isolated rel_concerned rel_deceased
```

```{stata save-data}
keep ID_ECS ESTADO_ENTREVISTA infected severity_raw severity rel_isolated rel_concerned rel_deceased

* Cohorte 2019, ola post-confinamiento:
saveold "~\UAM\marta.miret@uam.es - Documentacion Edad con Salud\Edad con salud - Subestudio COVID\Outcomes\Outcome datasets\Outcome_ailment.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Daniel Morillo
-   e-mail: [\<daniel.morillo\@cibersam.es>](mailto:daniel.morillo@cibersam.es){.email}
-   Phone number: (+34) 91 497 58 16

# References

(None)

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
