---
title: "Subestudio COVID -- cohort 2019 -- wave 1 -- Economic"
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: "`r ecs.data::get_proj_bib_file()`"
csl: "`r ecs.data::get_proj_csl_file()`"
---

```{r setup, include=FALSE}
library(Statamarkdown)

STATA_PATH <- r"(C:\Stata 16\Stata 16.exe)"

options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( # Para todos los 'chunks' de código:
  ## TODO: Descomentar si hace falta
  engine.path = list(stata = STATA_PATH),
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)
```

# Variable information

+---------------+------------------------------------------------------------------------------+-------------+---------------------+
| Variable name | Label                                                                        | Format      | Code                |
+===============+==============================================================================+=============+=====================+
| quietness     | Quietness at home (c2019_wCOVID)                                             | Dichotomous | 0 = No              |
|               |                                                                              |             |                     |
|               |                                                                              |             | 1 = Yes             |
+---------------+------------------------------------------------------------------------------+-------------+---------------------+
| internet      | Internet connection at home (c2019_wCOVID)                                   | Dichotomous | 0 = No              |
|               |                                                                              |             |                     |
|               |                                                                              |             | 1 = Yes             |
+---------------+------------------------------------------------------------------------------+-------------+---------------------+
| home_score    | Home score (measure of size and whether it is outward facing) (c2019_wCOVID) | Dichotomous | standardized (0, 1) |
+---------------+------------------------------------------------------------------------------+-------------+---------------------+
| economy       | Economic situation worsened due to COVID-19 measures (c2019_wCOVID)          | Dichotomous | 0 = No              |
|               |                                                                              |             |                     |
|               |                                                                              |             | 1 = Yes             |
+---------------+------------------------------------------------------------------------------+-------------+---------------------+
| unemployment  | (Temporally) unemployed due to COVID-19 measure (c2019_wCOVID)               | Dichotomous | 0 = No              |
|               |                                                                              |             |                     |
|               |                                                                              |             | 1 = Yes             |
+---------------+------------------------------------------------------------------------------+-------------+---------------------+

# Variable description

The following variables are related to characteristics of the house of the participant and the changes in the economy caused by the COVID-19 pandemic:

* **quietness**: Self-reported response of whether there is a lot of noise in the house and it does not let them rest.

* **internet**: Self-reported response of whether the participant's home has internet access. 

* **home_score**: Total of items related to the size of the house and if it has a view to the outside.


* **economy**: Self-reported response of whether the economy has worsened due to COVID-19.

* **unemployment**: Self-reported response of whether the participant is unemployed indefinitely or temporarily due to COVID-19

# Explanation of how the variables were created

* **quietness**: Reversed item `ECON1_1` recoded as  `0` (*Yes*) or `1` (*No*).

* **internet**: Item `ECON1_2` recoded as  `0` (*No*) or `1` (*Yes*).

* **home_score**: Standardization of the sum of the standardized scores `ECON1_3, ECON1_4, ECON1_5, ECON2 and ECON3`

* **economy**: Item `ECON4` recoded as  `0` (*No*) or `1` (*Yes*).

* **unemployment**: Item `ECON5` recoded as  `0` (*No*) or `1` (*Yes*).

# Syntax used to create the variable (Stata 16)

```{stata load-data}
* Cohorte 2019, ola post-confinamiento:
use "~\UAM\marta.miret@uam.es - Bases de datos maestras Edad con Salud\Subestudio_COVID\Edad_con_salud_Fichero_Completo.dta"
```

```{stata generate-outcomes, echo=TRUE}
* Generate variables from questionnaire and recoding
gen quietness = ECON1_1 //Reversed item
replace quietness = 0 if quietness == 1
replace quietness = 1 if quietness == 2
replace quietness = . if ECON1_1 == 9

gen internet = ECON1_2
replace internet = 0 if internet == 2
replace internet = . if ECON1_2 == 9

gen terrace = ECON1_3
replace terrace = 0 if terrace == 2
replace terrace = . if ECON1_3 == 9

gen garden = ECON1_4
replace garden = 0 if garden == 2
replace garden = . if ECON1_4 == 9

gen exterior = ECON1_5	//Reversed item
replace exterior = 0 if exterior == 1
replace exterior = 1 if exterior == 2
replace exterior = . if ECON1_5 == 9

gen area = ECON2
replace area = . if ECON2 == 999

gen bedrooms = ECON3
replace bedrooms = . if ECON3 == 99

gen economy = ECON4
replace economy = 0 if economy == 2
replace economy = . if ECON4 == 9

gen unemployment = ECON5
replace unemployment = 0 if unemployment == 2
replace unemployment = . if ECON5 == 9

* Generate home_score
* First standardization
egen float z_terrace = std(terrace), mean(0) std(1)
egen float z_garden = std(garden), mean(0) std(1)
egen float z_exterior = std(exterior), mean(0) std(1)
egen float z_area = std(area), mean(0) std(1)
egen float z_bedrooms = std(bedrooms), mean(0) std(1)
* Total 
egen home_score_suma = rowtotal(z_terrace z_garden z_exterior z_area z_bedrooms), missing
* If the interview status is 1 (access) or 9 (proxy)
gen home_score_filtro = .
replace home_score_filtro = home_score_suma if ESTADO_ENTREVISTA == 1 | ESTADO_ENTREVISTA == 9
* Total standarization
egen float home_score = std(home_score_filtro), mean(0) std(1)

* Labels 
label define BOOL_LABELS 0 "No" 1 "Yes"

label variable quietness "Quietness at home (c2019_wCOVID)"
label values quietness BOOL_LABELS

label variable internet "Internet connection at home (c2019_wCOVID)"
label values internet BOOL_LABELS

label variable home_score "Home score (c2019_wCOVID)"

label variable economy "Economic situation worsened due to COVID-19 measures (c2019_wCOVID)"
label values economy BOOL_LABELS

label variable unemployment "(Temporally) unemployed due to COVID-19 measures (c2019_wCOVID)"
label values unemployment BOOL_LABELS
```

<br> <!--# Salto de línea (i.e. párrafo vacío); añadir donde se necesite -->

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup'}
codebook quietness internet home_score economy unemployment
```


```{stata save-data}
keep ID_ECS ESTADO_ENTREVISTA quietness internet home_score economy unemployment

* Cohorte 2019, ola post-confinamiento:
saveold "~\UAM\marta.miret@uam.es - Documentacion Edad con Salud\Edad con salud - Subestudio COVID\Outcomes\Outcome datasets\Outcome_economic.dta", replace version(13)
```

<br>

# Contact person:

-   Name: Jesús Victor Sierralaya Baldeón
-   e-mail: [\ jesus.sierralaya\@inv.uam.es](mailto:jesus.sierralaya@inv.uam.es){.email}
-   Phone number: (+34) 91 497 58 09

<br>

<!--# TODO: Completar / eliminar "original version" según corresponda -->

\ Original version:

-   Name: Daniel Morillo
-   e-mail: [\ daniel.morillo\@cibersam.es](mailto:daniel.morillo@cibersam.es){.email}
-   Phone number: (+34) 91 497 58 16

# References

(None)

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
