---
title: "Edad con Salud -- cohort 2019 -- wave 1 -- List of threatening experience" 
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: "`r ecs.data::get_proj_bib_file()`"
csl: "`r ecs.data::get_proj_csl_file()`"
---

```{r setup, include=FALSE}
library(Statamarkdown)

STATA_PATH <- r"(C:\Stata 16\Stata 16.exe)"

options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( # Para todos los 'chunks' de código:
  ## TODO: Descomentar si hace falta
  engine.path = list(stata = STATA_PATH),
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE,  #   - Reutiliza el código de Stata de 'chunks' anteriores
  comment     = ''     #   - No añade ningún prefijo a los resultados
)
```

# Variable information

+--------------------+--------------------------------------------+-------------+-------------+
| Main variable name | Label                                      | Format      | Code        |
+:===================+:===========================================+:============+:============+
| LTE                | List of Threatening Experiences (c2019_w1) | Discrete    | 0-12        |
+--------------------+--------------------------------------------+-------------+-------------+


# Variable description

List of Threatening Experiences (LTE) provides a stressful event counter and is used to measure the risk of psychosomatic illnesses. It takes into account the following social situations: illness, emotional, family, work and legal relationships. A high number of LTEs can cause the appearance of mental illnesses (anxiety, depression) and biological illnesses (arterial hypertension, heart attacks, etc). Brugha marks that this shortened version is as accurate as one of 60 categories.

# Explanation of how the variables were created

The 12 items (q4659-q4670) of the LTE instrument have been collected directly in the questionnaire. In the sum, it considers the missing data as the absence of the event.

# Syntax used to create the variable (Stata 16)

```{stata load-data}
* Cohorte 2019, linea base:
use "~\UAM\marta.miret@uam.es - Bases de datos maestras Edad con Salud\Ola_3\Cohorte_2019\rawdata_c2019w1.dta"

```

```{stata generate-outcomes, echo=TRUE}
recode q4659_ill q4660_illrelative q4661_died q4662_diedrelative q4663_separation q4664_brokeoff q4665_problem q4666_unemployed q4667_sacked q4668_financial q4669_police q4670_stolen (888=.)

egen LTE = rowtotal(q4659_ill q4660_illrelative q4661_died q4662_diedrelative q4663_separation q4664_brokeoff q4665_problem q4666_unemployed q4667_sacked q4668_financial q4669_police q4670_stolen), missing

label variable LTE "List of Threatening Experiences (c2019_w1)"
```

<br> <!--# Salto de línea (i.e. párrafo vacío); añadir donde se necesite -->

```{stata descriptives, cleanlog=FALSE, collectcode=FALSE, results='markup'}
tab LTE, m
```

```{stata save-data}
* TODO: Sustituir '<variable(s) generada(s)>' por las variables del dataset
* TODO: Borrar `subsample_pre interview_pre` si no procede (dejar sólo para Cohorte 2019, Ola 1)
keep ID_ECS q0007a_result proxy subsample_pre interview_pre LTE

* Cohorte 2019, ola 1:
saveold "~\UAM\marta.miret@uam.es - Documentacion Edad con Salud\Edad con salud - Ola 3\Outcomes\Cohorte 2019\Outcome datasets\Outcome_LTE.dta", replace version(13)

```

<br>

# Contact person:

-   Name: Jesús Victor Sierralaya Baldeón
-   e-mail: [\<jesus.sierralaya\@inv.uam.es\>](mailto:jesus.sierralaya@inv.uam.es){.email}
-   Phone number: (+34) 91 497 58 09

<br>

# References

::: {#refs}
Brugha, T. S., & Cragg, D. (1990). The list of threatening experiences: the reliability and validity of a brief life events questionnaire. Acta Psychiatrica Scandinavica, 82(1), 77-81.
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
