---
title: "Edad con Salud -- cohort 2019 -- Wave 1 -- 'Health status'" ## TODO: Añadir "cohort", "assessment" y "variable name(s)/domain"
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
bibliography: "`r ecs.data::get_proj_bib_file()`"
csl: "`r ecs.data::get_proj_csl_file()`"
nocite: '@salomon2003, @wilson2006'
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(pacman)
p_load(haven, tidyverse)


# Opciones de configuración de las salidas de R y Stata:

options(width = 88) # Ancho de salida de texto en R

knitr::opts_chunk$set( # Para todos los 'chunks' de código:
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  message     = FALSE  #   - Omite los mensajes
)
```

# Variable information

+--------------------+-----------------------------+---------+---------------+
| Main variable name | Label                       | Format  | Code          |
+:===================+:============================+:========+:==============+
| health             | Health status (c2019_w1)    | Numeric | 0-100         |
+--------------------+-----------------------------+---------+---------------+

<br> <!--# Salto de línea (i.e. párrafo vacío); añadir donde se necesite -->

# Variable description

`health` is the health status score ranging in 0-100, with 0 meaning the worst health and 100 the perfect health. Health status was assessed with a set of health-related questions grouped into ten health domains: vision, mobility, self-care, cognition, interpersonal activities, pain and discomfort, sleep and energy, affect domestic life and work, and hearing. For each question, the responses were recorded on a five-point scale ranging from 1 (no difficulty/problem) to 5 (extreme difficulty/inability). A composite health status score was derived using Partial Credit Model.

# Explanation of how the variables were created

Items from Section 2000 with a five-category response scale were used when calibrating (health) item difficulties and person abilities (individual health levels) using a Partial Credit Model. A total of 41 items from Section 2000 have been included, after excluding items `q2009`, `q2026`, and `q2041`, as they were conditional on certain values of previous items. Also, it seems like items `q2053_effort` and `q2054_feel` were not present in waves 1 and 2 of the 2011 Cohort.

The method employed to generate the health status score is based on a Partial Credit Model. We have used R package `mirt` [@chalmers2012] to estimate the person ability estimate for each participant, with a marginal maximum likelihood algorithm. The score has been transformed into a 0-100 scale, where the maximum score of 100 represents best health.

# Syntax used to create the variable (R 4.1.1)

<br>

```{r generate-outcomes, results='hide', echo=TRUE}
library(mirt)

survey_data <- read_dta(
  file.path(
    "~/../UAM",
    "Marta Miret Garcia - Bases de datos maestras Edad con Salud",
    "Ola_3/Cohorte_2019",
    "rawdata_c2019w1.dta"
  )
)


HEALTH_VARS <- c(
  # General health status:
  "q2000_health", "q2001_activ",
  
  # Pain & discomfort:
  "q2007_pains",
  
  # Mobility:
  "q2003_vigor",  "q2036_walk",
  "q2029_climb",  "q2046_going",   "q2045_movement", "q2030_crouch",
  "q2027_stand",  "q2043_arising", "q2025_sitting",  "q2028_stand",
  "q2031_pickup", "q2034_reach",   "q2040_carry",
  
  # Personal care:
  "q2037_wash",   "q2038_dress", "q2044_toilet", "q2042_eating",
  "q2005_appear", "q2006_stay", 
  
  # Cognition:
  "q2035_concent", "q2011_learn",
  
  # Interpersonal activities:
  "q2033_activ",  "q2012_relat", "q2013_confl",
  "q2014_friend", "q2015_strang",
  
  # Domestic life & work:
  "q2032_hh_resp", "q2039_daily", 
  
  # Sleep & energy:
  "q2016_sleep", "q2017_rested",
  
  # Affect:
  "q2018_depr", "q2019_worry", "q2047_affect", "q2048_interf",
  
  # Vision:
  "q2023_vis_diff", "q2024_vis_diff",
  
  # Hearing:
  "q2051_hear_diff", "q2052_hear_diff"
)


survey_data <- survey_data %>%
  mutate(across(all_of(HEALTH_VARS), na_if, 888))

health_status_data <- survey_data %>% select_at(HEALTH_VARS)

hs_irt_model <- health_status_data %>%
  mirt(model = 1, itemtype = "Rasch")

hs_scores <- hs_irt_model %>% fscores() %>% drop()

# Invert the scores so a higher score means a better health status
hs_scores <- -hs_scores

# Normalize the range to 0-100
range_hs           <- hs_scores %>% range()
hs_scores_rescaled <- (hs_scores - range_hs[1]) /
  (range_hs[2] - range_hs[1]) *
  100
```

<br>


```{r descriptives, results='markup', echo=TRUE}
# Model statistics:

hs_irt_model %>% print()


## Goodness of fit:
hs_irt_model %>%
  M2(na.rm = TRUE) %>%
  pivot_longer(
    cols      = everything(),
    names_to  = "Statistic",
    values_to = "Value"
  )


# Score descriptives:
bind_cols(
  health          = hs_scores,
  health_rescaled = hs_scores_rescaled
) %>%
  psych::describe(skew = FALSE) %>%
  select(-vars)
```

```{r save-data}
output <- survey_data %>%
  select(ID_ECS, q0007a_result, proxy, subsample_pre, interview_pre) %>%
  add_column(
    health = hs_scores_rescaled %>%
      labelled(label = "Health status (c2019_w1)")
  )

output %>% write_dta(
  file.path(
    "~/../UAM",
    "Marta Miret Garcia - Documentacion Edad con Salud/Edad con salud - Ola 3",
    "Outcomes/Cohorte 2019/Outcome datasets",
    "Outcome_healthstatus.dta"
  ),
  version = 13
)
```

<br>

# Contact person:

-   Name: Daniel Morillo (adaptation for Cohort 2019, Wave 1)
-   e-mail: [\daniel.morillo\@inv.uam.es\>](mailto:daniel.morillo@inv.uam.es){.email}
-   Phone number: (+34) 91 497 58 16

<br>

Original version:

The creation of this outcome variable is based on the wave-2 equivalent of the 2011 Cohort from the Edad con Salud study (see document [Description_HealthStatus.doc](https://dauam-my.sharepoint.com/:w:/r/personal/marta_miret_uam_es/Documents/Edad%20con%20Salud/Documentacion%20Edad%20con%20Salud/Edad%20con%20salud%20-%20Ola%202/Outcomes/Outcome%20descriptions/Description_HealthStatus.doc?d=wde2429dfc1bf4bc884d84b8e461fda79&csf=1&web=1&e=3tovyX)). For the original document, please refer to:

-   Name: Félix Caballero
-   e-mail: [\felix.caballero\@uam.es\>](mailto:felix.caballero@uam.es){.email}
-   Phone number: (+34) 91 497 24 47

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
