---
title: "Edad con Salud -- cohort 2019 -- Wave 1 -- 'DISABILITY'" 
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
csl: "`r ecs.data::get_proj_csl_file()`"
# nocite: '@<cita_1>', '@<cita_2>' ## TODO: Descomentar y añadir cita(s) o borrar campo
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el código:

library(Statamarkdown)

# Ruta al ejecutable de Stata (Ejecuta `sysdir` en Stata para encontrar la ruta)

## TODO: sustituir '<path\to\>' por la ubicación de Stata
##       (ejecutar `sysdir` en Stata) y descomentar línea 34,
##       o eliminar linea 34:
# STATA_PATH <- r"(<path\to\>StataSE-64.exe)"


# Opciones de configuración de las salidas de R y Stata:

knitr::opts_chunk$set( # Para todos los 'chunks' de código:
## TODO: Descomentar si hace falta
# engine.path = list(stata = STATA_PATH),
echo        = FALSE, #   - No muestra la sintaxis en el documento
results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
collectcode = TRUE   #   - Reutiliza el código de Stata de 'chunks' anteriores
)
```

# Variable information

+------------------------+---------------------------------------------------------------+------------+------------+
| **Main variable name** | **Label**                                                     | **Format** | **Code**   |
+:=======================+:==============================================================+:===========+:===========+
| whodas12               | disability whodas 12 (c2019_w1)                               | Continuous | 0-100      |
+------------------------+---------------------------------------------------------------+------------+------------+
| adl                    | difficulties with activities of daily living (c2019_w1)       | Binary     | 0=no;1=yes |
+------------------------+---------------------------------------------------------------+------------+------------+
| iadl                   | difficulties with instrumental act of daily living (c2019_w1) | Binary     | 0=no;1=yes |
+------------------------+---------------------------------------------------------------+------------+------------+

# Variable description

The level of functional impairment has been assessed using the 12-item interviewer administered version of the World Health Organization (WHO) Disability Assessment Schedule version II (WHODAS 2.0). This instrument is grounded in the conceptual framework of the International Classification of Functioning, Disability and Health (ICF) and places health and disability on a continuum, with disability defined as "a decrement in each functioning domain". Participants were asked to report the level of difficulty they had in doing various activities during the previous 30 days using a five-point scale (none=1, mild=2, moderate=3, severe=4, extreme/cannot do=5). The global score ranges from 0 to 100, with higher scores indicating greater disability.
Level of independence was measured by means of Activities of Daily Living (ADLs) and Instrumental Activities of Daily Living (IADLs). ADLs describes a set of daily self-care activities and assesses the need for help with personal care activities such as eating, bathing, and dressing; while, IADLs describes somewhat higher-level functioning considered necessary to live independently, assessing the need for help with routine needs such as using transportation, housekeeping, and preparing food. An individual's ability to perform ADLs is typically considered normal functional status, with an inability to perform ADLs suggesting disability.

# Explanation of how the variables were created

The WHODAS items were selected and their scoring was determined according to the WHO guidelines. It follows an "item-response-theory" (IRT) based scoring where the summary score takes into account the levels of severity and weights of the items. First, the mean score across all items was calculated (per domain). Then, the summary scores were summed and converted into a metric ranging from 0 to 100. A score in this questionnaire was obtained for the participants who had answered to at least 80% of the items (i.e., at least 10 out of the 12 items). In those cases with one or two missing values, the score were obtained based on the items completed, replacing the missing values by the mean of the remaining items.

# Syntax used to create the variable (Stata 13)

```{stata load-data}

* Cohorte 2019, linea base:
use "~\UAM\marta.miret@uam.es - Bases de datos maestras Edad con Salud\Ola_3\Cohorte_2019\rawdata_c2019w1.dta"

```

```{stata generate-outcomes, echo=TRUE}

/*WHODAS-12 questionnaire*/

gen S1=q2028_stand
gen S2=q2032_hh_resp
gen S3=q2011_learn
gen S4=q2033_activ
gen S5=q2047_affect
gen S6=q2035_concent
gen S7=q2036_walk
gen S8=q2037_wash
gen S9=q2038_dress
gen S10=q2015_strang
gen S11=q2014_friend
gen S12=q2039_daily
recode S1 S2 S3 S4 S5 S6 S7 S8 S9 S10 S11 S12 (888=.) 
recode S1 S3 S5 S6 S7 S9 (1=0) (2=1) (3=2) (4=3) (5=4)
recode S2 S4 S8 S10 S11 S12 (1=0) (2=1) (3=1) (4=2) (5=2)
egen missing=rowmiss(S1-S12)
egen part1=rowmean(S1 S3 S5 S6 S7 S9) if missing <=2
egen part2=rowmean(S2 S4 S8 S10 S11 S12) if missing <=2
gen whodas12=(part1*25+part2*50)/2
label variable whodas12 "disability WHODAS 12 (c2019_w1)"

/*ADL and IADL*/

gen dq2025=q2025_sitting
gen dq2026=q2026_walk
gen dq2027=q2027_stand
gen dq2028=q2028_stand
gen dq2029=q2029_climb
gen dq2030=q2030_crouch
gen dq2031=q2031_pickup
gen dq2032=q2032_hh_resp
gen dq2033=q2033_activ
gen dq2034=q2034_reach
gen dq2035=q2035_concent
gen dq2036=q2036_walk
gen dq2037=q2037_wash
gen dq2038=q2038_dress
gen dq2039=q2039_daily
gen dq2040=q2040_carry
gen dq2041=q2041_movement
gen dq2042=q2042_eating
gen dq2043=q2043_arising
gen dq2044=q2044_toilet
gen dq2045=q2045_movement
gen dq2046=q2046_going
recode dq2025-dq2046 (1=0)(2=0)(3=0)(4=1)(5=1) (888=.)

egen adlcounter=rowtotal(dq2025 dq2026 dq2027 dq2028 dq2029 dq2030 dq2031 dq2034 dq2035 dq2036 dq2037 dq2038 dq2040 dq2041 dq2042 dq2043 dq2044)
egen missing_adl=rowmiss (dq2025 dq2026 dq2027 dq2028 dq2029 dq2030 dq2031 dq2034 dq2035 dq2036 dq2037 dq2038 dq2040 dq2041 dq2042 dq2043 dq2044)
gen adl=1 if adlcounter>=1 & adlcounter<.
replace adl=0 if adlcounter==0
replace adl=. if missing_adl==17
label variable adl "activities of daily living (c2019_w1)"

egen iadlcounter=rowtotal(dq2032 dq2033 dq2039 dq2045 dq2046)
egen missing_iadl=rowmiss (dq2032 dq2033 dq2039 dq2045 dq2046)
gen iadl=1 if iadlcounter>=1 & iadlcounter<.
replace iadl=0 if iadlcounter==0
replace iadl=. if missing_iadl==5
label variable iadl "instrumental act of daily living (c2019_w1)"

label define dAdl 0 "No" 1 "Yes"
label values adl iadl dAdl 

```

```{stata descriptives, collectcode=FALSE, results='markup', echo=TRUE}

sum whodas12, detail
codebook adl iadl

```

```{stata save-data}
keep ID_ECS q0007a_result proxy subsample_pre interview_pre whodas12 adl iadl

*Cohorte 2019, linea base:

saveold "~\UAM\marta.miret@uam.es - Documentacion Edad con Salud\Edad con salud - Ola 3\Outcomes\Cohorte 2019\Outcome datasets\Outcome_disability.dta", replace version(13)


```

# Contact person:

-   Name: Elvira Lara
-   e-mail: [elvira.lara\@uam.es](mailto:elvira.lara@uam.es){.email}
-   Phone number: (+34) 914974601

<br>

Original version:

-   Name: Félix Caballero 
-   e-mail: [felix.caballero@uam.es](mailto:felix.caballero@uam.es){.email}
-   Phone number: (+34) 914975088

# References

World Health Organization (2010).Measuring health and disability: manual for WHO Disability Assessment Schedule (WHODAS 2.0).Edited by T.B. Üstün, N. Kostanjsek, S. Chatterji and J. Rehm. World Health Organization Disability Assessment Schedule II (WHODAS-II). Available at http://whqlibdoc.who.int/publications/2010/9789241547598_eng.pdf

# Session info

```{r, results='markup'}
devtools::session_info()
```
