---
title: "Edad con Salud -- cohort 2011 -- Wave 3 -- 'Depression ICD-10'"
output:
  word_document:
    reference_docx: !expr ecs.data::get_pkg_docx_template()
csl: "`r ecs.data::get_proj_csl_file()`"
# bibliography: "`r ecs.data::get_proj_bib_file()`"
bibliography: "../../../../www/Edad_con_salud_outcome_vars.bib"
editor_options: 
  markdown: 
    wrap: 80
---

```{r setup, include=FALSE}
# Activar los paquetes que R necesita para ejecutar el c贸digo:

library(Statamarkdown)

# Ruta al ejecutable de Stata (Ejecuta `sysdir` en Stata para encontrar la ruta)

# Opciones de configuraci贸n de las salidas de R y Stata:

knitr::opts_chunk$set( # Para todos los 'chunks' de c贸digo:
  echo        = FALSE, #   - No muestra la sintaxis en el documento
  results     = 'hide',#   - Omite los resultados de ejecutar la sintaxis
  cleanlog    = TRUE,  #   - Omite repetir sintaxis en resultados de Stata
  collectcode = TRUE   #   - Reutiliza el c贸digo de Stata de 'chunks' anteriores
)
```

# Variable information

+---------------------------+----------------+------------+------------+
| **Main variable name**    | **Label**      | **Format** | **Code**   |
+:==========================+:===============+============+============+
| depression_lifetime_s2    | Lifetime       | Binary     | -   0=no;  |
|                           | depression     |            | -   1=yes  |
|                           | (c2011_w3)     |            |            |
+---------------------------+----------------+------------+------------+
| depression_12m_s2         | 12 month       | Binary     | -   0=no;  |
|                           | depression     |            | -   1=yes  |
|                           | (c2011_w3)     |            |            |
+---------------------------+----------------+------------+------------+
| depression_12treatment_s2 | Treatment for  | Binary     | -   0=no;  |
|                           | depression in  |            | -   1=yes  |
|                           | previous 12    |            |            |
|                           | monts          |            |            |
|                           | (c2011_w3)     |            |            |
+---------------------------+----------------+------------+------------+
| depress_self_s2           | depression,    | Binary     | -   0=no;  |
|                           | self reported  |            | -   1=yes  |
|                           | (c2011_w3)     |            |            |
+---------------------------+----------------+------------+------------+

# Variable description

Depression is a common mental disorder, characterized by sadness, loss of
interest or pleasure, feelings of guilt or low self-worth, disturbed sleep or
appetite, feelings of tiredness, and poor concentration. Depression can be
long-lasting or recurrent, substantially impairing an individual's ability to
function at work or school or cope with daily life. At its most severe,
depression can lead to suicide. When mild, people can be treated without
medicines but when depression is moderate or severe they may need medication and
professional talking treatments. Depression is a disorder that can be reliably
diagnosed and treated by non-specialists as part of primary health care.
Specialist care is needed for a small proportion of individuals with complicated
depression or those who do not respond to first-line treatments.

We used an adapted version of the Composite International Diagnostic Interview
(CIDI 3.0), a comprehensive, fully-structured interview designed to be used by
trained lay interviewers for the assessment of mental disorders according to the
definitions and criteria of DSM-V [@americanpsychiatricassociation2013]. It is
intended for use in epidemiological and cross-cultural studies as well as for
clinical and research purposes. The diagnostic section of the interview is based
on the World Health Organization's Composite International Diagnostic Interview.

# Explanation of how the variables were created

An algorithm based on the DSM-V was used for the endorsement of major depressive
episode at some point in their lives (lifetime prevalence) and in the previous
12 months (point prevalence). We also included a treatment variable for 12-month
depression. These algorithms were created only for the non-proxy respondents.

As for the proxy participants, original variables include `p4040_depression_s2`
and `p4040a_age_s2`. A diagnosis of depression was endorsed if the participant
answered "yes" to `p4040_depression_s2`, but not `997` in `p4040a_age_s2` (they
affirmed they had never had depression). We also included a treatment code
(`depress_12treatment_s2`) in those individuals who, apart from having reported
depression, have received treatment for it in the previous 12 months.

# Syntax used to create the variable (Stata 15)

```{stata load-data}
* Cohorte 2011, ola 3:
use "~\UAM\Marta Miret Garcia - Bases de datos maestras Edad con Salud\Ola_3\Cohorte_2011\FINAL\20190208_PES16028242_Seguimiento_final_completo_V6.dta"
```

```{stata generate-outcomes, echo=TRUE}

```

<br>

```{stata descriptives, collectcode=FALSE, results='markup'}
codebook depression_lifetime_s2 depression_12m_s2 depression_12treatment_s2 depress_self_s2
```

```{stata save-data}
keep ID_ECS q0007a_result proxy codebook depression_lifetime_s2 depression_12m_s2 depression_12treatment_s2 depress_self_s2

* Cohorte 2011, ola 3:
*saveold "~\UAM\Marta Miret Garcia - Documentacion Edad con Salud\Edad con salud - Ola 3\Outcomes\Cohorte 2011\Outcome datasets\Outcome_depression_DSM5.dta", replace version(13)
saveold "../../../../dat/Outcome_depression_DSM5.dta", replace version(13)
*# TODO: Delete previous line and uncomment the `saveold` command to the proper
*#       output path
```

<br>

# Contact person:

-   Name: Beatriz Olaya
-   e-mail: [beatriz.olaya\@pssjd.org](mailto:beatriz.olaya@pssjd.org)
-   Phone number: (+34) 936406350 ext (1) 2546

# References

::: {#refs}
:::

\newpage

# Session info

```{r, results='markup'}
devtools::session_info()
```
